---
title: "CDS Calculations for Paper Results"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(monocle3)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(gridExtra)
library(grid)
```

## Load the Final KRAB CDS Object

```{r load_data}
cds_path <- "./outputs/targeted_cCRE_combined_cds_all_genes_final.RDS"

full_experiment_cds <- readRDS(cds_path)
```

## Preliminary Checks

```{r check_metadata}
cds_column_data <- colData(full_experiment_cds) %>% as.data.frame()
nrow(cds_column_data)
```
Calculate the number of distinct sgRNAs detected in the targeted cCRE experiment:
```{r}
sgRNA_distinct <- length(unique(cds_column_data$sgRNA))
sgRNA_distinct

sgRNA_sequence_distinct <- length(unique(cds_column_data$sgRNA.sequence))
sgRNA_sequence_distinct

sgRNA_distinct == sgRNA_sequence_distinct
```
Calculate the total number of distinct genetic and chemical perturbations in the experiment:
```{r}
gxe_perturbations <- sgRNA_distinct * length(unique(cds_column_data$dose))
gxe_perturbations

gxe_perturbations == sgRNA_distinct*4
```

## Cell Counts per Dose

```{r cell_counts_per_dose}
# Count number of cells per dose
cells_per_dose <- cds_column_data %>%
  group_by(dose) %>%
  summarise(n_cells = n(), .groups = 'drop')

# Calculate median and mean number of cells per dose
median_cells_per_dose <- median(cells_per_dose$n_cells)
mean_cells_per_dose <- mean(cells_per_dose$n_cells)

# Format the table for printing and saving as a PDF
cells_per_dose_table <- cells_per_dose %>%
  mutate(dose = factor(dose, levels = c(0, 10, 50, 100))) %>%
  arrange(dose) %>%
  rename("TMZ dose (μM)" = dose, "cell count" = n_cells)

print(cells_per_dose_table)

print(paste("Median number of cells per dose:", median_cells_per_dose))
print(paste("Mean number of cells per dose:", mean_cells_per_dose))

tg <- tableGrob(cells_per_dose_table, theme = ttheme_default(base_size = 7))
width_in <- convertWidth(sum(tg$widths), "inches", valueOnly = TRUE)
height_in <- convertHeight(sum(tg$heights), "inches", valueOnly = TRUE)

cairo_pdf("./outputs/plots/cell_counts_per_dose.pdf", width = width_in + 0.2, height = height_in + 0.2)
grid.newpage()
grid.draw(tg)
dev.off()
```

## Cell Counts per sgRNA

```{r cell_counts_per_sgrna}
# Count number of cells per sgRNA
cells_per_sgrna <- cds_column_data %>%
  group_by(sgRNA) %>%
  summarise(n_cells = n(), .groups = 'drop')

# Calculate median and mean number of cells per sgRNA
median_cells_per_sgrna <- median(cells_per_sgrna$n_cells)
mean_cells_per_sgrna <- mean(cells_per_sgrna$n_cells)

print(paste("Median number of cells per sgRNA:", median_cells_per_sgrna))
print(paste("Mean number of cells per sgRNA:", mean_cells_per_sgrna))
```

## Cell Counts per ATAC_bin (this excludes exons and control groups so it is counting the mean and median sgRNAs per c-CRE)

```{r cell_counts_per_atac_bin}
# Define controls to exclude
controls_to_exclude <- c("AAVS1.control", "random.controls", "NTC")

# Count number of cells per ATAC_bin, excluding controls and exon targeting
cells_per_atac_bin <- cds_column_data %>%
  filter(!(ATAC_bin %in% controls_to_exclude)) %>%
  filter(!str_detect(df_name, "MSH2_exons|MSH6_exons")) %>%
  group_by(ATAC_bin) %>%
  summarise(n_cells = n(), .groups = 'drop')

# Calculate median and mean number of cells per ATAC_bin
median_cells_per_atac_bin <- median(cells_per_atac_bin$n_cells)
mean_cells_per_atac_bin <- mean(cells_per_atac_bin$n_cells)

print(paste("Median number of cells per ATAC_bin (excluding controls and msh2/msh6 exons):", median_cells_per_atac_bin))
print(paste("Mean number of cells per ATAC_bin (excluding controls and msh2/msh6 exons):", mean_cells_per_atac_bin))
```

## Gene Detection Analysis

Using `detect_genes` to count how many cells express each transcript (min_expr > 0) in the full experiment.

```{r detect_genes}
full_experiment_cds <- detect_genes(full_experiment_cds, min_expr = 0)

# Extract and sort gene metadata by detection frequency
gene_metadata <- rowData(full_experiment_cds) %>% 
  as.data.frame() %>%
  arrange(desc(num_cells_expressed)) %>%
  mutate(rank = row_number(),
         percentile_rank = round((rank / n()) * 100, 2))
```

## Gene Rank Analysis

Determining the "detectability rank" of key genes of interest in the full KRAB experiment.

```{r rank_analysis}
genes_to_check <- c("MSH2", "MSH6", "B2M", "PGK1", "HPRT1")

ranking_results <- gene_metadata %>%
  filter(gene_short_name %in% genes_to_check) %>%
  dplyr::select(gene_short_name, num_cells_expressed, rank, percentile_rank)

print(ranking_results)

# Create a new dataframe for PDF table formatting
ranking_results_table <- ranking_results %>%
  dplyr::select(gene_short_name, rank, percentile_rank) %>%
  rename(gene = gene_short_name, `percentile rank` = percentile_rank)

# Save table as PDF with 7pt font
tg_rank <- tableGrob(ranking_results_table, theme = ttheme_default(base_size = 7))
width_in <- convertWidth(sum(tg_rank$widths), "inches", valueOnly = TRUE)
height_in <- convertHeight(sum(tg_rank$heights), "inches", valueOnly = TRUE)

cairo_pdf("./outputs/plots/gene_ranking_results.pdf", width = width_in + 0.2, height = height_in + 0.2)
grid.newpage()
grid.draw(tg_rank)
dev.off()
```

