---
title: "100cells_150bpwindow_slide25_targeted_cCRE_sliding_window_analysis"
output: html_document
date: "2025-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(Biostrings)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(ggplot2)
library(htmlwidgets)
library(plotly)
library(SummarizedExperiment)
library(Matrix)
library(DESeq2)
library(monocle3)
library(boot.pval)
```



```{r}
working_dir <- "/Users/annaschoonen/Documents/JMF_lab_project/aim2/combined_analysis_of_all_working_KRAB_premade_probe_targeted_sci_experiments_feb_may_june/true_sliding_windows_with_overlap/bootstrap_sliding_windows/updated_bootstrap_code_9_4_25/rerun_same_code_1000bootstrap_150bpwindow_slide25_100cells/analysis_of_cluster_results/"

output_dir <- "./outputs/sliding_window_100cells/"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

paper_plots_dir <- "./outputs/sliding_window_100cells/final_plots_for_paper_main_and_supplement/"
dir.create(paper_plots_dir, recursive = TRUE, showWarnings = FALSE)
```




```{r}
# loading results from cluster back in here - Define the file path
file_path <- "../data/HPC_outputs/100cells_msh2_sliding_window_1000bootstrap_results.csv"

# Read the CSV into a data frame called 'results'
modeling_results <- read.csv(file_path)

# Calculate Benjamini-Hochberg Adjusted P-value on raw, full data (only for successful models)
modeling_results <- modeling_results %>%
  mutate(
    adjusted_p_value = if_else(
      status == "OK" & !is.na(median_p_value),
      p.adjust(median_p_value, method = "BH"),
      as.numeric(NA)
    )
  )

# Now, add the 'window' column to your dataframe
modeling_results <- modeling_results %>%
  mutate(
    window = if_else(
      # This is the condition to check
      window_scheme == "NA_control_window",

      # Value if the condition is TRUE
      ATAC_bin,

      # Value if the condition is FALSE
      # This part extracts the chromosome and pastes it with the window start/end
      paste0(sub(":.*", "", ATAC_bin), ":", window_start, "-", window_end)
    )
  )

# You can view the first few rows of your updated dataframe to check the result
head(modeling_results)

# Now, add the 'window' column to your dataframe
modeling_results <- modeling_results %>%
  mutate(
    window = if_else(
      # This is the condition to check
      window_scheme == "NA_control_window",

      # Value if the condition is TRUE
      ATAC_bin,

      # Value if the condition is FALSE
      # This part extracts the chromosome and pastes it with the window start/end
      paste0(sub(":.*", "", ATAC_bin), ":", window_start, "-", window_end)
    )
  )


# Relocate the 'window' column
modeling_results <- modeling_results %>%
  relocate(window, .after = window_end)

# You can check the new column order
colnames(modeling_results)

# You can view the first few rows of your updated dataframe to check the result
head(modeling_results)
dim(modeling_results)


# Create a data frame for control bins, sorted alphabetically
control_bins <- modeling_results %>%
  filter(is.na(window_start)) %>%
  arrange(window)

# Create a data frame for experimental bins, sorted by genomic position
experimental_bins <- modeling_results %>%
  filter(!is.na(window_start)) %>%
  arrange(window_start)

# Combine the two sets to create the final, ordered list for the x-axis
final_bin_order <- c(unique(control_bins$window), unique(experimental_bins$window))

# Convert the 'window' column into an ordered factor
modeling_results$window <- factor(modeling_results$window, levels = final_bin_order)

# Convert the 'dose' column to a factor with numerically sorted levels
# This will place the highest dose (100) at the top and the lowest (0) at the bottom.
modeling_results$dose <- factor(
  modeling_results$dose,
  levels = sort(unique(modeling_results$dose), decreasing = TRUE)
)

modeling_results_raw <- modeling_results
```

## Subset Significant Windows by Adjusted P-value

```{r subset_sig_p_val}
# Isolated subset of modeling_results_raw for windows with adjusted p-value < 0.05
sig_windows_by_pval_only <- modeling_results_raw %>%
  filter(!is.na(adjusted_p_value) & adjusted_p_value < 0.05)

# Count and report
num_sig_pval_only <- nrow(sig_windows_by_pval_only)
print(paste("Number of windows with adjusted p-value < 0.05:", num_sig_pval_only))

# Further subset for MSH2 exon 1 specifically (stricter 0.01 p-value cutoff)
MSH2exon1_sig_windows_by_pval_only <- modeling_results_raw %>%
  filter(!is.na(adjusted_p_value) & adjusted_p_value < 0.01) %>%
  filter(ATAC_bin == "chr2:47403155-47403402")

# Count and report for MSH2 exon 1
print(paste("Number of MSH2 exon 1 windows with adjusted p-value < 0.01:", nrow(MSH2exon1_sig_windows_by_pval_only)))

# Save modeling_results_raw as a csv
write.csv(modeling_results_raw,
  file = file.path(paper_plots_dir, "Supp_table_modeling_results_raw_100cells_v7script.csv"),
  row.names = FALSE
)
```

## Pivot Table for MSH2 Exon 1 Significant Windows

```{r pivot_table_msh2_exon1}
# Prepare and pivot the table
MSH2exon1_sig_windows_pivot <- MSH2exon1_sig_windows_by_pval_only %>%
  # Ensure numeric ordering by window_start
  arrange(as.numeric(window_start)) %>%
  # Select only the relevant columns for the table
  dplyr::select(window, dose, adjusted_p_value) %>%
  # Pivot wider to show doses as columns
  tidyr::pivot_wider(names_from = dose, values_from = adjusted_p_value)

# Display the table
print("MSH2 Exon 1 Significant Windows (adj. p-value < 0.01) by Dose:")
print(MSH2exon1_sig_windows_pivot)
```


```{r}
# Force the top 5% threshold to match the stricter 40 cell threshold (v6 script)
top_5_percent_threshold <- 0.5721355

# Print the result
print(top_5_percent_threshold)
```


```{r}


# Create the count-based histogram plot (no density line)
p <- ggplot(modeling_results, aes(x = mean_normalized_effect)) +
  # Add a histogram with counts on the y-axis (default)
  geom_histogram(bins = 100, fill = "dimgrey", alpha = 0.7) +

  # Add a vertical line for the lower threshold (-top 5%)
  geom_vline(
    xintercept = -top_5_percent_threshold,
    color = "blue",
    linetype = "dashed",
    size = 1
  ) +

  # Add a vertical line for the upper threshold (+top 5%)
  geom_vline(
    xintercept = top_5_percent_threshold,
    color = "red",
    linetype = "dashed",
    size = 1
  ) +

  # Add labels and a title
  labs(
    title = "Distribution of Mean Normalized Effect - 100 cell threshold",
    x = "Mean Normalized Effect",
    y = "Count",
    caption = paste0("Dashed lines indicate (+/- ", round(top_5_percent_threshold, 3), ")")
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(size = 7),
    axis.title.x = element_text(size = 7),
    axis.title.y = element_text(size = 7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    plot.caption = element_text(size = 6)
  )

# Print the plot
p

# Save the plot as a PDF with specific dimensions and font sizes
pdf_dir <- "./outputs/sliding_window_100cells/distribution_mean_norm_effect_modeling_results/"
dir.create(pdf_dir, recursive = TRUE, showWarnings = FALSE)
ggsave(
  filename = file.path(pdf_dir, "100cells_dist_mean_norm_effect.pdf"),
  plot = p,
  width = 3.1,
  height = 3.0,
  units = "in"
)

# Clean up variables to prevent any side effects on downstream chunks
rm(p, pdf_dir)
```






```{r}

# ---  Enhanced function to generate distinct colors (No change) ---
generate_distinct_colors <- function(n) {
  golden_ratio_conjugate <- 0.61803398875
  h <- runif(1)
  hues <- (h + (1:n) * golden_ratio_conjugate) %% 1
  sv_pairs <- list(
    c(s = 0.9, v = 0.85), c(s = 0.4, v = 0.95), c(s = 0.9, v = 0.6),
    c(s = 0.25, v = 0.98), c(s = 0.7, v = 0.7)
  )
  s_values <- rep(sapply(sv_pairs, `[`, 1), length.out = n)
  v_values <- rep(sapply(sv_pairs, `[`, 2), length.out = n)
  hsv(h = hues, s = s_values, v = v_values)
}

# ---  Create a High-Contrast Master Color Key (No change) ---
ordered_bins <- modeling_results %>%
  arrange(window) %>%
  pull(ATAC_bin) %>%
  unique()
num_bins <- length(ordered_bins)
distinct_colors <- generate_distinct_colors(n = num_bins)
atac_color_map <- data.frame(
  ATAC_bin = ordered_bins,
  bin_color = distinct_colors
)

# ---  Join the Color Map to the Main Dataframe (No change) ---
if (!"bin_color" %in% names(modeling_results)) {
modeling_results <- modeling_results %>%
  left_join(atac_color_map, by = "ATAC_bin")
}

# --- NEW: Define the Custom Color Scale for the Lollipops with new range and color ---
min_effect <- min(modeling_results$mean_normalized_effect, na.rm = TRUE)
max_effect <- max(modeling_results$mean_normalized_effect, na.rm = TRUE)
range_effect <- max_effect - min_effect

# Calculate symmetrical limits for color scale based on ACTUAL data range
actual_max_abs <- max(abs(range(modeling_results$mean_normalized_effect, na.rm = TRUE)))
# Ensure we cover at least the cap to avoid errors
actual_max_abs <- max(actual_max_abs, 1.5)

# Define the cap and threshold
cap_val <- 1.0
thresh_val <- top_5_percent_threshold

# Calculate normalized positions (0 to 1) for the gradient stops
# The scale spans from -actual_max_abs to +actual_max_abs
total_span <- 2 * actual_max_abs

# Function to normalize a value to [0, 1] given the range [-actual_max_abs, actual_max_abs]
norm_pos <- function(val) {
  (val - (-actual_max_abs)) / total_span
}

v_neg_cap <- norm_pos(-cap_val)
v_neg_thresh <- norm_pos(-thresh_val)
v_pos_thresh <- norm_pos(thresh_val)
v_pos_cap <- norm_pos(cap_val)

# Create the custom colorscale: blue -> blue -> lightgrey -> lightgrey -> red -> red
custom_colorscale <- list(
  list(0, "blue"),
  list(v_neg_cap, "blue"),
  list(v_neg_thresh, "lightgrey"),
  list(v_pos_thresh, "lightgrey"),
  list(v_pos_cap, "red"),
  list(1, "red")
)

# ---  Build the Header Plot (No change) ---
header_data <- modeling_results %>%
  dplyr::select(window, ATAC_bin, bin_color) %>%
  distinct()
p_header <- plot_ly(
  data = header_data,
  x = ~window, y = 1, type = "bar",
  marker = list(color = ~bin_color),
  hoverinfo = "text", text = ~ paste("ATAC Bin:", ATAC_bin)
) %>% layout(yaxis = list(showticklabels = FALSE, title = ""))

# ---  Build the Main Lollipop Plots (No change in this section, it uses custom_colorscale) ---
dose_plots <- list()
dose_levels <- levels(modeling_results$dose)
for (d in dose_levels) {
  dose_data <- filter(modeling_results, dose == d)
  dose_data_na <- filter(dose_data, is.na(mean_normalized_effect))
  dose_data_non_na <- filter(dose_data, !is.na(mean_normalized_effect))

  p_dose <- plot_ly(showlegend = FALSE) %>%
    add_segments(
      data = dose_data_non_na,
      x = ~window, y = 0, xend = ~window, yend = ~mean_normalized_effect,
      color = I("gray"), hoverinfo = "none"
    ) %>%
    add_markers(
      data = dose_data_non_na,
      x = ~window, y = ~mean_normalized_effect,
      marker = list(
        color = ~mean_normalized_effect, colorscale = custom_colorscale,
        cmin = -actual_max_abs, cmax = actual_max_abs,
        showscale = (d == "100"),
        colorbar = list(title = "Norm Effect"),
        size = 10
      ),
      hoverinfo = "text", text = ~ paste0(
        "Bin: ", ATAC_bin, "<br>",
        "Window: ", window, "<br>",
        "Scheme: ", window_scheme, "<br>",
        "Effect: ", round(mean_normalized_effect, 3), "<br>",
        "Adj. P-val: ", ifelse(is.na(adjusted_p_value), "NA", format.pval(adjusted_p_value, digits = 3)), "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    add_markers(
      data = dose_data_na, x = ~window, y = 0,
      marker = list(color = "black", size = 10),
      hoverinfo = "text", text = ~ paste0(
        "Bin: ", ATAC_bin, "<br>",
        "Window: ", window, "<br>",
        "Scheme: ", window_scheme, "<br>",
        "Effect: NA", "<br>",
        "Adj. P-val: ", ifelse(is.na(adjusted_p_value), "NA", format.pval(adjusted_p_value, digits = 3)), "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    add_annotations(
      text = paste(d), x = 1, y = 0.5,
      xref = "paper", yref = "paper",
      xanchor = "left", showarrow = FALSE,
      font = list(size = 14)
    )
  dose_plots[[as.character(d)]] <- p_dose
}

# ---  Combine All Plots and Enforce Order (No change) ---
final_plot <- subplot(
  p_header, dose_plots[["100"]], dose_plots[["50"]],
  dose_plots[["10"]], dose_plots[["0"]],
  nrows = 5, heights = c(0.1, 0.225, 0.225, 0.225, 0.225),
  shareX = TRUE, titleY = TRUE
) %>%
  layout(
    xaxis = list(
      title = "Genomic Window",
      categoryorder = "array",
      categoryarray = final_bin_order,
      tickangle = 60
    ),
    title = list(text = "Dose-Dependent Effect Across Genomic Windows-1000 bootstrap 150bp window", y = 0.99)
  )

# ---  Save the Final Interactive Plot ---
htmlwidgets::saveWidget(
  widget = final_plot,
  file = file.path(paper_plots_dir, "cells100_all_modeling_results_interactive_lollipop_final.html")
)
```


```{r}
# ---  Save the Global Static Lollipop Plot (All Modeling Results) ---
# This replicates the style of the significant windows version but for the complete dataset.

#  Prepare Data using the full results set
modeling_results_static_all <- modeling_results %>%
  mutate(dose = factor(as.character(dose), levels = c("100", "50", "10", "0"))) %>%
  mutate(window = factor(as.character(window), levels = final_bin_order))

#  Calculate dynamic scale limits (based on ALL results)
all_cell_range <- range(modeling_results_static_all$cell_count, na.rm = TRUE)
min_all_cells <- all_cell_range[1]
max_all_cells <- all_cell_range[2]

#  Define the Color Scale (Using the same logic as the interactive plot)
if (!exists("actual_max_abs") || is.null(actual_max_abs)) {
  actual_max_abs <- max(abs(range(modeling_results_static_all$mean_normalized_effect, na.rm = TRUE)))
  actual_max_abs <- max(actual_max_abs, 1.5)
}
cap_val_all <- 1.0
thresh_val_all <- top_5_percent_threshold
total_span_all <- 2 * actual_max_abs

norm_pos_all <- function(val) {
  (val - (-actual_max_abs)) / total_span_all
}

v_neg_cap_all <- norm_pos_all(-cap_val_all)
v_neg_thresh_all <- norm_pos_all(-thresh_val_all)
v_pos_thresh_all <- norm_pos_all(thresh_val_all)
v_pos_cap_all <- norm_pos_all(cap_val_all)

#  Create the Static Lollipop Plot
p_lollipop_all_static <- ggplot(modeling_results_static_all, aes(x = window)) +
  geom_linerange(
    data = subset(modeling_results_static_all, !is.na(mean_normalized_effect)),
    aes(ymin = 0, ymax = mean_normalized_effect),
    color = "gray"
  ) +
  geom_point(
    data = subset(modeling_results_static_all, !is.na(mean_normalized_effect)),
    aes(y = mean_normalized_effect, color = mean_normalized_effect),
    size = 2.0, alpha = 0.9
  ) +
  geom_point(
    data = subset(modeling_results_static_all, is.na(mean_normalized_effect)),
    aes(y = 0),
    color = "black",
    size = 2.0
  ) +
  facet_grid(dose ~ ., switch = "y", labeller = as_labeller(function(x) paste0(x, "~mu*'M'"), default = label_parsed)) +
  scale_color_gradientn(
    colors = c("blue", "blue", "lightgrey", "lightgrey", "red", "red"),
    values = c(0, v_neg_cap_all, v_neg_thresh_all, v_pos_thresh_all, v_pos_cap_all, 1),
    limits = c(-actual_max_abs, actual_max_abs),
    name = expression(beta[coef])
  ) +
  # Differentiate legend font sizes
  guides(
    color = guide_colorbar(title.theme = element_text(size = 14))
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.y.left = element_text(angle = 90, size = 13),
    strip.placement = "outside",
    legend.text = element_text(size = 10),
    legend.position = "right",
    panel.spacing = unit(1, "cm"),
    plot.margin = margin(t = 0, r = 0, b = 1, l = 0, unit = "cm")
  ) +
  labs(y = NULL) +
  scale_x_discrete(limits = final_bin_order, drop = FALSE) +
  coord_cartesian(expand = FALSE, clip = "off")

#  Create Header (Matches the interactive p_header but in ggplot)
header_data_static_all <- header_data %>%
  mutate(window = factor(as.character(window), levels = final_bin_order))

p_header_static_all <- ggplot(header_data_static_all, aes(x = window, y = 1, fill = ATAC_bin)) +
  geom_tile() +
  scale_fill_manual(values = setNames(header_data_static_all$bin_color, header_data_static_all$ATAC_bin)) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    plot.margin = margin(b = 0, unit = "cm")
  ) +
  coord_cartesian(expand = FALSE)

# Combine and Save
if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  combined_lollipop_all_static <- p_header_static_all / p_lollipop_all_static +
    plot_layout(heights = c(1, 10), guides = "collect")
} else {
  library(gridExtra)
  combined_lollipop_all_static <- grid.arrange(p_header_static_all, p_lollipop_all_static, ncol = 1, heights = c(1, 10))
}

ggsave(
  filename = file.path(paper_plots_dir, "Final_Supp_fig_cells100_all_modeling_results_static_lollipop_final.pdf"),
  plot = combined_lollipop_all_static,
  width = 15, height = 10
)

ggsave(
  filename = file.path(paper_plots_dir, "Final_Supp_fig_cells100_all_modeling_results_static_lollipop_final.png"),
  plot = combined_lollipop_all_static,
  width = 15, height = 10, dpi = 300
)
```


```{r}
head(modeling_results)
```



```{r}
# SET YOUR THRESHOLD HERE FOR ENTIRE SCRIPT
mean_norm_effect_threshold <- top_5_percent_threshold
```


```{r}


# The result is assigned to the new dataframe name
modeling_results_conf_int <- modeling_results %>%
  mutate(
    confidence_interval_contain_zero = if_else(
      # This condition checks if the two numbers cross zero
      conf_low_95 * conf_high_95 <= 0,

      # Value if TRUE (they have different signs or one is zero)
      "yes_contains_0",

      # Value if FALSE (they have the same sign)
      "NO"
    )
  )

# Check the first few rows of the new dataframe
head(modeling_results_conf_int)
```





keep only windows that pass both the confidence interval and mean_normalized_effect for 2 subsequent windows for sliding window or just pass both filters if small or control window

```{r}
# --- Part 1: Filtering to create 'significant_windows_only' dataframe ---
# (This section assumes your 'modeling_results_conf_int' dataframe is available)
sliding_windows_to_keep <- modeling_results_conf_int %>%
  filter(window_scheme == "sliding_window") %>%
  group_by(ATAC_bin, dose) %>%
  arrange(window_start) %>%
  filter(
    (
      (abs(mean_normalized_effect) >= mean_norm_effect_threshold & confidence_interval_contain_zero == "NO") &
        (abs(lead(mean_normalized_effect)) >= mean_norm_effect_threshold & lead(confidence_interval_contain_zero) == "NO")
    ) |
      (
        (abs(mean_normalized_effect) >= mean_norm_effect_threshold & confidence_interval_contain_zero == "NO") &
          (abs(lag(mean_normalized_effect)) >= mean_norm_effect_threshold & lag(confidence_interval_contain_zero) == "NO")
      )
  ) %>%
  ungroup() %>%
  pull(window) %>%
  unique()

small_windows_to_keep <- modeling_results_conf_int %>%
  filter(window_scheme == "small_window") %>%
  filter(abs(mean_normalized_effect) >= mean_norm_effect_threshold & confidence_interval_contain_zero == "NO") %>%
  pull(window) %>%
  unique()

control_windows_to_keep <- modeling_results_conf_int %>%
  filter(window_scheme == "NA_control_window") %>%
  filter(abs(mean_normalized_effect) >= mean_norm_effect_threshold & confidence_interval_contain_zero == "NO") %>%
  pull(window) %>%
  unique()

all_windows_to_keep <- c(
  sliding_windows_to_keep,
  small_windows_to_keep,
  control_windows_to_keep
) %>%
  unique()

significant_windows_only_w_random_controls <- modeling_results_conf_int %>%
  filter(window %in% all_windows_to_keep)

significant_windows_only <- significant_windows_only_w_random_controls %>%
  filter(ATAC_bin != "random.controls")

# ---  Prepare the 'significant_windows_only' Data for Plotting ---
control_bins_sig <- significant_windows_only %>%
  filter(is.na(window_start)) %>%
  arrange(window)
experimental_bins_sig <- significant_windows_only %>%
  filter(!is.na(window_start)) %>%
  arrange(window_start)
final_bin_order_sig <- c(
  unique(control_bins_sig$window),
  unique(experimental_bins_sig$window)
)
significant_windows_only$window <- factor(
  significant_windows_only$window,
  levels = final_bin_order_sig
)
significant_windows_only$dose <- factor(
  significant_windows_only$dose,
  levels = sort(unique(significant_windows_only$dose), decreasing = TRUE)
)

# ---  Load Manual Coloring from CSV ---
# We now use pre-assigned colors for global consistency
manual_colors_path <- "../data/ATAC_bin_color_map/manual_colors_ATAC_bins.csv"
manual_colors_df <- read.csv(manual_colors_path, stringsAsFactors = FALSE)

# Create the color map using ATAC_bin and final_colors from the CSV
atac_color_map <- manual_colors_df %>%
  dplyr::select(ATAC_bin, bin_color = final_colors) %>%
  distinct()
# Force update of bin_color to use the new colors generated for the filtered set
if ("bin_color" %in% names(significant_windows_only)) {
  significant_windows_only <- significant_windows_only %>% dplyr::select(-bin_color)
}

significant_windows_only <- significant_windows_only %>%
  left_join(atac_color_map, by = "ATAC_bin")

# Export significant ATAC bins with colors for audit and comparison
atac_bin_color_map_for_audit <- significant_windows_only %>%
  dplyr::select(ATAC_bin, bin_color) %>%
  distinct()

write.csv(atac_bin_color_map_for_audit,
  file = file.path(output_dir, "significant_atac_bins_100cells_w_colors.csv"),
  row.names = FALSE
)

write.csv(significant_windows_only,
  file = "./outputs/sliding_window_100cells/significant_windows_only_100cells.csv",
  row.names = FALSE
)

# --- Define the Custom Color Scale ---
# Calculate symmetrical limits for color scale based on ACTUAL data range
actual_max_abs <- max(abs(range(significant_windows_only$mean_normalized_effect, na.rm = TRUE)))
# Ensure we cover at least the cap to avoid errors
actual_max_abs <- max(actual_max_abs, 1.5)

# Define the cap and threshold
cap_val <- 1.0
thresh_val <- mean_norm_effect_threshold

# Calculate normalized positions (0 to 1) for the gradient stops
# The scale spans from -actual_max_abs to +actual_max_abs
total_span <- 2 * actual_max_abs

# Function to normalize a value to [0, 1] given the range [-actual_max_abs, actual_max_abs]
norm_pos <- function(val) {
  (val - (-actual_max_abs)) / total_span
}

v_neg_cap <- norm_pos(-cap_val)
v_neg_thresh <- norm_pos(-thresh_val)
v_pos_thresh <- norm_pos(thresh_val)
v_pos_cap <- norm_pos(cap_val)

# Create the custom colorscale: blue -> blue -> lightgrey -> lightgrey -> red -> red
custom_colorscale <- list(
  list(0, "blue"),
  list(v_neg_cap, "blue"),
  list(v_neg_thresh, "lightgrey"),
  list(v_pos_thresh, "lightgrey"),
  list(v_pos_cap, "red"),
  list(1, "red")
)

# --- Build the Header and Lollipop Plots ---
header_data <- significant_windows_only %>%
  dplyr::select(window, ATAC_bin, bin_color) %>%
  distinct()
p_header <- plot_ly(
  data = header_data, x = ~window, y = 1, type = "bar",
  marker = list(color = ~bin_color),
  hoverinfo = "text", text = ~ paste("ATAC Bin:", ATAC_bin)
) %>%
  layout(yaxis = list(showticklabels = FALSE, title = ""))

dose_plots <- list()
dose_levels <- levels(significant_windows_only$dose)

for (d in dose_levels) {
  dose_data <- filter(significant_windows_only, dose == d)
  dose_data_na <- filter(dose_data, is.na(mean_normalized_effect))
  dose_data_non_na <- filter(dose_data, !is.na(mean_normalized_effect))

  p_dose <- plot_ly(showlegend = FALSE) %>%
    add_segments(
      data = dose_data_non_na, x = ~window, y = 0,
      xend = ~window, yend = ~mean_normalized_effect,
      color = I("gray"), hoverinfo = "none"
    ) %>%
    # --- THIS SECTION IS UPDATED ---
    add_markers(
      data = dose_data_non_na, x = ~window, y = ~mean_normalized_effect,
      marker = list(
        color = ~mean_normalized_effect, colorscale = custom_colorscale,
        cmin = -actual_max_abs, cmax = actual_max_abs,
        showscale = (d == "100"),
        colorbar = list(title = "Norm Effect"),
        size = 10
      ),
      hoverinfo = "text",
      # Create a detailed, multi-line tooltip
      text = ~ paste0(
        "Bin: ", ATAC_bin, "<br>",
        "Window: ", window, "<br>",
        "Scheme: ", window_scheme, "<br>",
        "Effect: ", round(mean_normalized_effect, 3), "<br>",
        "Adj. P-val: ", ifelse(is.na(adjusted_p_value), "NA", format.pval(adjusted_p_value, digits = 3)), "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "CI contains 0: ", confidence_interval_contain_zero, "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    # --- THIS SECTION IS ALSO UPDATED ---
    add_markers(
      data = dose_data_na, x = ~window, y = 0,
      marker = list(color = "black", size = 10),
      hoverinfo = "text",
      # Create a similar tooltip for NA values
      text = ~ paste0(
        "Bin: ", ATAC_bin, "<br>",
        "Window: ", window, "<br>",
        "Scheme: ", window_scheme, "<br>",
        "Effect: NA", "<br>",
        "Adj. P-val: ", ifelse(is.na(adjusted_p_value), "NA", format.pval(adjusted_p_value, digits = 3)), "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "CI contains 0: ", confidence_interval_contain_zero, "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    add_annotations(
      text = paste(d), x = 1, y = 0.5,
      xref = "paper", yref = "paper",
      xanchor = "left", showarrow = FALSE,
      font = list(size = 14)
    )

  dose_plots[[as.character(d)]] <- p_dose
}

# --- Combine All Plots and Enforce Order ---
final_plot_significant <- subplot(
  p_header, dose_plots[["100"]], dose_plots[["50"]],
  dose_plots[["10"]], dose_plots[["0"]],
  nrows = 5, heights = c(0.1, 0.225, 0.225, 0.225, 0.225),
  shareX = TRUE, titleY = TRUE
) %>%
  layout(
    xaxis = list(
      title = "Genomic Window",
      categoryorder = "array",
      categoryarray = final_bin_order_sig, # Use the new order for significant windows
      tickangle = 60
    ),
    title = list(text = "Dose-Dependent Effect for Significant Windows Only_detailed hover_FINAL", y = 0.99)
  )

# --- Save the Final Interactive Plot ---
htmlwidgets::saveWidget(
  widget = final_plot_significant,
  file = file.path(paper_plots_dir, "cells100_interactive_lollipop_nominal_hit_windows_detailed_hover_FINAL.html")
)
```








calculating # of sgRNAs from significant_windows_only

```{r}
all_sgRNA_across_significant_windows <- significant_windows_only %>%
  dplyr::select(ATAC_bin, window, window_start, window_end, sgRNA_per_window) %>%
  tidyr::separate_rows(sgRNA_per_window, sep = ",") %>%
  dplyr::rename(sgRNA.sequence = sgRNA_per_window)

write.csv(all_sgRNA_across_significant_windows,
  file = "./outputs/sliding_window_100cells/all_sgRNA_across_significant_windows_100cells.csv",
  row.names = FALSE
)

head(all_sgRNA_across_significant_windows)

# Get the final count of unique sgRNAs
num_unique_sgrnas <- length(unique(all_sgRNA_across_significant_windows$sgRNA.sequence))

print(paste("Total number of unique sgRNAs across all significant windows:", num_unique_sgrnas))
```




```{r}
# --- Static Heatmap ---

if (!requireNamespace("patchwork", quietly = TRUE)) {
  if (!requireNamespace("gridExtra", quietly = TRUE)) {
    stop("Please install 'patchwork' or 'gridExtra' to generate the static heatmap.")
  }
  library(gridExtra)
  use_patchwork <- FALSE
} else {
  library(patchwork)
  use_patchwork <- TRUE
}

static_data <- significant_windows_only

static_data$window <- factor(static_data$window, levels = final_bin_order_sig)
static_data$dose <- factor(static_data$dose, levels = sort(unique(static_data$dose), decreasing = FALSE))

header_data_static <- static_data %>%
  dplyr::select(window, ATAC_bin, bin_color) %>%
  distinct()

p_header_static <- ggplot(header_data_static, aes(x = window, y = 1, fill = ATAC_bin)) +
  geom_tile() +
  scale_fill_manual(values = setNames(atac_color_map$bin_color, atac_color_map$ATAC_bin)) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(b = 0, unit = "cm"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  coord_cartesian(expand = FALSE)
```


```{r}
# --- Static Lollipop Plot (Replicating Interactive Style) ---

static_data_lolli <- static_data %>%
  mutate(dose = factor(as.character(dose), levels = c("100", "50", "10", "0")))

actual_max_abs_lolli <- max(abs(range(static_data$mean_normalized_effect, na.rm = TRUE)))
actual_max_abs_lolli <- max(actual_max_abs_lolli, 1.5)
cap_val_lolli <- 1.0
thresh_val_lolli <- mean_norm_effect_threshold
total_span_lolli <- 2 * actual_max_abs_lolli

norm_pos_lolli <- function(val) {
  (val - (-actual_max_abs_lolli)) / total_span_lolli
}

v_neg_cap_lolli <- norm_pos_lolli(-cap_val_lolli)
v_neg_thresh_lolli <- norm_pos_lolli(-thresh_val_lolli)
v_pos_thresh_lolli <- norm_pos_lolli(thresh_val_lolli)
v_pos_cap_lolli <- norm_pos_lolli(cap_val_lolli)

p_lollipop_static <- ggplot(static_data_lolli, aes(x = window)) +
  geom_segment(
    data = subset(static_data_lolli, !is.na(mean_normalized_effect)),
    aes(xend = window, y = 0, yend = mean_normalized_effect),
    color = "gray"
  ) +
  # Points (colored by effect)
  geom_point(
    data = subset(static_data_lolli, !is.na(mean_normalized_effect)),
    aes(y = mean_normalized_effect, color = mean_normalized_effect),
    size = 2.0
  ) +
  # NA values (points at 0, #CCCCCC)
  geom_point(
    data = subset(static_data_lolli, is.na(mean_normalized_effect)),
    aes(y = 0),
    color = "black",
    size = 2.0
  ) +
  # Facet by Dose (stacked vertically) with microMolar unit mathematically parsed
  facet_grid(dose ~ ., switch = "y", labeller = as_labeller(function(x) paste0(x, "~mu*'M'"), default = label_parsed)) +
  # Custom Color Scale
  scale_color_gradientn(
    colors = c("blue", "blue", "lightgrey", "lightgrey", "red", "red"),
    values = c(0, v_neg_cap_lolli, v_neg_thresh_lolli, v_pos_thresh_lolli, v_pos_cap_lolli, 1),
    limits = c(-actual_max_abs_lolli, actual_max_abs_lolli),
    name = expression(beta[coef])
  ) +
  guides(
    color = guide_colorbar(title.theme = element_text(size = 14))
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.text.y.left = element_text(angle = 90, size = 13),
    strip.placement = "outside",
    legend.text = element_text(size = 10),
    legend.position = "right",
    panel.spacing = unit(1, "cm"),
    plot.margin = margin(t = 0, r = 1, b = 1, l = 1, unit = "cm")
  ) +
  labs(y = NULL) +
  coord_cartesian(expand = FALSE, clip = "off")

if (use_patchwork) {
  combined_lollipop_static <- p_header_static / p_lollipop_static +
    plot_layout(heights = c(1, 10), guides = "collect")
} else {
  combined_lollipop_static <- grid.arrange(p_header_static, p_lollipop_static, ncol = 1, heights = c(1, 10))
}

ggsave(
  filename = file.path(output_dir, "cells100_static_lollipop_significant_windows_inverted_v5.pdf"),
  plot = combined_lollipop_static,
  width = 12, height = 8 
)

ggsave(
  filename = file.path(output_dir, "cells100_static_lollipop_significant_windows_inverted_v5.png"),
  plot = combined_lollipop_static,
  width = 12, height = 8, dpi = 300
)
```

```{r}
# --- Static Lollipop Plot with lollipop scaled to cell number (Replicating Interactive Style) ---

static_data_lolli <- static_data %>%
  mutate(dose = factor(as.character(dose), levels = c("100", "50", "10", "0")))

sig_cell_range <- range(significant_windows_only$cell_count, na.rm = TRUE)
min_sig_cells <- sig_cell_range[1] 
max_sig_cells <- sig_cell_range[2]

actual_max_abs_lolli <- max(abs(range(static_data$mean_normalized_effect, na.rm = TRUE)))
actual_max_abs_lolli <- max(actual_max_abs_lolli, 1.5)
cap_val_lolli <- 1.0
thresh_val_lolli <- mean_norm_effect_threshold
total_span_lolli <- 2 * actual_max_abs_lolli

norm_pos_lolli <- function(val) {
  (val - (-actual_max_abs_lolli)) / total_span_lolli
}

v_neg_cap_lolli <- norm_pos_lolli(-cap_val_lolli)
v_neg_thresh_lolli <- norm_pos_lolli(-thresh_val_lolli)
v_pos_thresh_lolli <- norm_pos_lolli(thresh_val_lolli)
v_pos_cap_lolli <- norm_pos_lolli(cap_val_lolli)

p_lollipop_static <- ggplot(static_data_lolli, aes(x = window)) +
  geom_segment(
    data = subset(static_data_lolli, !is.na(mean_normalized_effect)),
    aes(xend = window, y = 0, yend = mean_normalized_effect),
    color = "gray"
  ) +
  geom_point(
    data = subset(static_data_lolli, !is.na(mean_normalized_effect)),
    aes(y = mean_normalized_effect, color = mean_normalized_effect, size = cell_count),
    alpha = 0.9
  ) +
  geom_point(
    data = subset(static_data_lolli, is.na(mean_normalized_effect)),
    aes(y = 0, size = cell_count),
    color = "black"
  ) +
  facet_grid(dose ~ ., switch = "y", labeller = as_labeller(function(x) paste0(x, "~mu*'M'"), default = label_parsed)) +
  scale_color_gradientn(
    colors = c("blue", "blue", "lightgrey", "lightgrey", "red", "red"),
    values = c(0, v_neg_cap_lolli, v_neg_thresh_lolli, v_pos_thresh_lolli, v_pos_cap_lolli, 1),
    limits = c(-actual_max_abs_lolli, actual_max_abs_lolli),
    name = expression(beta[coef])
  ) +
  scale_size_continuous(
    limits = c(min_sig_cells, max_sig_cells),
    range  = c(1, 6),
    name   = "Cell\ncount",
    breaks = c(min_sig_cells, 500, 1000, max_sig_cells)
  ) +
  guides(
    color = guide_colorbar(title.theme = element_text(size = 14)),
    size = guide_legend(title.theme = element_text(size = 12))
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.text.y.left = element_text(angle = 90, size = 13),
    strip.placement = "outside",
    legend.text = element_text(size = 10),
    legend.position = "right",
    panel.spacing = unit(0.7, "cm"),
    plot.margin = margin(t = 0, r = 1, b = 1, l = 1, unit = "cm")
  ) +
  labs(y = NULL) +
  coord_cartesian(expand = FALSE, clip = "off")

if (use_patchwork) {
  combined_lollipop_static <- p_header_static / p_lollipop_static +
    plot_layout(heights = c(1, 10), guides = "collect")
} else {
  combined_lollipop_static <- grid.arrange(p_header_static, p_lollipop_static, ncol = 1, heights = c(1, 10))
}

ggsave(
  filename = file.path(paper_plots_dir, "main_fig2_FINAL_100cell_threshold_cell_number_static_lollipop_significant_windows.pdf"),
  plot = combined_lollipop_static,
  width = 12, height = 6
)

ggsave(
  filename = file.path(paper_plots_dir, "main_fig2_FINAL_100cell_threshold_cell_number_static_lollipop_significant_windows.png"),
  plot = combined_lollipop_static,
  width = 12, height = 8, dpi = 600
)
```

-----------------------------------
# Create a dataframe with a single row for each distinct sgRNA, concatenating overlapping windows to merge with the CDS object and plot the percentage of positive cells per window.


```{r}
sgRNA_long <- modeling_results_raw %>%
  tidyr::separate_rows(sgRNA_per_window, sep = ",") %>%
  dplyr::rename(sgRNA.sequence = sgRNA_per_window) %>%
  dplyr::select(ATAC_bin, window, window_scheme, sgRNA.sequence)

sgRNA_for_cds_merge <- sgRNA_long %>%
  group_by(sgRNA.sequence) %>%
  summarize(
    window = paste(unique(window), collapse = ","),
    window_scheme = paste(unique(window_scheme), collapse = ","),
    ATAC_bin = paste(unique(ATAC_bin), collapse = "!!"),
    .groups = "drop"
  )

print(paste("Dimensions before removing empty sgRNA.sequence:", paste(dim(sgRNA_for_cds_merge), collapse = " x ")))

sgRNA_for_cds_merge <- sgRNA_for_cds_merge %>%
  filter(!is.na(sgRNA.sequence) & trimws(sgRNA.sequence) != "")

print(paste("Dimensions after removing empty sgRNA.sequence:", paste(dim(sgRNA_for_cds_merge), collapse = " x ")))

num_multiple_bins <- sum(grepl("!!", sgRNA_for_cds_merge$ATAC_bin))
print(paste("Number of sgRNAs with multiple ATAC_bins assigned:", num_multiple_bins))

problematic_sgrnas <- sgRNA_for_cds_merge %>%
  filter(grepl("!!", ATAC_bin))
print("Problematic sgRNAs with multiple ATAC_bins:")
print(problematic_sgrnas)

head(sgRNA_for_cds_merge)
```


# The CDS object was made in targeted_cCRE_combine_cds_preliminary_analysis.Rmd

```{r}
cds_path <- "./outputs/targeted_cCRE_combined_cds_all_genes_final.RDS"
cds <- readRDS(cds_path)

cd_df <- as.data.frame(colData(cds))

#  Check for mismatches before merging
# Find combinations in our new dataframe that do not exist in the cds colData
mismatches <- sgRNA_for_cds_merge %>%
  anti_join(cd_df, by = c("sgRNA.sequence", "ATAC_bin"))

if (nrow(mismatches) > 0) {
  warning("Warning: Some sgRNA.sequence and ATAC_bin combinations in sgRNA_for_cds_merge do not match any cells in the CDS colData!")
  print(head(mismatches))
} else {
  message("All sgRNA+ATAC_bin combinations match successfully with the CDS colData.")
}

cd_joined <- cd_df %>%
  left_join(sgRNA_for_cds_merge, by = c("sgRNA.sequence", "ATAC_bin")) %>%
  mutate(window = if_else(is.na(window), ATAC_bin, window))

rownames(cd_joined) <- rownames(cd_df)

colData(cds) <- DataFrame(cd_joined)

head(colData(cds)[, c("sgRNA.sequence", "ATAC_bin", "window", "window_scheme")])
```

```{r}
cds_for_sgRNA_enrichment <- cds
column_df_cds_for_sgRNA_enrichment <- colData(cds_for_sgRNA_enrichment) %>% as.data.frame()
```

```{r}
# --- sgRNA Enrichment Analysis for Significant Windows ---
# This analysis looks at how the distribution of individual sgRNAs within significant ATAC bins changes across doses.
# We focus on the proportion of cells with each guide relative to the total cells in that bin and dose.

#  Filter the colData to keep only cells belonging to significant windows
# Note: In 'column_df_cds_for_sgRNA_enrichment$window', multiple windows may be concatenated (comma-separated)
significant_windows_list <- unique(significant_windows_only$window)

if (!require(stringr)) library(stringr)
sig_pattern <- paste(significant_windows_list, collapse = "|")

df_sig_cells <- column_df_cds_for_sgRNA_enrichment %>%
  filter(str_detect(window, sig_pattern))

generate_sgrna_palette <- function(n) {
  sgrna_colors_long <- c(
    "#3C3CD3", "#FE8F42", "#5D3FD3", "#FF00FF", "#555555", "#FFD300", "#009FFF", "#FFFF00", "#00FFBE", "#AE02DD",
    "#9A4D42", "#B1CC71", "#FFACFD", "#1F9698", "#F1085C", "#766C95", "#FF0000", "#00FFFF", "#00479E", "#A3FF00",
    "#009ACD", "#FF4500", "#32CD32", "#FF69B4", "#98FB98", "#A10300", "#14F9FF", "#858567", "#DC5E93", "#93D4FF",
    "#004CFF", "#C47A04", "#8701AC", "#00BFFF", "#FFD700", "#00C2A0", "#FFB79F", "#A0E9FF", "#006326", "#FF1493",
    "#00FF1556", "#a05fdd", "#19c4ea", "#D57CD1", "#5D5948", "#92EA58", "#6C2006", "#BB8169", "#85d1a2", "#860c3a"
  )

  if (n > length(sgrna_colors_long)) {
    warning("More guides than distinct colors. Recycling palette.")
    return(rep(sgrna_colors_long, length.out = n))
  }
  return(sgrna_colors_long[1:n])
}

plot_sgrna_enrichment_proportions <- function(df, target_atac_bin, cohort_label = "cells100") {
  bin_df <- df %>% filter(ATAC_bin == target_atac_bin)

  if (nrow(bin_df) == 0) {
    message(paste("Note: No cells found for bin:", target_atac_bin))
    return(NULL)
  }

  group_cols <- c("dose", "sgRNA.sequence")
  if ("sgRNA" %in% names(bin_df)) group_cols <- c(group_cols, "sgRNA")
  if ("genomic_start" %in% names(bin_df)) group_cols <- c(group_cols, "genomic_start")

  enrichment_stats <- bin_df %>%
    group_by(across(all_of(group_cols))) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(dose) %>%
    mutate(
      total_cells_in_dose_bin = sum(count),
      proportion = count / total_cells_in_dose_bin
    ) %>%
    ungroup()

  enrichment_stats$dose <- factor(enrichment_stats$dose, levels = c(0, 10, 50, 100))

  # Order Guides by Genomic Position (Numeric)
  if ("genomic_start" %in% names(enrichment_stats)) {
    guide_order_df <- enrichment_stats %>%
      dplyr::select(sgRNA.sequence, genomic_start) %>%
      distinct() %>%
      arrange(as.numeric(genomic_start))

    enrichment_stats$sgRNA.sequence <- factor(
      enrichment_stats$sgRNA.sequence,
      levels = guide_order_df$sgRNA.sequence
    )
  }

  if ("sgRNA" %in% names(enrichment_stats)) {
    enrichment_stats$hover_text <- paste0(
      "sgRNA Name: ", enrichment_stats$sgRNA, "<br>",
      "Sequence: ", enrichment_stats$sgRNA.sequence, "<br>",
      "Genomic Start: ", enrichment_stats$genomic_start, "<br>",
      "Proportion: ", round(enrichment_stats$proportion * 100, 2), "% (", enrichment_stats$count, "/", enrichment_stats$total_cells_in_dose_bin, ")<br>",
      "Dose: ", enrichment_stats$dose, " ng/mL"
    )
  } else {
    enrichment_stats$hover_text <- paste0(
      "Sequence: ", enrichment_stats$sgRNA.sequence, "<br>",
      "Genomic Start: ", enrichment_stats$genomic_start, "<br>",
      "Proportion: ", round(enrichment_stats$proportion * 100, 2), "% (", enrichment_stats$count, "/", enrichment_stats$total_cells_in_dose_bin, ")<br>",
      "Dose: ", enrichment_stats$dose, " ng/mL"
    )
  }

  num_guides <- length(unique(enrichment_stats$sgRNA.sequence))
  custom_colors <- generate_sgrna_palette(num_guides)

  p <- ggplot(enrichment_stats, aes(x = dose, y = proportion, fill = sgRNA.sequence, text = hover_text)) +
    geom_bar(stat = "identity", position = "stack", color = "white", linewidth = 0.2, alpha = 0.9) +
    scale_fill_manual(values = custom_colors) +
    theme_minimal() +
    theme(
      legend.position = "right",
      legend.text = element_text(size = 6),
      legend.title = element_text(size = 8, face = "bold"),
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 10, color = "gray30")
    ) +
    labs(
      title = paste("sgRNA Enrichment Distribution:", target_atac_bin),
      subtitle = "Relative proportions of unique sgRNAs per dosage group (stacked to 100%)",
      x = "Dose (ng/mL)",
      y = "Proportion of Captured Cells",
      fill = "sgRNA Sequence"
    ) +
    guides(fill = guide_legend(ncol = ceiling(num_guides / 20)))

  sub_dir <- file.path(output_dir, "sgRNA_proportion")
  if (!dir.exists(sub_dir)) dir.create(sub_dir, recursive = TRUE, showWarnings = FALSE)

  safe_bin_name <- gsub(":", "_", target_atac_bin)
  plot_filename <- file.path(sub_dir, paste0("sci_sgRNA_proportions_over_dose_", cohort_label, "_", safe_bin_name, ".png"))
  ggsave(plot_filename, plot = p, width = 12, height = 7, dpi = 300)

  if (require(plotly) && require(htmlwidgets)) {
    p_interactive <- ggplotly(p, tooltip = "text") %>%
      layout(
        title = list(text = paste0("Interactive sgRNA Enrichment: ", target_atac_bin), x = 0.05),
        margin = list(t = 100)
      )

    html_filename <- file.path(sub_dir, paste0("sci_sgRNA_proportions_over_dose_", cohort_label, "_", safe_bin_name, "_INTERACTIVE.html"))
    saveWidget(p_interactive, file = html_filename, selfcontained = TRUE)

    print(p_interactive)
  }

  return(p)
}

p_enrichment_1 <- plot_sgrna_enrichment_proportions(df_sig_cells, "chr2:47403155-47403402")
if (!is.null(p_enrichment_1)) print(p_enrichment_1)

p_enrichment_2 <- plot_sgrna_enrichment_proportions(df_sig_cells, "chr2:48530040-48530316")
if (!is.null(p_enrichment_2)) print(p_enrichment_2)

p_enrichment_3 <- plot_sgrna_enrichment_proportions(df_sig_cells, "chr2:47402900-47403140")
if (!is.null(p_enrichment_3)) print(p_enrichment_3)
```



# START plot percent cells positive plotting
```{r}
# --- plot_percent_cells_positive for specific windows ---
# Define the specific windows of interest
win1 <- "chr2:48530040-48530190"
win2 <- "chr2:48530065-48530215"

# Extract colData to manipulate and classify cells
cd_df <- as.data.frame(colData(cds))

# Filter for dose == 100 and gracefully map duplicating index overlaps for adjacent blocks
cd_df_100_base <- as.data.frame(colData(cds)) %>% filter(dose == 100)

cds_cols_to_extract <- character()
group_assignments <- character()

# Extract NTC_100
ntc_cells_100 <- rownames(cd_df_100_base %>% filter(ATAC_bin == "NTC"))
cds_cols_to_extract <- c(cds_cols_to_extract, ntc_cells_100)
group_assignments <- c(group_assignments, rep("NTC_100", length(ntc_cells_100)))

# Extract win1 to securely prevent starvation
w1_cells_100 <- rownames(cd_df_100_base %>% filter(grepl(win1, window, fixed = TRUE)))
if (length(w1_cells_100) > 0) {
  cds_cols_to_extract <- c(cds_cols_to_extract, w1_cells_100)
  group_assignments <- c(group_assignments, rep(paste0(win1, "_100"), length(w1_cells_100)))
}

# Extract win2 mapped independently to prevent identical starvation completely
w2_cells_100 <- rownames(cd_df_100_base %>% filter(grepl(win2, window, fixed = TRUE)))
if (length(w2_cells_100) > 0) {
  cds_cols_to_extract <- c(cds_cols_to_extract, w2_cells_100)
  group_assignments <- c(group_assignments, rep(paste0(win2, "_100"), length(w2_cells_100)))
}

# Enforce the correct factor ordering for plotting natively
desired_order <- c("NTC_100", paste0(win1, "_100"), paste0(win2, "_100"))

# Subset the main CDS natively wrapping duplicate clones properly structurally
# --- CELL CLONING STRATEGY FOR SLIDING WINDOW OVERLAPS ---
# Because 150bp sliding windows taking 25bp steps naturally overlap, a single cell often maps into multiple adjoining windows.
# Assigning a cell exclusively to its *first* matched window artificially starves adjacent target windows of real data.
# To maintain perfect data integrity unconditionally, we safely dynamically duplicate any single cell hitting multiple
# windows inside the extraction array natively. By explicitly appending a strictly unique suffix (`_dup_X`) to each duplicated
# cell barcode below, Monocle3 legally treats them as mathematically distinct independent entities, cleanly sorting each
# hit securely into its correct respective 'plot_group' X-axis bar without cross-contaminating or structurally double-counting
# internally within any individual distinct window.
cds_subset <- cds[, cds_cols_to_extract]
colnames(cds_subset) <- paste0(colnames(cds_subset), "_dup_", seq_along(cds_cols_to_extract))

colData(cds_subset)$plot_group <- factor(group_assignments, levels = desired_order)

# Subset specifically for the MSH2 gene explicitly
cds_subset_MSH2 <- cds_subset[rowData(cds_subset)$gene_short_name == "MSH2", ]

plot <- plot_percent_cells_positive(
  cds_subset_MSH2,
  group_cells_by = "plot_group"
)

colors <- c(
  "NTC_100" = "darkgrey",
  setNames(rep("#FF1493", 2), c(paste0(win1, "_100"), paste0(win2, "_100")))
)

plot <- plot +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot)

STON1_plot_out_dir <- "./outputs/window_p_val_plot_percent_pos/STON1_hits/"
dir.create(STON1_plot_out_dir, recursive = TRUE, showWarnings = FALSE)

ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_percent_cells_pos_MSH2_dose100_STON1_hit_windows.png"),
  plot = plot,
  width = 1.2 + 0.2 * length(desired_order), height = 4.5, dpi = 300
)
```


make plot percent cells positive function to do dose matched pairwise comparision with p value
```{r}
# --- Modified function to plot percent cells positive with Dose-Matched Pairwise P-values ---
plot_percent_cells_positive_pvals_multi <- function(
  cds_subset,
  group_cells_by = NULL,
  min_expr = 0,
  nrow = NULL,
  ncol = 1,
  panel_order = NULL,
  label_by_short_name = TRUE,
  normalize = TRUE,
  bootstrap_samples = 1000,
  conf_int_alpha = 0.95
) {
  # Standard Monocle3 setup
  assertthat::assert_that(methods::is(cds_subset, "cell_data_set"))
  marker_exprs <- SingleCellExperiment::counts(cds_subset)

  if (normalize) {
    marker_exprs <- Matrix::t(Matrix::t(marker_exprs) / size_factors(cds_subset))
    marker_exprs_melted <- reshape2::melt(round(as.matrix(marker_exprs), digits = 4))
  } else {
    marker_exprs_melted <- reshape2::melt(as.matrix(marker_exprs))
  }

  colnames(marker_exprs_melted) <- c("f_id", "cell", "expression")
  c_data <- as.data.frame(colData(cds_subset))
  r_data <- as.data.frame(rowData(cds_subset))
  c_data$cell_id_internal <- rownames(c_data)
  r_data$feat_id_internal <- rownames(r_data)

  marker_exprs_melted <- merge(marker_exprs_melted, c_data, by.x = "cell", by.y = "cell_id_internal")
  marker_exprs_melted <- merge(marker_exprs_melted, r_data, by.x = "f_id", by.y = "feat_id_internal")

  if (label_by_short_name && !is.null(marker_exprs_melted$gene_short_name)) {
    marker_exprs_melted$feature_label <- marker_exprs_melted$gene_short_name
    marker_exprs_melted$feature_label[is.na(marker_exprs_melted$feature_label)] <- marker_exprs_melted$f_id
  } else {
    marker_exprs_melted$feature_label <- marker_exprs_melted$f_id
  }

  if (!is.null(panel_order)) {
    marker_exprs_melted$feature_label <- factor(marker_exprs_melted$feature_label, levels = panel_order)
  }

  if (is.null(group_cells_by)) {
    stop("group_cells_by must be provided for a between-group comparison.")
  }

  gene_list <- split(marker_exprs_melted, marker_exprs_melted$feature_label, drop = TRUE)

  marker_counts <- lapply(gene_list, function(sub_data) {
    if (nrow(sub_data) == 0) {
      return(NULL)
    }

    lvls <- unique(as.character(sub_data[[group_cells_by]]))

    group_stats <- do.call(rbind, lapply(lvls, function(l) {
      grp_data <- sub_data[sub_data[[group_cells_by]] == l, ]

      b_grp <- boot::boot(data = grp_data, statistic = function(d, i) {
        mean(d[i, ]$expression > min_expr)
      }, R = bootstrap_samples)

      dose_val <- sub(".*_([^_]+)$", "\\1", l)
      is_ntc <- grepl("^NTC_", l)

      data.frame(
        feature_label = sub_data$feature_label[1],
        group_var = l,
        dose_val = dose_val,
        is_ntc = is_ntc,
        target_fraction_mean = b_grp$t0,
        target_fraction_low = stats::quantile(b_grp$t, (1 - conf_int_alpha) / 2),
        target_fraction_high = stats::quantile(b_grp$t, 1 - (1 - conf_int_alpha) / 2),
        p_value = NA_real_,
        is_imposed_pval = FALSE,
        stringsAsFactors = FALSE
      )
    }))

    # Now that we have all basic stats, perform pairwise p-value calculation
    # For every group that is NOT an NTC, find the matching NTC for its dose
    for (i in 1:nrow(group_stats)) {
      if (!group_stats$is_ntc[i]) {
        target_group <- group_stats$group_var[i]
        curr_dose <- group_stats$dose_val[i]
        ntc_group <- paste0("NTC_", curr_dose)

        if (ntc_group %in% group_stats$group_var) {
          combined_data <- sub_data[sub_data[[group_cells_by]] %in% c(target_group, ntc_group), ]

          b_obj_diff <- boot::boot(data = combined_data, statistic = function(d, boot_idx) {
            d_sub <- d[boot_idx, ]
            m_ntc <- mean(d_sub$expression[d_sub[[group_cells_by]] == ntc_group] > min_expr)
            m_tgt <- mean(d_sub$expression[d_sub[[group_cells_by]] == target_group] > min_expr)
            return(m_tgt - m_ntc)
          }, R = bootstrap_samples, strata = as.factor(combined_data[[group_cells_by]]))

          # Execute ONLY the boot.pval percentile method
          # If this fails, we return NA to maintain methodological purity.
          p_val <- tryCatch(
            {
              boot.pval::boot.pval(b_obj_diff, type = "perc", theta_null = 0)
            },
            error = function(e) {
              return(NA_real_)
            }
          )

          # Audit Flag: If it failed, mark it for your report
          group_stats$is_imposed_pval[i] <- is.na(p_val)
          group_stats$p_value[i] <- p_val
        }
      }
    }
    return(group_stats)
  }) %>% do.call(rbind, .)

  # --- Dynamic Benjamini-Hochberg (BH) Correction ---
  # We explicitly set n to the count of non-NA p-values. 
  # This ensures we only correct for the actual number of Target vs. NTC tests.
  marker_counts <- marker_counts %>%
    dplyr::mutate(
      p_value_adj = p.adjust(
        p_value, 
        method = "BH", 
        n = sum(!is.na(p_value))
      )
    )

  colnames(marker_counts)[which(names(marker_counts) == "group_var")] <- group_cells_by

  # Format P-value labels to display only significant bold asterisks
  marker_counts$p_label <- sapply(marker_counts$p_value_adj, function(p) {
    if (is.na(p)) {
      return("")
    }

    if (p < 0.01) {
      return("**")
    } # Check 0.01 first
    if (p < 0.05) {
      return("*")
    }

    return("") # Return empty string if not significant
  })

  # Final Plot construction scale to percentages
  marker_counts$target_fraction_mean <- marker_counts$target_fraction_mean * 100
  marker_counts$target_fraction_low <- marker_counts$target_fraction_low * 100
  marker_counts$target_fraction_high <- marker_counts$target_fraction_high * 100

  # Ensure the group variable is a factor for proper x-axis ordering
  marker_counts[[group_cells_by]] <- factor(marker_counts[[group_cells_by]], levels = levels(c_data[[group_cells_by]]))

  qp <- ggplot(marker_counts, aes_string(x = group_cells_by, y = "target_fraction_mean", fill = group_cells_by)) +
    facet_wrap(~feature_label, nrow = nrow, ncol = ncol, scales = "free_y") +
    geom_bar(stat = "identity", width = 0.8) +
    geom_linerange(aes(ymin = target_fraction_low, ymax = target_fraction_high)) +
    geom_text(aes(label = p_label, y = target_fraction_high + 2), vjust = 0, size = 4, fontface = "bold") +
    ylab("Cells (percent)") +
    monocle3:::monocle_theme_opts()

  return(list(plot = qp, stats = marker_counts))
}

# -------------------------------------------------------------------------
# Initialize master audit list mapping statistically imposed values dynamically
all_plot_audit_data <- list()

set.seed(42) # Set seed for stable bootstrapping

# Plot P-values for Dose 100 only
res_dose100 <- plot_percent_cells_positive_pvals_multi(
  cds_subset_MSH2,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_pval_dose100 <- res_dose100$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_dose100$stats

plot_pval_dose100 <- plot_pval_dose100 +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot_pval_dose100)

ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_A_plot_percent_cells_pos_MSH2_dose100_STON1_hit_windows_with_pvals.png"),
  plot = plot_pval_dose100,
  width = 1.2 + 0.2 * length(desired_order), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_A_plot_percent_cells_pos_MSH2_dose100_STON1_hit_windows_with_pvals.pdf"),
  plot = plot_pval_dose100,
  width = 1.2 + 0.2 * length(desired_order), height = 4.5, dpi = 300, device = "pdf"
)

plot_pval_dose100_no_xlabel <- plot_pval_dose100 + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_A_plot_percent_cells_pos_MSH2_dose100_STON1_hit_windows_with_pvals_no_xlabel.pdf"),
  plot = plot_pval_dose100_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order), height = 1.0, dpi = 600, device = "pdf"
)
```

```{r}
# ==============================================================================
# STON1: sgRNA Distribution per Window
# ==============================================================================

cat("\n--- sgRNA Distribution across Doses for STON1 Hit Windows ---\n")

# Safely extract strictly native dataframes dynamically without modifying downstream CDS objects
ston1_windows <- c("chr2:48530040-48530190", "chr2:48530065-48530215")
all_cd_df_ston1 <- as.data.frame(colData(cds))

for (base_w in ston1_windows) {
  cat(paste0("\n========================================\n"))
  cat(paste0("Distribution for STON1 Window: ", base_w, "\n"))
  cat(paste0("========================================\n"))

  # Filter strictly for cells targeting this exact base window safely mapped
  cells_in_window <- all_cd_df_ston1 %>%
    filter(grepl(base_w, window, fixed = TRUE))

  # Group mathematically and pivot cleanly to long matrix layout
  sgRNA_dist <- cells_in_window %>%
    group_by(sgRNA.sequence, dose) %>%
    summarise(cell_count = n(), .groups = "drop") %>%
    pivot_wider(
      names_from = dose,
      names_prefix = "dose_",
      values_from = cell_count,
      values_fill = list(cell_count = 0)
    )

  # Force all expected doses to exist cleanly even if technically zero counts
  expected_cols <- paste0("dose_", c(0, 10, 50, 100))
  for (col in expected_cols) {
    if (!col %in% names(sgRNA_dist)) {
      sgRNA_dist[[col]] <- 0
    }
  }

  sgRNA_dist$window <- base_w

  sgRNA_dist <- sgRNA_dist %>%
    dplyr::select(window, sgRNA.sequence, all_of(expected_cols)) %>%
    arrange(sgRNA.sequence)

  totals <- data.frame(
    window = base_w,
    sgRNA.sequence = "TOTAL",
    stringsAsFactors = FALSE
  )
  for (col in expected_cols) {
    totals[[col]] <- sum(sgRNA_dist[[col]], na.rm = TRUE)
  }
  sgRNA_dist <- bind_rows(sgRNA_dist, totals)

  print(as.data.frame(sgRNA_dist))

  safe_name <- gsub("[:-]", "_", base_w) 
  env_var_name <- paste0("sgRNA_dist_", safe_name)
  assign(env_var_name, as.data.frame(sgRNA_dist))
  cat(paste0("   -> Saved dataframe safely to environment as: '", env_var_name, "'\n"))
}
```


```{r}
# --- Combined Windows (win1 & win2) vs NTC across ALL Doses ---

# Define windows of interest
win1 <- "chr2:48530040-48530190"
win2 <- "chr2:48530065-48530215"

cd_df_comb_all <- as.data.frame(colData(cds)) %>%
  mutate(base_group = as.character(NA))

# Classify cells into NTC or STON1_hit_windows base group
cd_df_comb_all$base_group[cd_df_comb_all$ATAC_bin == "NTC"] <- "NTC"

is_target_window <- grepl(win1, cd_df_comb_all$window, fixed = TRUE) | grepl(win2, cd_df_comb_all$window, fixed = TRUE)
cd_df_comb_all$base_group[is_target_window & is.na(cd_df_comb_all$base_group)] <- "STON1_hit_windows"

# Filter out unassigned cells
cd_df_comb_all <- cd_df_comb_all %>% filter(!is.na(base_group))

# Combine with dose to make the final plot_group factor
cd_df_comb_all <- cd_df_comb_all %>%
  mutate(plot_group = paste0(base_group, "_", dose))

# Enforce correct factor ordering (0, 10, 50, 100)
doses <- c("0", "10", "50", "100")
desired_order_comb_all <- c(rbind(
  paste0("NTC_", doses),
  paste0("STON1_hit_windows_", doses)
))
cd_df_comb_all$plot_group <- factor(cd_df_comb_all$plot_group, levels = desired_order_comb_all)

# Subset CDS
cds_subset_comb_all <- cds[, rownames(cd_df_comb_all)]
colData(cds_subset_comb_all) <- DataFrame(cd_df_comb_all)

# Subset for MSH2
cds_subset_MSH2_comb_all <- cds_subset_comb_all[rowData(cds_subset_comb_all)$gene_short_name == "MSH2", ]

# Colors and Styling
colors_comb_all <- c(
  setNames(rep("darkgrey", 4), paste0("NTC_", doses)),
  setNames(rep("#FF1493", 4), paste0("STON1_hit_windows_", doses))
)

# Generate Normal Plot
plot_comb_all_normal <- plot_percent_cells_positive(
  cds_subset_MSH2_comb_all,
  group_cells_by = "plot_group"
)

plot_comb_all_normal <- plot_comb_all_normal +
  scale_fill_manual(values = colors_comb_all) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot_comb_all_normal)
ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_percent_cells_pos_MSH2_all_doses_STON1_hit_windows_combined_windows_normal.png"),
  plot = plot_comb_all_normal,
  width = 1.2 + 0.2 * length(desired_order_comb_all), height = 4.5, dpi = 300
)

# Generate P-Value Plot with Custom Function
set.seed(42)
res_comb_all <- plot_percent_cells_positive_pvals_multi(
  cds_subset_MSH2_comb_all,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_comb_all_pval <- res_comb_all$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_comb_all$stats

plot_comb_all_pval <- plot_comb_all_pval +
  scale_fill_manual(values = colors_comb_all) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot_comb_all_pval)
ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_B_plot_percent_cells_pos_MSH2_all_doses_STON1_hit_windows_combined_windows_with_pvals.png"),
  plot = plot_comb_all_pval,
  width = 1.2 + 0.2 * length(desired_order_comb_all), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_B_plot_percent_cells_pos_MSH2_all_doses_STON1_hit_windows_combined_windows_with_pvals.pdf"),
  plot = plot_comb_all_pval,
  width = 1.2 + 0.2 * length(desired_order_comb_all), height = 4.5, dpi = 300, device = "pdf"
)

plot_comb_all_pval_no_xlabel <- plot_comb_all_pval + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(STON1_plot_out_dir, "plot_B_plot_percent_cells_pos_MSH2_all_doses_STON1_hit_windows_combined_windows_with_pvals_no_xlabel.pdf"),
  plot = plot_comb_all_pval_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order_comb_all), height = 1.0, dpi = 600, device = "pdf"
)
```



-------------------------
Plot % cells positive MSH2 exon1

```{r}
# ==============================================================================
# MSH2 EXON 1 HIT WINDOWS vs NTC (Dose 100)
# ==============================================================================

# Output Directory Configuration
MSH2_exon1_plot_out_dir <- "./outputs/window_p_val_plot_percent_pos/MSH2_exon1_hits/"
dir.create(MSH2_exon1_plot_out_dir, recursive = TRUE, showWarnings = FALSE)

# Extract specific MSH2 Exon 1 hits
msh2_exon1_hits_df <- significant_windows_only %>%
  filter(ATAC_bin == "chr2:47403155-47403402") %>%
  dplyr::select(ATAC_bin, window, window_start) %>%
  distinct() %>%
  arrange(as.numeric(window_start))

print("Found MSH2 Exon 1 hit windows:")
print(msh2_exon1_hits_df)

# Base window names (strip any potential trailing _dose if they exist)
msh2_exon1_windows <- as.character(msh2_exon1_hits_df$window)
msh2_exon1_windows_base <- sub("_[0-9]+$", "", msh2_exon1_windows)

# ==============================================================================
# MSH2 EXON 1 ALL HIT WINDOWS POOLED ALL DOSES vs NTC ALL DOSES
# ==============================================================================

# Extract all unique NTC cells across all doses
cd_df_pooled_base <- as.data.frame(colData(cds))
ntc_cells_all_pooled <- rownames(cd_df_pooled_base %>% filter(ATAC_bin == "NTC"))

pattern <- paste(msh2_exon1_windows_base, collapse = "|")
target_cells_all_pooled <- rownames(cd_df_pooled_base %>% filter(grepl(pattern, window)))

cds_cols_pooled <- c(ntc_cells_all_pooled, target_cells_all_pooled)
group_assignments_pooled <- c(
  rep("NTC_all", length(ntc_cells_all_pooled)),
  rep("Combined_MSH2_exon1_all", length(target_cells_all_pooled))
)

desired_order_pooled <- c("NTC_all", "Combined_MSH2_exon1_all")

cds_subset_pooled_msh2 <- cds[, cds_cols_pooled]
colData(cds_subset_pooled_msh2)$plot_group <- factor(group_assignments_pooled, levels = desired_order_pooled)
cds_subset_pooled_msh2 <- cds_subset_pooled_msh2[rowData(cds_subset_pooled_msh2)$gene_short_name == "MSH2", ]

colors_pooled <- setNames(c("darkgrey", "deepskyblue3"), desired_order_pooled)

plot_pooled_normal <- plot_percent_cells_positive(
  cds_subset_pooled_msh2,
  group_cells_by = "plot_group"
) +
  scale_fill_manual(values = colors_pooled) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_percent_cells_pos_MSH2_all_doses_pooled_vs_NTC_all_doses_normal.png"),
  plot = plot_pooled_normal,
  width = 1.2 + 0.2 * length(desired_order_pooled), height = 4.5, dpi = 300
)

# Generate P-Value Plot via unified function
set.seed(42)
res_pooled_pval <- plot_percent_cells_positive_pvals_multi(
  cds_subset_pooled_msh2,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_pooled_pval <- res_pooled_pval$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_pooled_pval$stats

plot_pooled_pval <- plot_pooled_pval +
  scale_fill_manual(values = colors_pooled) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_C_plot_percent_cells_pos_MSH2_all_doses_pooled_vs_NTC_all_doses_with_pvals.png"),
  plot = plot_pooled_pval,
  width = 1.2 + 0.2 * length(desired_order_pooled), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_C_plot_percent_cells_pos_MSH2_all_doses_pooled_vs_NTC_all_doses_with_pvals.pdf"),
  plot = plot_pooled_pval,
  width = 1.2 + 0.2 * length(desired_order_pooled), height = 4.5, dpi = 300, device = "pdf"
)

plot_pooled_pval_no_xlabel <- plot_pooled_pval + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_C_plot_percent_cells_pos_MSH2_all_doses_pooled_vs_NTC_all_doses_with_pvals_no_xlabel.pdf"),
  plot = plot_pooled_pval_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order_pooled), height = 1.0, dpi = 600, device = "pdf"
)

# -----------------------------------------------------------------------------

# Prepare data for Dose = 100
cd_df_100_base <- as.data.frame(colData(cds)) %>% filter(dose == 100)

cds_cols_to_extract_100 <- character()
group_assignments_100 <- character()

# Extract NTC_100
ntc_cells_100 <- rownames(cd_df_100_base %>% filter(ATAC_bin == "NTC"))
cds_cols_to_extract_100 <- c(cds_cols_to_extract_100, ntc_cells_100)
group_assignments_100 <- c(group_assignments_100, rep("NTC_100", length(ntc_cells_100)))

# Extract each individual window individually to clone cells mapped natively to multiple windows
for (base_w in msh2_exon1_windows_base) {
  w_cells_100 <- rownames(cd_df_100_base %>% filter(grepl(base_w, window, fixed = TRUE)))
  if (length(w_cells_100) > 0) {
    cds_cols_to_extract_100 <- c(cds_cols_to_extract_100, w_cells_100)
    group_assignments_100 <- c(group_assignments_100, rep(paste0(base_w, "_100"), length(w_cells_100)))
  }
}

# Enforce numeric ordering securely
desired_order_msh2_100 <- c("NTC_100", paste0(msh2_exon1_windows_base, "_100"))

# Subset CDS securely wrapping duplicating indexed elements
# --- CELL CLONING STRATEGY FOR SLIDING WINDOW OVERLAPS ---
# Because 150bp sliding windows taking 25bp steps naturally overlap, a single cell often maps into multiple adjoining windows.
# Assigning a cell exclusively to its *first* matched window artificially starves adjacent target windows of real data.
# To maintain perfect data integrity unconditionally, we safely dynamically duplicate any single cell hitting multiple
# windows inside the extraction array natively. By explicitly appending a strictly unique suffix (`_dup_X`) to each duplicated
# cell barcode below, Monocle3 legally treats them as mathematically distinct independent entities, cleanly sorting each
# hit securely into its correct respective 'plot_group' X-axis bar without cross-contaminating or structurally double-counting
# internally within any individual distinct window.
cds_subset_100_msh2 <- cds[, cds_cols_to_extract_100]
colnames(cds_subset_100_msh2) <- paste0(colnames(cds_subset_100_msh2), "_dup_", seq_along(cds_cols_to_extract_100))

colData(cds_subset_100_msh2)$plot_group <- factor(group_assignments_100, levels = desired_order_msh2_100)
cds_subset_100_msh2 <- cds_subset_100_msh2[rowData(cds_subset_100_msh2)$gene_short_name == "MSH2", ]

# Colors: darkgrey for NTC, deepskyblue3 for all target windows
colors_msh2_100 <- setNames(
  c("darkgrey", rep("deepskyblue3", length(msh2_exon1_windows_base))),
  desired_order_msh2_100
)

# Normal Plot
plot_msh2_100_normal <- plot_percent_cells_positive(
  cds_subset_100_msh2,
  group_cells_by = "plot_group"
) +
  scale_fill_manual(values = colors_msh2_100) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot_msh2_100_normal)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_percent_cells_pos_MSH2_dose100_MSH2_exon1_hit_windows_normal.png"),
  plot = plot_msh2_100_normal,
  width = 1.2 + 0.2 * length(desired_order_msh2_100), height = 4.5, dpi = 300
)

# P-value Plot
set.seed(42)
res_msh2_100 <- plot_percent_cells_positive_pvals_multi(
  cds_subset_100_msh2,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_msh2_100_pval <- res_msh2_100$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_msh2_100$stats

plot_msh2_100_pval <- plot_msh2_100_pval +
  scale_fill_manual(values = colors_msh2_100) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print(plot_msh2_100_pval)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_D_plot_percent_cells_pos_MSH2_dose100_MSH2_exon1_hit_windows_with_pvals.png"),
  plot = plot_msh2_100_pval,
  width = 1.2 + 0.2 * length(desired_order_msh2_100), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_D_plot_percent_cells_pos_MSH2_dose100_MSH2_exon1_hit_windows_with_pvals.pdf"),
  plot = plot_msh2_100_pval,
  width = 1.2 + 0.2 * length(desired_order_msh2_100), height = 4.5, dpi = 300, device = "pdf"
)

plot_msh2_100_pval_no_xlabel <- plot_msh2_100_pval + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_D_plot_percent_cells_pos_MSH2_dose100_MSH2_exon1_hit_windows_with_pvals_no_xlabel.pdf"),
  plot = plot_msh2_100_pval_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order_msh2_100), height = 1.0, dpi = 600, device = "pdf"
)
```

```{r}
# ==============================================================================
# MSH2 EXON 1 SET 2: Individual Windows (All Doses) vs NTC (All Doses)
# ==============================================================================

# Prepare data dynamically duplicating explicitly overlapping indexed arrays
cd_df_all_base <- as.data.frame(colData(cds))

cds_cols_to_extract_all <- character()
group_assignments_all <- character()

# Extract NTC_all mapping
ntc_cells_all <- rownames(cd_df_all_base %>% filter(ATAC_bin == "NTC"))
cds_cols_to_extract_all <- c(cds_cols_to_extract_all, ntc_cells_all)
group_assignments_all <- c(group_assignments_all, rep("NTC_all", length(ntc_cells_all)))

# Extract each individual window individually to cleanly duplicate mapped cell overlapping nodes natively
for (base_w in msh2_exon1_windows_base) {
  w_cells_all <- rownames(cd_df_all_base %>% filter(grepl(base_w, window, fixed = TRUE)))
  if (length(w_cells_all) > 0) {
    cds_cols_to_extract_all <- c(cds_cols_to_extract_all, w_cells_all)
    group_assignments_all <- c(group_assignments_all, rep(paste0(base_w, "_all"), length(w_cells_all)))
  }
}

desired_order_msh2_all_indiv <- c("NTC_all", paste0(msh2_exon1_windows_base, "_all"))

# Subset CDS supporting overlapping mapped logic cleanly
# --- CELL CLONING STRATEGY FOR SLIDING WINDOW OVERLAPS ---
# Because 150bp sliding windows taking 25bp steps naturally overlap, a single cell often maps into multiple adjoining windows.
# Assigning a cell exclusively to its *first* matched window artificially starves adjacent target windows of real data.
# To maintain perfect data integrity unconditionally, we safely dynamically duplicate any single cell hitting multiple
# windows inside the extraction array natively. By explicitly appending a strictly unique suffix (`_dup_X`) to each duplicated
# cell barcode below, Monocle3 legally treats them as mathematically distinct independent entities, cleanly sorting each
# hit securely into its correct respective 'plot_group' X-axis bar without cross-contaminating or structurally double-counting
# internally within any individual distinct window.
cds_subset_all_msh2_indiv <- cds[, cds_cols_to_extract_all]
colnames(cds_subset_all_msh2_indiv) <- paste0(colnames(cds_subset_all_msh2_indiv), "_dup_", seq_along(cds_cols_to_extract_all))

colData(cds_subset_all_msh2_indiv)$plot_group <- factor(group_assignments_all, levels = desired_order_msh2_all_indiv)
cds_subset_all_msh2_indiv <- cds_subset_all_msh2_indiv[rowData(cds_subset_all_msh2_indiv)$gene_short_name == "MSH2", ]

# Colors
colors_msh2_all_indiv <- setNames(
  c("darkgrey", rep("deepskyblue3", length(msh2_exon1_windows_base))),
  desired_order_msh2_all_indiv
)

# Normal Plot Mapping natively
plot_msh2_all_indiv_normal <- plot_percent_cells_positive(
  cds_subset_all_msh2_indiv,
  group_cells_by = "plot_group"
) +
  scale_fill_manual(values = colors_msh2_all_indiv) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print("Generating Set 2 Normal Plot...")
print(plot_msh2_all_indiv_normal)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_percent_cells_pos_MSH2_all_doses_pooled_MSH2_exon1_hit_windows_normal.png"),
  plot = plot_msh2_all_indiv_normal,
  width = 1.2 + 0.2 * length(desired_order_msh2_all_indiv), height = 4.5, dpi = 300
)

# P-value Plot generating directly
set.seed(42)
res_msh2_all_indiv <- plot_percent_cells_positive_pvals_multi(
  cds_subset_all_msh2_indiv,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_msh2_all_indiv_pval <- res_msh2_all_indiv$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_msh2_all_indiv$stats

plot_msh2_all_indiv_pval <- plot_msh2_all_indiv_pval +
  scale_fill_manual(values = colors_msh2_all_indiv) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), breaks = seq(0, 12, by = 4))

print("Generating Set 2 P-value Plot...")
print(plot_msh2_all_indiv_pval)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_E_plot_percent_cells_pos_MSH2_all_doses_pooled_MSH2_exon1_hit_windows_with_pvals.png"),
  plot = plot_msh2_all_indiv_pval,
  width = 1.2 + 0.2 * length(desired_order_msh2_all_indiv), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_E_plot_percent_cells_pos_MSH2_all_doses_pooled_MSH2_exon1_hit_windows_with_pvals.pdf"),
  plot = plot_msh2_all_indiv_pval,
  width = 1.2 + 0.2 * length(desired_order_msh2_all_indiv), height = 4.5, dpi = 300, device = "pdf"
)

plot_msh2_all_indiv_pval_no_xlabel <- plot_msh2_all_indiv_pval + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_E_plot_percent_cells_pos_MSH2_all_doses_pooled_MSH2_exon1_hit_windows_with_pvals_no_xlabel.pdf"),
  plot = plot_msh2_all_indiv_pval_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order_msh2_all_indiv), height = 1.0, dpi = 600, device = "pdf"
)


# ==============================================================================
# MSH2 EXON 1 SET 3: COMBINED 5 Windows vs NTC (Separated by Dose)
# ==============================================================================

# Extract colData and explicitly initialize dynamically
cd_df_comb_msh2_doses <- as.data.frame(colData(cds)) %>%
  mutate(base_group = as.character(NA))

# Classify cells natively to NTC string directly
cd_df_comb_msh2_doses$base_group[cd_df_comb_msh2_doses$ATAC_bin == "NTC"] <- "NTC"

# Find any cell containing explicitly ANY of the 5 windows explicitly mapped
is_target_window_msh2 <- rep(FALSE, nrow(cd_df_comb_msh2_doses))
for (base_w in msh2_exon1_windows_base) {
  is_target_window_msh2 <- is_target_window_msh2 | grepl(base_w, cd_df_comb_msh2_doses$window, fixed = TRUE)
}
cd_df_comb_msh2_doses$base_group[is_target_window_msh2 & is.na(cd_df_comb_msh2_doses$base_group)] <- "Combined_MSH2_exon1"

# Filter out explicitly unassigned mapped cells natively
cd_df_comb_msh2_doses <- cd_df_comb_msh2_doses %>% filter(!is.na(base_group))

# Combine safely over dose to make final plot formatting cleanly grouped directly
cd_df_comb_msh2_doses <- cd_df_comb_msh2_doses %>%
  mutate(plot_group = paste0(base_group, "_", dose))

#  Enforce exactly structured numeric dose arrays explicitly (0, 10, 50, 100)
doses <- c("0", "10", "50", "100")
desired_order_comb_msh2_doses <- c(rbind(
  paste0("NTC_", doses),
  paste0("Combined_MSH2_exon1_", doses)
))
cd_df_comb_msh2_doses$plot_group <- factor(cd_df_comb_msh2_doses$plot_group, levels = desired_order_comb_msh2_doses)

# Subset CDS objects properly
cds_subset_comb_msh2_doses <- cds[, rownames(cd_df_comb_msh2_doses)]
colData(cds_subset_comb_msh2_doses) <- DataFrame(cd_df_comb_msh2_doses)

# Subset cleanly for specifically MSH2 mapped native arrays
cds_subset_MSH2_comb_msh2_doses <- cds_subset_comb_msh2_doses[rowData(cds_subset_comb_msh2_doses)$gene_short_name == "MSH2", ]

# Native custom Colors
colors_comb_msh2_doses <- c(
  setNames(rep("darkgrey", 4), paste0("NTC_", doses)),
  setNames(rep("deepskyblue3", 4), paste0("Combined_MSH2_exon1_", doses))
)

# Standard Percent Native Cells Output Plot directly assigned natively
plot_comb_msh2_doses_normal <- plot_percent_cells_positive(
  cds_subset_MSH2_comb_msh2_doses,
  group_cells_by = "plot_group"
) +
  scale_fill_manual(values = colors_comb_msh2_doses) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), labels = function(x) sprintf("%.0f", x))

print("Generating Set 3 Normal Plot...")
print(plot_comb_msh2_doses_normal)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_percent_cells_pos_MSH2_all_doses_separated_MSH2_exon1_combined_windows_normal.png"),
  plot = plot_comb_msh2_doses_normal,
  width = 1.2 + 0.2 * length(desired_order_comb_msh2_doses), height = 4.5, dpi = 300
)

# P-Value Empirical Calculation generated over mapped explicit strings
set.seed(42)
res_comb_msh2_doses <- plot_percent_cells_positive_pvals_multi(
  cds_subset_MSH2_comb_msh2_doses,
  group_cells_by = "plot_group",
  bootstrap_samples = 1000
)
plot_comb_msh2_doses_pval <- res_comb_msh2_doses$plot
all_plot_audit_data[[length(all_plot_audit_data) + 1]] <- res_comb_msh2_doses$stats

plot_comb_msh2_doses_pval <- plot_comb_msh2_doses_pval +
  scale_fill_manual(values = colors_comb_msh2_doses) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", strip.text = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8)) +
  scale_y_continuous(limits = c(0, 12.5), labels = function(x) sprintf("%.0f", x))

print("Generating Set 3 P-Value Plot...")
print(plot_comb_msh2_doses_pval)
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_F_plot_percent_cells_pos_MSH2_all_doses_separated_MSH2_exon1_combined_windows_with_pvals.png"),
  plot = plot_comb_msh2_doses_pval,
  width = 1.2 + 0.2 * length(desired_order_comb_msh2_doses), height = 4.5, dpi = 300
)

ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_F_plot_percent_cells_pos_MSH2_all_doses_separated_MSH2_exon1_combined_windows_with_pvals.pdf"),
  plot = plot_comb_msh2_doses_pval,
  width = 1.2 + 0.2 * length(desired_order_comb_msh2_doses), height = 4.5, dpi = 300, device = "pdf"
)

plot_comb_msh2_doses_pval_no_xlabel <- plot_comb_msh2_doses_pval + theme(axis.text.x = element_blank(), axis.title.x = element_blank())
ggsave(
  filename = file.path(MSH2_exon1_plot_out_dir, "plot_F_plot_percent_cells_pos_MSH2_all_doses_separated_MSH2_exon1_combined_windows_with_pvals_no_xlabel.pdf"),
  plot = plot_comb_msh2_doses_pval_no_xlabel,
  width = 0.55 + 0.2 * length(desired_order_comb_msh2_doses), height = 1.0, dpi = 600, device = "pdf"
)
```

```{r}
# ==============================================================================
# MSH2 EXON 1: sgRNA Distribution per Window
# ==============================================================================
library(tidyr)
library(dplyr)

cat("\n--- sgRNA Distribution across Doses for MSH2 Exon 1 Hit Windows ---\n")

# Extract the complete working dataframe cleanly
all_cd_df_msh2 <- as.data.frame(colData(cds))

for (base_w in msh2_exon1_windows_base) {
  cat(paste0("\n========================================\n"))
  cat(paste0("Distribution for Window: ", base_w, "\n"))
  cat(paste0("========================================\n"))

  # Filter strictly for cells targeting this exact base window safely mapped
  cells_in_window <- all_cd_df_msh2 %>%
    filter(grepl(base_w, window, fixed = TRUE))

  # Group mathematically and pivot cleanly to long matrix layout
  sgRNA_dist <- cells_in_window %>%
    group_by(sgRNA.sequence, dose) %>%
    summarise(cell_count = n(), .groups = "drop") %>%
    pivot_wider(
      names_from = dose,
      names_prefix = "dose_",
      values_from = cell_count,
      values_fill = list(cell_count = 0)
    )

  # Force all expected doses to exist cleanly even if technically zero counts
  expected_cols <- paste0("dose_", c(0, 10, 50, 100))
  for (col in expected_cols) {
    if (!col %in% names(sgRNA_dist)) {
      sgRNA_dist[[col]] <- 0
    }
  }

  # Prepend the window name specifically identically as requested explicitly mapped
  sgRNA_dist$window <- base_w

  # Format correctly sequenced ordering mathematically mapped precisely
  sgRNA_dist <- sgRNA_dist %>%
    dplyr::select(window, sgRNA.sequence, all_of(expected_cols)) %>%
    arrange(sgRNA.sequence)

  # ADD SUMMATION TOTAL ROW
  totals <- data.frame(
    window = base_w,
    sgRNA.sequence = "TOTAL",
    stringsAsFactors = FALSE
  )
  for (col in expected_cols) {
    totals[[col]] <- sum(sgRNA_dist[[col]], na.rm = TRUE)
  }
  sgRNA_dist <- bind_rows(sgRNA_dist, totals)

  print(as.data.frame(sgRNA_dist))

  safe_name <- gsub("[:-]", "_", base_w) 
  env_var_name <- paste0("sgRNA_dist_", safe_name)
  assign(env_var_name, as.data.frame(sgRNA_dist))
  cat(paste0("   -> Saved dataframe safely to environment as: '", env_var_name, "'\n"))
}
```

```{r}
# ==============================================================================
# FINAL BOOTSTRAP STATISTICAL AUDIT
# ==============================================================================
cat("\n\n********************************************************\n")
cat("   STATISTICAL AUDIT: IMPOSED P-VALUE LOG\n")
cat("********************************************************\n")

# Combine all results from all plot calls
if (length(all_plot_audit_data) > 0) {
  master_stats <- do.call(rbind, all_plot_audit_data)

  # Filter for windows where p-value was forced due to zero variance
  imposed_hits <- master_stats %>%
    filter(is_imposed_pval == TRUE) %>%
    dplyr::select(feature_label, plot_group, dose_val, p_value) %>%
    distinct()

  if (nrow(imposed_hits) > 0) {
    cat("The following windows triggered a Zero-Variance fail-safe.\n")
    cat("This means the bootstrap distribution did not fluctuate, likely due to\n")
    cat("extremely low expression (all 0s) or uniform expression in the window.\n\n")
    print(as.data.frame(imposed_hits))
  } else {
    cat("AUDIT PASSED: No windows used imposed P-values.\n")
    cat("All significant results were derived from a fluctuating bootstrap distribution.\n")
  }
} else {
  cat("No audit data found. Ensure you are saving 'res$stats' from your plot calls.\n")
}
cat("********************************************************\n")
```
