---
title: "combine_cds_preliminary_plots_prep_cds_for_sliding_window"
output: html_document
date: "2025-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(Biostrings)
library(DESeq2)
library(Matrix)
library(SummarizedExperiment)
library(dplyr)
library(ggplot2)
library(ggrastr)
library(gridExtra)
library(htmlwidgets)
library(monocle3)
library(plotly)
library(readr)
library(stringr)
library(tidyr)
```




```{r}
working_dir <- "./outputs/"
```

```{r}
path_for_plots <- "./outputs/plots/"
```


```{r}
targeted_cCRE_1_cds <- readRDS("./outputs/targeted_cCRE_1_final.RDS")
targeted_cCRE_2_cds <- readRDS("./outputs/targeted_cCRE_2_final.RDS")
targeted_cCRE_3_cds <- readRDS("./outputs/targeted_cCRE_3_final.RDS")
```

```{r}
# Combine the three CDS objects into one
CDS_w_HASH_CROP <- combine_cds(
  list(targeted_cCRE_1_cds, targeted_cCRE_2_cds, targeted_cCRE_3_cds),
  sample_col_name = "cds_identifier"
)
```

```{r}
coldata_CDS_w_HASH_CROP <- colData(CDS_w_HASH_CROP) %>% as.data.frame()

sgRNA_l <- read.csv("../data/common_references/sgRNA_library_FINAL_w_sgOPTI_CROP_adapters.csv", stringsAsFactors = FALSE)

CROP_samplesheet <- read.table("../data/common_references/final_CROP_samplesheet_unique_numbers.txt",
  header = FALSE, sep = "\t", stringsAsFactors = FALSE
)
colnames(CROP_samplesheet) <- c("sgRNA", "sgRNA.sequence", "axis")
CROP_samplesheet <- CROP_samplesheet %>% dplyr::select(sgRNA, sgRNA.sequence)

CROP_samplesheet <- CROP_samplesheet %>%
  mutate(
    sgRNA_w_NTC = case_when(


      grepl("non_targeting_control", sgRNA) ~ "NTC",
      grepl("AAVS1", sgRNA) ~ "AAVS1.control",
      grepl("random_targeting_controls", sgRNA) ~ "random.controls",



      TRUE ~ sgRNA
    )
  )
```
# Resolve overlapping ATAC peaks for sgRNAs assigned to multiple bins by selecting the largest peak length.
# Parse the Gene.Symbol column to determine ATAC seq peak coordinates and calculate peak length.

```{r}
MSH2_CRE_sgRNA_library_parsed <- sgRNA_l %>%
  mutate(Gene.Symbol_to_parse = Gene.Symbol) %>%
  separate(Gene.Symbol_to_parse, into = c("chr", "start_end"), sep = ":") %>%
  separate(start_end, into = c("Gene.Symbol_start", "Gene.Symbol_end"), sep = "-") %>%
  mutate(
    Gene.Symbol_start = as.numeric(Gene.Symbol_start),
    Gene.Symbol_end = as.numeric(Gene.Symbol_end)
  )

MSH2_CRE_sgRNA_library_parsed <- MSH2_CRE_sgRNA_library_parsed %>%
  mutate(ATAC_peak_length = Gene.Symbol_end - Gene.Symbol_start)
```


# Collapse by unique sgRNA.sequence and retain the larger of the two ATAC peaks in the Gene.Symbol column.
# This ensures only the Gene.Symbol with the maximum ATAC_peak_length is assigned to each sgRNA.

```{r}
# Part 1: Collapse the dataframe by unique sgRNA, selecting for the max peak length
collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet <- MSH2_CRE_sgRNA_library_parsed %>%
  group_by(sgRNA.sequence) %>%
  summarise(
    df_name = paste(df_name, collapse = " / "),
    Gene.Symbol = {
      if (all(is.na(ATAC_peak_length))) {
        paste(Gene.Symbol, collapse = " *!! ")
      } else {
        tied <- Gene.Symbol[ATAC_peak_length == max(ATAC_peak_length, na.rm = TRUE)]
        tied <- sort(tied)
        if (length(tied) > 1) {
          paste("ATAC_peak_length tie:", paste(tied, collapse = " * "))
        } else {
          tied
        }
      }
    },
    Gene.Symbol_note = {
      if (all(is.na(ATAC_peak_length))) {
        "All ATAC_peak_length values are NA"
      } else {
        tied <- Gene.Symbol[ATAC_peak_length == max(ATAC_peak_length, na.rm = TRUE)]
        if (length(tied) > 1) {
          "Tie in ATAC_peak_length"
        } else {
          num_valid_peaks <- sum(!is.na(ATAC_peak_length))
          if (num_valid_peaks == 1) {
            "unique ATAC_bin"
          } else {
            "sgRNA duplicated_max ATAC_peak_length"
          }
        }
      }
    },
    sgRNA.Strand = paste(unique(sgRNA.Strand), collapse = " ** "),
    X..edits = ifelse(all(is.na(X..edits)), NA, max(X..edits, na.rm = TRUE)),
    X.silent.edits = paste(unique(X.silent.edits), collapse = " * "),
    Nucleotide.edits = paste(unique(Nucleotide.edits), collapse = " * "),
    Amino.acid.edits = paste(unique(Amino.acid.edits), collapse = " * "),
    Mutation.category = paste(unique(Mutation.category), collapse = " * "),
    Clinical.significance = paste(unique(Clinical.significance), collapse = " * "),
    .groups = "drop"
  )

collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet <- collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet %>%
  mutate(
    ATAC_bin = Gene.Symbol, # duplicate Gene.Symbol column
    ATAC_bin = case_when(
      grepl("hCRISPRi_non-targeting", ATAC_bin) ~ "NTC",
      grepl("hCRISPRa_non-targeting", ATAC_bin) ~ "NTC",
      grepl("AAVS1-controls", ATAC_bin) ~ "AAVS1.control",
      grepl("random_region", ATAC_bin) ~ "random.controls",
      TRUE ~ ATAC_bin # keep original if no match
    )
  )
```

```{r}
full_sgRNA_lib_supplement <- MSH2_CRE_sgRNA_library_parsed

full_sgRNA_lib_supplement <- full_sgRNA_lib_supplement %>%
  dplyr::mutate(library_summary = dplyr::case_when(
    df_name %in% c(
      "chr2_crisprpath_3x_enhancers",
      "0_edits_chr2_crisprpath_3x_enhancers"
    ) ~ "chr2_crisprpath_3x_enhancers",
    grepl("DNase_ATAC_matches_within_2Mb_MSH2_no_rescues", df_name) |
      grepl("DAS_rescued", df_name) ~ "cCREs",
    grepl("MSH6_exons", df_name) ~ "MSH6_exons",
    grepl("MSH2_exons", df_name) ~ "MSH2_exons",
    df_name %in% c(
      "CRISPRpath_3x_enhancers_negative_controls",
      "0_edits_CRISPRpath_3x_enhancers_negative_controls"
    ) ~ "CRISPRpath_3x_enhancers_negative_controls",
    grepl("AAVS1_negative_control", df_name) ~ df_name,
    grepl("hCRISPRa_non_targeting_control", df_name) ~ df_name,
    grepl("hCRISPRi_non_targeting_control", df_name) ~ df_name,
    grepl("random_targeting_controls", df_name) ~ df_name,
    TRUE ~ NA_character_
  ))

library_controls <- full_sgRNA_lib_supplement %>%
  dplyr::filter(grepl("control", df_name, ignore.case = TRUE))

library_non_controls <- full_sgRNA_lib_supplement %>%
  dplyr::filter(!grepl("control", df_name, ignore.case = TRUE))
```

```{r}
# ---- ISOLATED: sgRNA library distribution bar plots & tables ----
# All variables are local to this chunk; nothing written to existing objects.


# Output directory (relative to this Rmd's location)
supp_dist_dir <- "./outputs/supp_figs_sgRNA_library_distribution"
if (!dir.exists(supp_dist_dir)) dir.create(supp_dist_dir, recursive = TRUE)

# Shared theme: all text at 7pt
dist_theme <- theme_classic(base_size = 7) +
  theme(
    axis.text.x  = element_text(size = 7, angle = 45, hjust = 1),
    axis.text.y  = element_text(size = 7),
    axis.title   = element_text(size = 7),
    plot.title   = element_text(size = 7)
  )

# ── Plot 1: by library_summary ────────────────────────────────────────────────
lib_summary_counts <- full_sgRNA_lib_supplement %>%
  dplyr::count(library_summary, name = "n_sgRNAs") %>%
  dplyr::arrange(dplyr::desc(n_sgRNAs))

p_lib_summary <- ggplot(
  lib_summary_counts,
  aes(x = reorder(library_summary, -n_sgRNAs), y = n_sgRNAs)
) +
  geom_bar(stat = "identity", fill = "#4d7eb8") +
  labs(
    title = "sgRNA Library Distribution by library_summary",
    x = "library_summary", y = "Number of sgRNAs"
  ) +
  dist_theme

ggsave(file.path(supp_dist_dir, "barplot_by_library_summary.pdf"),
  plot = p_lib_summary, width = 4, height = 3, units = "in"
)

tbl_lib_summary <- gridExtra::tableGrob(
  lib_summary_counts,
  rows = NULL,
  theme = gridExtra::ttheme_default(base_size = 7)
)
pdf(file.path(supp_dist_dir, "table_by_library_summary.pdf"), width = 5, height = 3)
gridExtra::grid.arrange(tbl_lib_summary)
dev.off()

# ── Plot 2: by df_name ────────────────────────────────────────────────────────
df_name_counts <- full_sgRNA_lib_supplement %>%
  dplyr::count(df_name, name = "n_sgRNAs") %>%
  dplyr::arrange(dplyr::desc(n_sgRNAs))

p_df_name <- ggplot(
  df_name_counts,
  aes(x = reorder(df_name, -n_sgRNAs), y = n_sgRNAs)
) +
  geom_bar(stat = "identity", fill = "#4d7eb8") +
  labs(
    title = "sgRNA Library Distribution by df_name",
    x = "df_name", y = "Number of sgRNAs"
  ) +
  dist_theme

ggsave(file.path(supp_dist_dir, "barplot_by_df_name.pdf"),
  plot = p_df_name, width = 7, height = 4, units = "in"
)

tbl_df_name <- gridExtra::tableGrob(
  df_name_counts,
  rows = NULL,
  theme = gridExtra::ttheme_default(base_size = 7)
)
pdf(file.path(supp_dist_dir, "table_by_df_name.pdf"), width = 7, height = 7)
gridExtra::grid.arrange(tbl_df_name)
dev.off()

message("Saved sgRNA distribution plots and tables to: ", supp_dist_dir)
```

# Add genomic coordinate mapping for each sgRNA sequence.
```{r}
all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet <- collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet %>%
  mutate(
    sgRNA_all_SENSE = case_when(
      sgRNA.Strand == "antisense" ~ as.character(reverseComplement(DNAStringSet(sgRNA.sequence))),
      TRUE ~ sgRNA.sequence
    )
  )

all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet <- all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet %>% dplyr::select(
  sgRNA.sequence,
  sgRNA_all_SENSE,
  sgRNA.Strand,
  everything()
)
```


# Load the FASTA file containing the 2Mb region surrounding MSH2.
```{r}
full_path <- "../data/targeted_cCRE_combine_cds_preliminary_analysis/df_2Mb_surrounding_MSH2.fa"

#  1. Load the entire FASTA file
msh2_sequences <- readDNAStringSet(full_path)

# 2. Convert all sequences to uppercase (this works on the whole set)
msh2_sequences <- toupper(msh2_sequences)

msh2_fasta_final <- msh2_sequences[[1]]
```

# Load the corresponding BED file for the MSH2 region.
```{r}
full_path <- "../data/targeted_cCRE_combine_cds_preliminary_analysis/df_2Mb_surrounding_MSH2.bed"
msh2_bed <- read.table(full_path, header = FALSE, sep = "\t", stringsAsFactors = FALSE)
colnames(msh2_bed) <- c("chr", "start", "end")
```


# Map the sgRNA sequences to the reference FASTA to determine their exact genomic coordinates.

```{r}
msh2_fasta_char <- as.character(msh2_fasta_final)

map_sgRNA_to_fasta <- function(sgRNA_seq, fasta_seq) {
  match_positions <- gregexpr(sgRNA_seq, fasta_seq)[[1]]
  if (match_positions[1] != -1) {
    data.frame(
      start = match_positions,
      end = match_positions + nchar(sgRNA_seq) - 1
    )
  } else {


    NULL
  }
}


# --- Step 1: Find all positions for every sgRNA ---
sgRNA_positions_list <- lapply(
  all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet$sgRNA_all_SENSE,
  map_sgRNA_to_fasta,
  fasta_seq = msh2_fasta_char
)


# --- Step 2: Combine all matches into a single long data frame ---
names(sgRNA_positions_list) <- 1:nrow(all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet)

all_matches_df <- bind_rows(sgRNA_positions_list, .id = "original_index")


# --- Step 3: Join the matches back to the original sgRNA data ---
original_df_with_index <- all_sense_collapsed_unique_MSH2_CRE_sgRNA_library_sample_sheet %>%
  mutate(original_index = as.character(row_number()))

mapped_df <- left_join(original_df_with_index, all_matches_df, by = "original_index") %>%
  dplyr::select(-original_index) # Clean up the temporary index


# --- Step 4: Convert to genomic coordinates ---
fasta_start_coord <- msh2_bed$start[1]
final_mapped_guides <- mapped_df %>%
  mutate(
    genomic_start = start + fasta_start_coord - 1,
    genomic_end = end + fasta_start_coord - 1
  )
```

```{r}
guides_with_multiple_matches <- final_mapped_guides %>%
  add_count(sgRNA.sequence) %>%
  dplyr::filter(n > 1) %>%
  arrange(sgRNA.sequence) # Sorting makes it easy to see the groups together
```

```{r}
guides_with_atac_coords <- final_mapped_guides %>%
  extract(
    col = ATAC_bin,
    into = c("ATAC_bin_start", "ATAC_bin_end"),
    regex = "chr2:(\\d+)-(\\d+)",
    remove = FALSE,
    convert = TRUE
  )

print(head(guides_with_atac_coords))
```
# Retain sgRNA mappings strictly within their assigned ATAC_bin boundaries (controls mapped to NA are retained).
```{r}
guides_with_controls_kept <- guides_with_atac_coords %>%
  dplyr::filter(


    (genomic_start >= ATAC_bin_start & genomic_end <= ATAC_bin_end) |



      is.na(ATAC_bin_start)
  )

head(guides_with_controls_kept)
dim(guides_with_controls_kept)

guide_information_w_ATAC_bins <- guides_with_controls_kept %>% mutate(ATAC_bin_length = ATAC_bin_end - ATAC_bin_start)
head(guide_information_w_ATAC_bins)
dim(guide_information_w_ATAC_bins)
```

# Calculate mean and median lengths of the targeted ATAC_bins.
```{r}
mean(guide_information_w_ATAC_bins$ATAC_bin_length, na.rm = TRUE)
median(guide_information_w_ATAC_bins$ATAC_bin_length, na.rm = TRUE)
min(guide_information_w_ATAC_bins$ATAC_bin_length, na.rm = TRUE)
```


```{r}
ggplot(data = guide_information_w_ATAC_bins, aes(x = ATAC_bin_length)) +
  geom_histogram(binwidth = 5, fill = "coral", color = "black") + # Adjust binwidth as needed
  labs(
    title = "Histogram of ATAC Bin Length",
    x = "ATAC Bin Length",
    y = "Frequency"
  ) +
  theme_minimal()
```



# Merge valid guide mappings with the CROP samplesheet metadata.

```{r}
merged_crop_library_samplesheet <- inner_join(
  CROP_samplesheet,
  guide_information_w_ATAC_bins,
  by = "sgRNA.sequence"
)


full_path <- file.path(working_dir, "merged_crop_library_samplesheet.csv")

write.csv(
  merged_crop_library_samplesheet,
  file = full_path,
  row.names = FALSE

)



write.csv(
  merged_crop_library_samplesheet,
  file = "./outputs/merged_crop_library_samplesheet.csv",
  row.names = FALSE
)
```

# Calculate the total number of unique sgRNAs assigned to each ATAC_bin.
```{r}
sgRNAs_per_ATAC_bin <- merged_crop_library_samplesheet %>%
  group_by(ATAC_bin) %>%
  summarise(
    sgRNA_count = n_distinct(sgRNA.sequence),
    .groups = "drop"
  ) %>%
  arrange(ATAC_bin)
```

# Merge the validated sgRNA library metadata into the primary CDS colData.

```{r}
joined_df <- inner_join(coldata_CDS_w_HASH_CROP, merged_crop_library_samplesheet, by = "sgRNA")
new_colData <- as(joined_df, "DataFrame")

rownames(new_colData) <- new_colData$cell_ID.x

colData(CDS_w_HASH_CROP) <- new_colData
```

estimate_size_factors
```{r}
filtered_KRAB_cds <- estimate_size_factors(CDS_w_HASH_CROP)
```

```{r}
# isolated chunk: plot_percent_cells_positive — one PDF per gene for Illustrator assembly

pcp_allcells_genes <- c("MSH2", "MSH6", "B2M", "HPRT1", "PGK1")
pcp_allcells_outdir <- "./outputs/plots/plot_percent_cells_pos_all_cells"
dir.create(pcp_allcells_outdir, showWarnings = FALSE, recursive = TRUE)

pcp_allcells_shared_theme <- theme(
  text            = element_text(size = 7),
  axis.text       = element_text(size = 7),
  axis.title      = element_text(size = 7),
  strip.text      = element_text(size = 7),
  legend.position = "none"
)

for (gene in pcp_allcells_genes) {
  cds_sub <- filtered_KRAB_cds[rowData(filtered_KRAB_cds)$gene_short_name == gene, ]
  p <- plot_percent_cells_positive(cds_sub) +
    scale_fill_manual(values = "#00CD66") +
    pcp_allcells_shared_theme

  ggsave(
    filename = file.path(
      pcp_allcells_outdir,
      paste0("all_cells_pcp_", gene, ".pdf")
    ),
    plot = p,
    width = 1,
    height = 1.4,
    units = "in",
    device = cairo_pdf
  )
}

rm(pcp_allcells_genes, pcp_allcells_outdir, pcp_allcells_shared_theme, gene, cds_sub, p)
```

```{r}
filtered_KRAB_cds_w_bins <- filtered_KRAB_cds

colData(filtered_KRAB_cds_w_bins)$dose <- as.character(colData(filtered_KRAB_cds_w_bins)$dose)
colData(filtered_KRAB_cds_w_bins)$bin_dose <- paste(colData(filtered_KRAB_cds_w_bins)$ATAC_bin, colData(filtered_KRAB_cds_w_bins)$dose, sep = "_")
```

# Finalize the CDS objects for downstream cluster execution.

```{r}
saveRDS(filtered_KRAB_cds_w_bins, "./outputs/targeted_cCRE_combined_cds_all_genes_final.RDS")

# First, subset for your gene of interest as before

msh2_cds <- filtered_KRAB_cds_w_bins[rowData(filtered_KRAB_cds_w_bins)$gene_short_name == "MSH2", ]

saveRDS(msh2_cds, "./outputs/targeted_cCRE_combined_cds_msh2_only_final.RDS")
```


##############################################
Make UMAPs 

```{r}
UMAP_output_dir <- "./outputs/plots/UMAPs"
final_UMAP_plots_dir <- "./outputs/plots/UMAPs/final_main_figure_plots"
supplemental_UMAP_plots_dir <- "./outputs/plots/UMAPs/final_supplemental_figs"
dir.create(final_UMAP_plots_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(supplemental_UMAP_plots_dir, recursive = TRUE, showWarnings = FALSE)
```

```{r}
preprocess_KRAB_cds <- preprocess_cds(filtered_KRAB_cds_w_bins,
  method = "PCA",
  num_dim = 15
)
plot_pc_variance_explained(preprocess_KRAB_cds)

ggsave(file.path(UMAP_output_dir, "Variance_explained_15_PCs.png"))

# 4. VARIANCE EXPLAINED PLOT FOR SUPPLEMENTAL (Moved from bottom of script)
p_var <- plot_pc_variance_explained(preprocess_KRAB_cds) +
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 7)
  )

ggsave(file.path(supplemental_UMAP_plots_dir, "Variance_explained_15_PCs.pdf"),
  plot = p_var,
  width = 2.5,
  height = 1.5
)
```


```{r}
preprocess_KRAB_cds <- reduce_dimension(preprocess_KRAB_cds,
  max_components = 2,
  reduction_method = "UMAP",
  preprocess_method = "PCA",
  umap.min_dist = 0.1,
  umap.n_neighbors = 20,
  umap.metric = "cosine"
)
```



```{r}
# 1. CREATE THE BATCH PLOT
p_batch <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "sample",
  label_cell_groups = FALSE,
  cell_size = 0.2,
  alpha = 0.8
) +
  facet_wrap(~sample, scales = "fixed", nrow = 1) + # Force one row for a "Figure 1" look
  theme_void() + # Matches your MSH2 and Dose aesthetics
  theme(
    legend.position = "none",
    strip.text = element_text(size = 8, face = "bold")
  ) # Keeps sample names legible

# 2. RASTERIZE THE DOTS (600 DPI is the sweet spot for 6-inch plots)
p_batch_raster <- rasterise(p_batch, dpi = 600)

# 3. SAVE THE HYBRID PDF
# We use a height of 2.5 to keep the facets from looking "stretched" vertically
ggsave(file.path(UMAP_output_dir, "Batch_A172_Facets_6in_raster.pdf"),
  plot = p_batch_raster,
  width = 6,
  height = 2.5
)

# 4. SAVE A HIGH-RES PNG FOR SLIDES
ggsave(file.path(UMAP_output_dir, "Batch_A172_Facets_6in_600dpi.png"),
  plot = p_batch, # No need for rasterise() on a PNG
  width = 6,
  height = 2.5,
  dpi = 600
)
```



```{r}
plot_cells(preprocess_KRAB_cds, color_cells_by = "log10.umi", label_cell_groups = FALSE, cell_size = 0.5)
```





#plots below are colorblind friendly from Paul Tol vibrant palette https://sronpersonalpages.nl/~pault/
  




```{r}
# Convert dose to a factor with the correct numerical order
colData(preprocess_KRAB_cds)$dose <- factor(colData(preprocess_KRAB_cds)$dose,
  levels = c("0", "10", "50", "100")
)

# Define your named color vector - From Paul Tol color blind friendly palette
dose_colors <- c(
  "0" = "#0077BB",
  "10" = "#33BBEE",
  "50" = "#EE7733",
  "100" = "#EE3377"
)

# Create the plot
p_dose_small <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "dose",
  label_cell_groups = FALSE,
  cell_size = 0.25
) +
  theme_void() +
  ggtitle("TMZ") +
  scale_color_manual(values = dose_colors, name = expression(Dose ~ "(" * mu * M * ")")) +
  theme(
    legend.position = "right",
    text = element_text(size = 7),
    plot.title = element_text(hjust = 0.5, size = 7),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  ) +
  guides(color = guide_legend(override.aes = list(size = 2))) # Smaller legend for 2-inch plot

ggsave(file.path(UMAP_output_dir, "Dose_A172_UMAP_small_color3_900dpi.png"),
  plot = p_dose_small,
  width = 2.14,
  height = 1,
  dpi = 900
)

# --- 2. rasterized pdf DOSE PLOT  ---

p_dose_small <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "dose",
  label_cell_groups = FALSE,
  cell_size = 0.25
) +
  theme_void() +
  ggtitle("TMZ") +
  scale_color_manual(values = dose_colors, name = expression(Dose ~ "(" * mu * M * ")")) +
  theme(
    legend.position = "right",
    text = element_text(size = 7),
    plot.title = element_text(hjust = 0.5, size = 7),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  ) +
  guides(color = guide_legend(override.aes = list(size = 2))) # Smaller legend for 2-inch plot

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 600 DPI
p_dose_raster_fixed <- rasterise(p_dose_small, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "Dose_A172_UMAP_small_color3_900dpi_raster_fixed.pdf"),
  plot = p_dose_raster_fixed,
  width = 2.14,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_Dose_A172_UMAP_small_color3_900dpi_raster_fixed.pdf"),
  plot = p_dose_raster_fixed,
  width = 1.1,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_without_legend_Dose_A172_UMAP_small_color3_900dpi_raster_fixed.pdf"),
  plot = p_dose_raster_fixed + theme(legend.position = "none"),
  width = 1.1,
  height = 1
)


# --- 2. PRESENTATION DOSE PLOT (Large PNG) ---
p_dose_large <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "dose",
  label_cell_groups = FALSE,
  cell_size = 0.5
) +
  theme_void() +
  ggtitle("TMZ") +
  scale_color_manual(values = dose_colors) +
  theme(
    legend.position = "right",
    text = element_text(size = 12), # Visible on slides
    plot.title = element_text(hjust = 0.5, size = 12),
    legend.key.width = unit(0.8, "line"),
    legend.key.height = unit(0.8, "line")
  ) +
  guides(color = guide_legend(override.aes = list(size = 4)))

ggsave(file.path(UMAP_output_dir, "Dose_A172_Presentation_Large_color3.png"),
  plot = p_dose_large,
  width = 6,
  height = 5,
  dpi = 600
)
```




```{r}
# 2. Plot with better labels and rasterization
plot_cells(preprocess_KRAB_cds,
  color_cells_by = "dose",
  label_cell_groups = FALSE,
  cell_size = 0.1,
  alpha = 0.2, # Lower alpha helps see density better for 800k
  rasterize = TRUE
) + # Crucial for 800k cells
  facet_wrap(~dose, ncol = 2) +
  theme_classic(base_size = 8) +
  theme(
    legend.position = "none",
    strip.background = element_blank(), # Cleaner facet titles
    strip.text = element_text(face = "bold")
  )
```





#PLOT genes of interest

#PLOT MSH2
```{r}
plot_cells(cds = preprocess_KRAB_cds, genes = "MSH2", scale_to_range = FALSE, cell_size = .4, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)")
```

```{r}
p_small_manuscript <- plot_cells(cds = preprocess_KRAB_cds, genes = "MSH2", scale_to_range = FALSE, cell_size = 0.25, label_cell_groups = FALSE) +
  theme_void() +
  ggtitle("MSH2") +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 7),
    plot.title = element_text(hjust = 0.5, size = 7),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line"),
    strip.text = element_blank(),
    strip.background = element_blank()
  )

# Export the small high dpi png
ggsave(file.path(UMAP_output_dir, "MSH2_A172_UMAP_small_900dpi.png"),
  plot = p_small_manuscript,
  width = 2.14,
  height = 1,
  dpi = 900
)

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 900 DPI
p_raster_fixed <- rasterise(p_small_manuscript, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "MSH2_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 2.14,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_MSH2_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 1.1,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_without_legend_MSH2_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed + theme(legend.position = "none"),
  width = 1.1,
  height = 1
)


# larger .png version
p_large <- plot_cells(cds = preprocess_KRAB_cds, genes = "MSH2", scale_to_range = FALSE, cell_size = .5, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

ggsave(file.path(UMAP_output_dir, "MSH2_A172_Presentation_Large.png"),
  plot = p_large,
  width = 6,
  height = 5,
  dpi = 600
)
```



#PLOT B2M

```{r}
plot_cells(cds = preprocess_KRAB_cds, genes = "B2M", scale_to_range = FALSE, cell_size = .4, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)")
```

```{r}
p_small_manuscript <- plot_cells(cds = preprocess_KRAB_cds, genes = "B2M", scale_to_range = FALSE, cell_size = 0.25, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 6),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

# Export the small high dpi png
ggsave(file.path(UMAP_output_dir, "B2M_A172_UMAP_small_900dpi.png"),
  plot = p_small_manuscript,
  width = 2.14,
  height = 1,
  dpi = 900
)

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 900 DPI
p_raster_fixed <- rasterise(p_small_manuscript, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "B2M_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 2.14,
  height = 1
)


# larger .png version
p_large <- plot_cells(cds = preprocess_KRAB_cds, genes = "B2M", scale_to_range = FALSE, cell_size = .5, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

ggsave(file.path(UMAP_output_dir, "B2M_A172_Presentation_Large.png"),
  plot = p_large,
  width = 6,
  height = 5,
  dpi = 600
)
```


#PLOT MSH6
```{r}
plot_cells(cds = preprocess_KRAB_cds, genes = "MSH6", scale_to_range = FALSE, cell_size = .4, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)")
```

```{r}
p_small_manuscript <- plot_cells(cds = preprocess_KRAB_cds, genes = "MSH6", scale_to_range = FALSE, cell_size = 0.25, label_cell_groups = FALSE) +
  theme_void() +
  ggtitle("MSH6") +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 7),
    plot.title = element_text(hjust = 0.5, size = 7),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line"),
    strip.text = element_blank(),
    strip.background = element_blank()
  )

# Export the small high dpi png
ggsave(file.path(UMAP_output_dir, "MSH6_A172_UMAP_small_900dpi.png"),
  plot = p_small_manuscript,
  width = 2.14,
  height = 1,
  dpi = 900
)

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 900 DPI
p_raster_fixed <- rasterise(p_small_manuscript, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "MSH6_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 2.14,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_MSH6_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 1.1,
  height = 1
)

ggsave(file.path(final_UMAP_plots_dir, "width1_1_height1_without_legend_MSH6_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed + theme(legend.position = "none"),
  width = 1.1,
  height = 1
)


# larger .png version
p_large <- plot_cells(cds = preprocess_KRAB_cds, genes = "MSH6", scale_to_range = FALSE, cell_size = .5, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

ggsave(file.path(UMAP_output_dir, "MSH6_A172_Presentation_Large.png"),
  plot = p_large,
  width = 6,
  height = 5,
  dpi = 600
)
```



# Plot PGK1

```{r}
plot_cells(cds = preprocess_KRAB_cds, genes = "PGK1", scale_to_range = FALSE, cell_size = .4, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)")
```

```{r}
p_small_manuscript <- plot_cells(cds = preprocess_KRAB_cds, genes = "PGK1", scale_to_range = FALSE, cell_size = 0.25, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 6),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

# Export the small high dpi png
ggsave(file.path(UMAP_output_dir, "PGK1_A172_UMAP_small_900dpi.png"),
  plot = p_small_manuscript,
  width = 2.14,
  height = 1,
  dpi = 900
)

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 900 DPI
p_raster_fixed <- rasterise(p_small_manuscript, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "PGK1_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 2.14,
  height = 1
)


# larger .png version
p_large <- plot_cells(cds = preprocess_KRAB_cds, genes = "PGK1", scale_to_range = FALSE, cell_size = .5, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

ggsave(file.path(UMAP_output_dir, "PGK1_A172_Presentation_Large.png"),
  plot = p_large,
  width = 6,
  height = 5,
  dpi = 600
)
```



# PLOT HPRT1
```{r}
plot_cells(cds = preprocess_KRAB_cds, genes = "HPRT1", scale_to_range = FALSE, cell_size = .4, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)")
```

```{r}
p_small_manuscript <- plot_cells(cds = preprocess_KRAB_cds, genes = "HPRT1", scale_to_range = FALSE, cell_size = 0.25, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 6),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

# Export the small high dpi png
ggsave(file.path(UMAP_output_dir, "HPRT1_A172_UMAP_small_900dpi.png"),
  plot = p_small_manuscript,
  width = 2.14,
  height = 1,
  dpi = 900
)

# 3. THE FIX: Wrap the plot object in rasterise()
# This forces the points to flatten at 900 DPI
p_raster_fixed <- rasterise(p_small_manuscript, dpi = 900)

# 4. Save the rasterized version
ggsave(file.path(UMAP_output_dir, "HPRT1_A172_UMAP_small_900dpi_raster.pdf"),
  plot = p_raster_fixed,
  width = 2.14,
  height = 1
)


# larger .png version
p_large <- plot_cells(cds = preprocess_KRAB_cds, genes = "HPRT1", scale_to_range = FALSE, cell_size = .5, label_cell_groups = FALSE) +
  theme_void() +
  viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    legend.key.width = unit(0.5, "line"),
    legend.key.height = unit(0.5, "line")
  )

ggsave(file.path(UMAP_output_dir, "HPRT1_A172_Presentation_Large.png"),
  plot = p_large,
  width = 6,
  height = 5,
  dpi = 600
)
```















```{r}
# Define supplemental output directory


# 1. NEW BATCH PLOTS FOR SUPPLEMENTAL
# Reorder the sample factor as requested
colData(preprocess_KRAB_cds)$sample <- factor(colData(preprocess_KRAB_cds)$sample,
  levels = c(
    "KRAB_new_DEPC_premade_probes",
    "targeted_sci_KRAB_5_28_25",
    "KRAB_7_16_25_P7cols_2_7_8_11_12"
  )
)

# --- Faceted version ---
p_batch_supp <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "sample",
  label_cell_groups = FALSE,
  cell_size = 0.03
) +
  facet_wrap(~sample, scales = "fixed", nrow = 1) +
  theme_void() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 7, face = "plain")
  ) # Size 7, not bold

p_batch_supp_raster <- rasterise(p_batch_supp, dpi = 900)

ggsave(file.path(supplemental_UMAP_plots_dir, "width3_09_height0_97_Batch_A172_Facets_900dpi_raster.pdf"),
  plot = p_batch_supp_raster,
  width = 3.09,
  height = 0.97
)

# --- Overlapping version (No Facets) ---
p_batch_overlap <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "sample",
  label_cell_groups = FALSE,
  cell_size = 0.025,
  alpha = 0.6
) +
  ggtitle("sci reps") + # Updated title
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 7), # Size 7, not bold
    legend.position = "none"
  )

p_batch_overlap_raster <- rasterise(p_batch_overlap, dpi = 900)

ggsave(file.path(supplemental_UMAP_plots_dir, "width1_height1_sci_reps_A172_Overlapping_900dpi_raster.pdf"),
  plot = p_batch_overlap_raster,
  width = 1,
  height = 1
)


# 2. DOSE PLOT FOR SUPPLEMENTAL
# Create a formatted dose factor with plotmath expression for 'µM' safely, robust to re-running the chunk
colData(preprocess_KRAB_cds)$dose_formatted <- factor(
  gsub(" ~ mu \\* M", "", gsub(" \u03BCM", "", gsub(" \u00B5M", "", as.character(colData(preprocess_KRAB_cds)$dose)))),
  levels = c("0", "10", "50", "100")
)
levels(colData(preprocess_KRAB_cds)$dose_formatted) <- paste0(
  levels(colData(preprocess_KRAB_cds)$dose_formatted), " ~ mu * M"
)

dose_colors <- c(
  "0 ~ mu * M" = "#0077BB",
  "10 ~ mu * M" = "#33BBEE",
  "50 ~ mu * M" = "#EE7733",
  "100 ~ mu * M" = "#EE3377"
)

p_dose_supp <- plot_cells(preprocess_KRAB_cds,
  color_cells_by = "dose_formatted",
  label_cell_groups = FALSE,
  cell_size = 0.03,
  alpha = 0.3
) +
  facet_wrap(~dose_formatted, nrow = 1, labeller = label_parsed) +
  theme_void() +
  scale_color_manual(values = dose_colors) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 7, face = "plain")
  )

p_dose_supp_raster <- rasterise(p_dose_supp, dpi = 900)

ggsave(file.path(supplemental_UMAP_plots_dir, "width4_12_height1_Dose_A172_Facets_900dpi_raster.pdf"),
  plot = p_dose_supp_raster,
  width = 4.12,
  height = 1
)


# 3. GENE PLOTS FOR SUPPLEMENTAL
# List of genes to plot
genes_to_plot <- c("MSH6", "B2M", "PGK1", "HPRT1")

for (gene_name in genes_to_plot) {
  # Create the plot
  p_gene <- plot_cells(
    cds = preprocess_KRAB_cds,
    genes = gene_name,
    scale_to_range = FALSE,
    cell_size = 0.25,
    label_cell_groups = FALSE
  ) +
    theme_void() +
    ggtitle(gene_name) +
    viridis::scale_color_viridis(option = "magma", name = "log10(norm\n expr + 0.1)") +
    theme(
      legend.position = "right",
      text = element_text(size = 7),
      plot.title = element_text(hjust = 0.5, size = 7),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.key.width = unit(0.5, "line"),
      legend.key.height = unit(0.5, "line"),
      strip.text = element_blank(),
      strip.background = element_blank()
    )

  # Rasterize the plot at 900 DPI
  p_gene_raster <- rasterise(p_gene, dpi = 900)

  # Save the plot with 1.55 inch width and 1.05 inch height
  file_name <- paste0("width1_55_height1_05_", gene_name, "_A172_UMAP_small_900dpi_raster.pdf")
  ggsave(file.path(supplemental_UMAP_plots_dir, file_name),
    plot = p_gene_raster,
    width = 1.55,
    height = 1.05
  )
}

```
