---
title: "Aim2_MSH2_CRE_analysis"
output: html_document
date: "2024-07-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressPackageStartupMessages({
  library(tidyr)
  library(readr)
  library(dplyr)
  library(ggplot2)
  library(stringr)
  library(monocle3)
})
```

## Initializing CDS Object
```{r}
cds.pre <- readRDS("../data/targeted_cCRE_1/GSM9825201_targeted_cCRE_1_cDNA_cds_precell_prehash.RDS")

cds_coldata_df <- colData(cds.pre) %>% as.data.frame()
dim(cds_coldata_df)
```



```{r}
# knee plot for UMIs, visualizing cutoff at 12
# horizontal line is at log10(12), i.e. n.umi = 12, matching the UMI cutoff below

# ggrastr::rasterise() rasterizes only the points at 600 DPI;
# all text, axes, hline, and title remain as editable vectors in the PDF
umi_knee_plot <- ggplot(cds_coldata_df %>% arrange(desc(n.umi)) %>% mutate(cell_rank = dplyr::row_number()), aes(x= log10(cell_rank), y = log10(n.umi))) +
  geom_hline(size = 0.2, yintercept = log10(12)) + #input what n.umi you want to cutoff for visualization
  ggrastr::rasterise(geom_point(size = 0.5, stroke = 0), dpi = 600) +
  monocle3:::monocle_theme_opts() +
  scale_y_continuous(breaks = c(1,2,3,4,5)) +
  xlab("Log10 CellRank") +
  ylab("Log10 UMI") +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1,"line"),
    legend.key.height = unit(0.2,"line"),
    text = element_text(size = 6)
  )

# print for notebook display (unchanged)
print(umi_knee_plot)

# save PDF copy for paper figures (2in wide, 7pt fonts, rasterized points)
umi_knee_plot_paper <- umi_knee_plot +
  theme(text = element_text(size = 7))

output_dir <- "./outputs/"
dir.create(file.path(output_dir, "cds_knee_plots"), showWarnings = FALSE, recursive = TRUE)

ggsave(
  filename = file.path(output_dir, "cds_knee_plots/targeted_cCRE_1_knee_plot.pdf"),
  plot     = umi_knee_plot_paper,
  width    = 2,
  height   = 2,
  units    = "in",
  device   = cairo_pdf
)

rm(umi_knee_plot_paper, knee_plot_sample_title)
```



```{r}
cds.pre <- cds.pre[,colData(cds.pre)$n.umi >= 12]
dim(cds.pre)

```



```{r}
cols <- colData(cds.pre) %>% as.data.frame() %>%
  mutate(cell_ID=Cell) %>%
  tidyr::separate(col = Cell, into = c('P7', 'P5', NA, NA, 'n1', NA, NA, 'n2'), sep = "_") %>%
  mutate(RT = paste('RT_BC_', n1, sep = '')) %>%
  mutate(Lig = paste('Lig_BC_', n2, sep = '')) %>%
  select(-n1, -n2)
```



## Merging Hash Data into CDS










```{r}
hash_table <- read.table(gzfile("../data/targeted_cCRE_1/GSM9825201_targeted_cCRE_1_hashTable.out.txt.gz"), col.names = c("sample", "cell_ID", "oligo", "hash_umis"))
```








Calculate Hash Proportions and Top-to-Second-Best Ratios. 
Note: Single-hash cells are assigned a default ratio of 500 to pass subsequent ratio filters.

```{r}
hashTable_summary_raw <- hash_table %>%
  group_by(cell_ID) %>%
  mutate(proportion = hash_umis/sum(hash_umis),total_hash_umis_per_cell_ID = sum(hash_umis)) %>%
  arrange(desc(proportion)) %>%
  mutate(rank = row_number()) %>%
  dplyr::filter(rank %in% c(1,2)) %>%
  mutate(top_to_second_best_ratio = ifelse(sum(rank) > 1, proportion[rank == 1]/proportion[rank == 2], 500))%>%
  dplyr::filter(rank == 1)

unique(hashTable_summary_raw$rank)
```






```{r}
# Filter the hashTable data frame to keep only rows that have a hash_umis value of 4 or more
filtered_hashTable <- hashTable_summary_raw %>%
  filter(hash_umis >= 4)
```



Filter `hash_table_summary` based on `top_to_second_best_ratio` threshold.
```{r}
# Filter the hashTable_summary data frame to keep only rows that meet the threshold
hashTable_summary <- filtered_hashTable %>%
  filter(top_to_second_best_ratio >= 2)
```



























```{r}
# Load the tidyverse package
library(tidyverse)


# Duplicate the oligo column
hashTable_summary <- hashTable_summary %>%
  mutate(oligo_temp = oligo)

# Split the oligo_temp column
hashTable_summary <- hashTable_summary %>%
  separate(oligo_temp, into = c("cell_line_part1", "cell_line_part2", "dose", "other1", "drug", "well_ID"), sep = "_") %>%
  mutate(cell_line = paste(cell_line_part1, cell_line_part2, sep = "_"),
         dose = as.numeric(dose)) %>%
  select(-other1,-cell_line_part2,-cell_line_part1)
```

```{r}
# Convert tibble to dataframe
hashTable_summary <- as.data.frame(hashTable_summary)
```

Merge pre-hash CDS with processed hash information.
```{r}
dim(cols)

cols <- cols %>%
  left_join(hashTable_summary, by = c("cell_ID" = "cell_ID", "sample" = "sample")) %>%
  select(-Size_Factor, -proportion, -rank)
```


```{r}
rownames(cols) <- cols$cell_ID
```

```{r}
# assiging our cols dataframe as colData(cds)
colData(cds.pre) <- DataFrame(cols)
```

```{r}
# adding columns for log10(n.umi) and % mito
colData(cds.pre)$log10.umi <- colData(cds.pre)$n.umi %>% log10()

mt_genes <- rowData(cds.pre) %>% as.data.frame() %>%
  filter(grepl("^MT-",gene_short_name) == TRUE)

mt_genes_id <- rownames(mt_genes)
colData(cds.pre)$percent_mito <- 100 * (colSums(exprs(cds.pre)[mt_genes_id,])/colSums(exprs(cds.pre)))

```

```{r}
cds_w_hash_df <- colData(cds.pre) %>% as.data.frame()
```

Quantify cells missing hash assignments.
```{r}
# Count rows with NA in the hash_umis column
na_count_hash <- cds_w_hash_df %>%
  filter(is.na(hash_umis))

na_count_hash

#this is the percent of cells in the cds that do not have hash assigned
nrow(na_count_hash)/nrow(cds_w_hash_df)*100

nrow(cds_w_hash_df)-nrow(na_count_hash)
```










## Processing CROP-seq sgRNA Data








```{r}
# Read in crop table
cropTable <- read.table(gzfile("../data/targeted_cCRE_1/GSM9825202_targeted_cCRE_1_CROPTable.out.txt.gz"), col.names = c("sample", "cell_ID", "sgRNA", "axis", "reads"))

arrange(cropTable, desc(reads))
```

```{r}
length(unique(cropTable$cell_ID))
```







Parse `cell_ID` to extract P7 and P5 indices for filtering.

```{r}
cropTable <- cropTable %>%
  mutate(
    split_parts = str_split(cell_ID, "_"),
    P7 = sapply(split_parts, `[`, 1),
    P5 = sapply(split_parts, `[`, 2)
  ) %>%
  select(-split_parts)  # Clean up

```



```{r}
unique(cropTable$P7)
```


```{r}
length(unique(cropTable$sgRNA))
```
Filter for relevant P7 indices corresponding to the target sequencing run.
```{r}
cropTable_real_P7s <- cropTable %>%
  filter(P7 %in% c( "10G","10H"))
```

Process CROP dataframe and map valid P7 indices.
```{r}
cropTable_mapped <- cropTable_real_P7s

cropTable_mapped <- cropTable_mapped %>%
  mutate(new_cell = case_when(
    grepl("10H", cell_ID) ~ gsub("10H", "09", cell_ID),
    grepl("10G", cell_ID) ~ gsub("10G", "10", cell_ID),
    TRUE ~ cell_ID
  ))
```




```{r}
cds_w_hash <- cds.pre
```




Standardize cell IDs by stripping the 3rd character from the P7 prefix.
```{r}
# if you did columns of P7
colData(cds_w_hash)$new_cell <- gsub("^(.{2})(.)(.*)$", "\\1\\3", colData(cds_w_hash)$cell_ID)
```

Initialize `cds_w_hash_filtered` for downstream merging.
```{r}
cds_w_hash_filtered <- cds_w_hash
```

```{r}
cds_hash_filtered_df <- colData(cds_w_hash_filtered) %>% as.data.frame()
```

Calculate percentage of unassigned cells.
```{r}
# Count rows with NA in the hash_umis column
na_count_hash <- cds_hash_filtered_df %>%
  filter(is.na(hash_umis))

# this is the percent of cells in the cds that do not have hash assigned
nrow(na_count_hash) / nrow(cds_hash_filtered_df) * 100
```

Filter `cropTable` to retain only cells present in the CDS.
```{r}
crop_mapped_cds_intersect <- cropTable_mapped
crop_mapped_cds_intersect <- crop_mapped_cds_intersect[crop_mapped_cds_intersect$new_cell %in% colData(cds_w_hash_filtered)$new_cell, ]
```

Check unique cell IDs present in CDS with an assigned CROP sgRNA.
```{r}
length(unique(crop_mapped_cds_intersect$cell_ID))
```
```{r}
length(unique(crop_mapped_cds_intersect$sgRNA))
```



Calculate sgRNA proportions and top-to-second-best ratios. 
Single sgRNA cells receive a default ratio of 100 to pass subsequent ratio filters.

```{r}
cropTable_summary_raw <- crop_mapped_cds_intersect %>%
  group_by(new_cell) %>%
  mutate(sgRNA_proportion = reads/sum(reads),total_sgrna_read_per_cell = sum(reads)) %>%
  dplyr::arrange(desc(sgRNA_proportion)) %>%
  mutate(rank = row_number()) %>%
  dplyr::filter(rank %in% c(1,2)) %>%
  mutate(sgRNA_top_to_second_best_ratio = ifelse(sum(rank) > 1, sgRNA_proportion[rank == 1]/sgRNA_proportion[rank == 2], 100)) %>%
  dplyr::filter(rank == 1)
```


```{r}
# Filter the CROPTable data frame to keep only rows that have a reads value of 2 or more
cropTable_filtered_reads <- cropTable_summary_raw %>%
  filter(reads >= 2)
```


Assign final CROP summary table.
```{r}
cropTable_summary <- cropTable_filtered_reads
```










Merge CROP data with hash-annotated CDS.
```{r}
cols_CDS_hash <- colData(cds_w_hash_filtered) %>% as.data.frame()
```
```{r}
cols_merge_CROP <- cols_CDS_hash %>%
  left_join(cropTable_summary, by = c("new_cell" = "new_cell", "sample" = "sample"))
```

Use `cell_ID.x` as row name because it matches the CDS expression matrix.
```{r}
rownames(cols_merge_CROP) <- cols_merge_CROP$cell_ID.x
```


Verify row name alignment.
```{r}
identical(colnames(cds_w_hash_filtered), row.names(cols_merge_CROP))
```


```{r}
# assigning our cols dataframe as colData(cds)
colData(cds_w_hash_filtered) <- DataFrame(cols_merge_CROP)
```

```{r}
head(exprs(cds_w_hash_filtered))
```


```{r}
print(dim(cds_w_hash_filtered))
print(dim(cds_w_hash_filtered[,!is.na(colData(cds_w_hash_filtered)$sgRNA)])) # cells with a sgRNA 
```

Quantify percentage of cells with successful CROP assignment.
```{r}
cds_hash_crop_df <- colData(cds_w_hash_filtered) %>% as.data.frame()
no_NA_CROP <- cds_hash_crop_df %>%
  filter(!is.na(sgRNA))

no_NA_CROP

nrow(no_NA_CROP)/nrow(cds_hash_crop_df)*100
```






```{r}
# First remove cells from CDS that have no CROP assignment

# Define the condition to exclude cells with NA in the 'sgRNA' column
condition_to_exclude_sgRNA_NA <- is.na(colData(cds_w_hash_filtered)$sgRNA)

# Filter the cells based on the condition
cds_crop_assigned <- cds_w_hash_filtered[, !condition_to_exclude_sgRNA_NA]

# Verify the result
colData(cds_crop_assigned) %>% as.data.frame()




# remove cells from CDS that have no HASH assignment

# Define the condition to exclude cells with NA in the 'sgRNA' column
condition_to_exclude_HASH_NA <- is.na(colData(cds_crop_assigned)$oligo)

# Filter the cells based on the condition
cds_final <- cds_crop_assigned[, !condition_to_exclude_HASH_NA]

# Verify the result
cds_final_df<-colData(cds_final) %>% as.data.frame()
cds_final_df


nrow(colData(cds_crop_assigned))-nrow(cds_final_df) # Cells lost due to hash post filtering for crop
```


```{r}
# umi for cdna
median(cds_final_df$n.umi)
mean(cds_final_df$n.umi)

# hash umi
median(cds_final_df$hash_umis)
mean(cds_final_df$hash_umis)

# crop umi (aka reads)
median(cds_final_df$reads)
mean(cds_final_df$reads)

# Number unique sgRNAs
length(unique(cds_final_df$sgRNA))
```




```{r}
saveRDS(cds_final, file.path(output_dir, "targeted_cCRE_1_final.RDS"))
```

```{r}
dim(cds_final)
```






