---
title: "boostrapped_lollipop_plots"
output: html_document
date: "2025-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
library(Biostrings)   
library(dplyr)        
library(tidyr)        
library(stringr)     
library(readr)
library(ggplot2)
library(htmlwidgets)
library(plotly)
library(SummarizedExperiment)
library(Matrix)
library(DESeq2)
library(monocle3)
library(janitor)
```



```{r}
# Load 40-cell sliding window analysis outputs
inputs_dir <- "./outputs/sliding_window_40cells"
modeling_results_conf_int <- read.csv(file.path(inputs_dir, "modeling_results_conf_int_40cells.csv"))
significant_windows_only <- read.csv(file.path(inputs_dir, "significant_windows_only_40cells.csv"))
all_sgRNA_across_significant_windows <- read.csv(file.path(inputs_dir, "all_sgRNA_across_significant_windows_40cells.csv"))
```


##########  START random local controls for regions that are not significant hits 

These will serve as negative control guides for the TMZ survival screen follow up experiment of targete-cCRE nominal hits
these sgRNAs ideally target regions that do not have an effect on MSH2 expression in the targeted-cCRE screen

Steps: choose random local control for guides that are not in the ATAC bins of any significant window, with the smallest normalized effect and confidence interval that contain 0 


```{r}
random_local_controls <- modeling_results_conf_int %>%
  # 1. Keep only rows from the first dataframe with no match in the second
  anti_join(significant_windows_only, by = "ATAC_bin") %>%
  group_by(window) %>%
  
  # 3. Keep only entire window groups that meet your criteria across all rows
  filter(
    all(abs(mean_normalized_effect) < 0.1, na.rm = TRUE) &
    all(confidence_interval_contain_zero == "yes_contains_0")
  ) %>%
  ungroup() %>%

  arrange(num_unique_sgrnas)
```


```{r}
length(unique(random_local_controls$window))
```




plot the potential local controls that seem to have very little signal
```{r}
# # --- 1. Define and Validate the ATAC_bin Subset ---

local_controls <- random_local_controls

# --- 2. Prepare the 'local_controls' Data for Plotting ---
control_bins_sub <- local_controls %>%
  filter(is.na(window_start)) %>%
  arrange(window)
experimental_bins_sub <- local_controls %>%
  filter(!is.na(window_start)) %>%
  arrange(window_start)
final_bin_order_sub <- c(
  unique(control_bins_sub$window),
  unique(experimental_bins_sub$window)
)
local_controls$window <- factor(
  local_controls$window,
  levels = final_bin_order_sub
)
local_controls$dose <- factor(
  local_controls$dose,
  levels = sort(unique(local_controls$dose), decreasing = TRUE)
)

# --- Helper Function: Generate Distinct Colors ---
generate_distinct_colors <- function(n) {
  golden_ratio_conjugate <- 0.61803398875
  h <- runif(1)
  hues <- (h + (1:n) * golden_ratio_conjugate) %% 1
  sv_pairs <- list(
    c(s = 0.9, v = 0.85), c(s = 0.4, v = 0.95), c(s = 0.9, v = 0.6),
    c(s = 0.25, v = 0.98), c(s = 0.7, v = 0.7)
  )
  s_values <- rep(sapply(sv_pairs, `[`, 1), length.out = n)
  v_values <- rep(sapply(sv_pairs, `[`, 2), length.out = n)
  hsv(h = hues, s = s_values, v = v_values)
}

# --- 3. Create a High-Contrast Master Color Key for the Subset ---
ordered_bins_sub <- local_controls %>%
  arrange(window) %>%
  pull(ATAC_bin) %>%
  unique()
num_bins_sub <- length(ordered_bins_sub)
distinct_colors_sub <- generate_distinct_colors(n = num_bins_sub)
atac_color_map_sub <- data.frame(
  ATAC_bin = ordered_bins_sub,
  bin_color = distinct_colors_sub
)

# Use a different column name temporarily to avoid issues with the existing 'bin_color'
atac_color_map_sub <- rename(atac_color_map_sub, bin_color_sub = bin_color)

local_controls <- local_controls %>%
  left_join(atac_color_map_sub, by = "ATAC_bin")


# --- 4. Define the Custom Color Scale for the Subset ---
# Calculate the threshold from the top 5% of all normalized effects
mean_norm_effect_threshold <- quantile(
  abs(modeling_results_conf_int$mean_normalized_effect),
  probs = 0.95,
  na.rm = TRUE
)
min_effect_sub <- min(local_controls$mean_normalized_effect, na.rm = TRUE)
max_effect_sub <- max(local_controls$mean_normalized_effect, na.rm = TRUE)
range_effect_sub <- max_effect_sub - min_effect_sub
norm_neg_threshold_sub <- (-mean_norm_effect_threshold - min_effect_sub) / range_effect_sub
norm_pos_threshold_sub <- (mean_norm_effect_threshold - min_effect_sub) / range_effect_sub
custom_colorscale_sub <- list(
  list(0, "blue"),
  list(norm_neg_threshold_sub, "lightgrey"),
  list(norm_pos_threshold_sub, "lightgrey"),
  list(1, "red")
)

# --- 5. Build the Header and Lollipop Plots for the Subset ---
header_data_sub <- local_controls %>%
  select(window, ATAC_bin, bin_color_sub) %>%
  distinct()
p_header_sub <- plot_ly(data = header_data_sub, x = ~window, y = 1, type = 'bar',
                    marker = list(color = ~bin_color_sub),
                    hoverinfo = 'text', text = ~paste("ATAC Bin:", ATAC_bin)) %>%
                layout(yaxis = list(showticklabels = FALSE, title = ""))

dose_plots_sub <- list()
dose_levels_sub <- levels(local_controls$dose)

for(d in dose_levels_sub) {
  dose_data <- filter(local_controls, dose == d)
  dose_data_na <- filter(dose_data, is.na(mean_normalized_effect))
  dose_data_non_na <- filter(dose_data, !is.na(mean_normalized_effect))

  p_dose <- plot_ly(showlegend = FALSE) %>%
    add_segments(data = dose_data_non_na, x = ~window, y = 0,
                 xend = ~window, yend = ~mean_normalized_effect,
                 color = I("gray"), hoverinfo = "none") %>%
    add_markers(
      data = dose_data_non_na, x = ~window, y = ~mean_normalized_effect,
      marker = list(color = ~mean_normalized_effect, colorscale = custom_colorscale_sub,
                    cmin = min_effect_sub, cmax = max_effect_sub,
                    showscale = (d == "100"),
                    colorbar = list(title = "Norm Effect")),
      hoverinfo = 'text',
      text = ~paste0(
        "Effect: ", round(mean_normalized_effect, 3), "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "CI contains 0: ", confidence_interval_contain_zero, "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    add_markers(
      data = dose_data_na, x = ~window, y = 0,
      marker = list(color = "black"),
      hoverinfo = 'text',
      text = ~paste0(
        "Effect: NA", "<br>",
        "CI: ", round(conf_low_95, 3), " - ", round(conf_high_95, 3), "<br>",
        "CI contains 0: ", confidence_interval_contain_zero, "<br>",
        "Cell Count: ", cell_count, "<br>",
        "Unique sgRNAs: ", num_unique_sgrnas, "<br>",
        "Status: ", status
      )
    ) %>%
    add_annotations(text = paste(d), x = 1, y = 0.5,
                    xref = "paper", yref = "paper",
                    xanchor = "left", showarrow = FALSE,
                    font = list(size = 14))

  dose_plots_sub[[as.character(d)]] <- p_dose
}

# --- 6. Combine All Subset Plots ---
final_plot_sub <- subplot(
  p_header_sub, dose_plots_sub[["100"]], dose_plots_sub[["50"]],
  dose_plots_sub[["10"]], dose_plots_sub[["0"]],
  nrows = 5, heights = c(0.1, 0.225, 0.225, 0.225, 0.225),
  shareX = TRUE, titleY = TRUE
) %>%
  layout(
    xaxis = list(
      title = "Genomic Window (Subset)",
      categoryorder = "array",
      categoryarray = final_bin_order_sub,
      tickangle = 60
    ),
    title = list(text = "Dose-Dependent Effect for flow sub libary options", y = 0.99)
  )

# --- 7. Save the Final Interactive Subset Plot ---
htmlwidgets::saveWidget(
  widget = final_plot_sub,
  file = "./outputs/TMZ_survival_screen_sgRNA_library/interactive_lollipop_local_control_options_for_flow.html"
)
```


local controls to use - randomly sample sgRNAs from these windows:
windows were chosen from manual inspection of interactive_lollipop_local_control_options_for_flow.html
```{r}
# 1. Define the specific windows you want to select
target_windows <- c(
  "chr2:48440510-48440660",
  "chr2:48314106-48314256",
  "chr2:47176827-47176977",
  "chr2:47176152-47176302",
  "chr2:47176102-47176252",
  "chr2:47175652-47175802",
  "chr2:47175902-47176052",
  "chr2:47175677-47175827",
  "chr2:47175877-47176027",
  "chr2:46542573-46542723",
  "chr2:47368968-47369118"
)

# 2. Check for any windows that are in your list but not in the dataframe
# setdiff() finds the elements in the first vector that are not in the second
missing_windows <- setdiff(target_windows, random_local_controls$window)

# If the list of missing windows has any items, throw a warning
if (length(missing_windows) > 0) {
  warning("The following target windows were not found: ", 
          paste(missing_windows, collapse = ", "))
}

# 3. Filter the dataframe to keep only the rows with matching windows
final_local_controls_windows <- random_local_controls %>%
  filter(window %in% target_windows)

length(unique(final_local_controls_windows))

# Filter for dose == 0 and select your desired columns
final_controls_dose_zero <- final_local_controls_windows %>%
  filter(dose == 0) %>%
  select(ATAC_bin, window, dose, sgRNA_per_window)

length(unique(final_controls_dose_zero$ATAC_bin))



# Convert the comma-separated column to a long format
long_format_local_controls <- final_controls_dose_zero %>%
  separate_rows(sgRNA_per_window, sep = ",")



#make a bedfile so you can upload to guidescan

# 1. Start with your dataframe and create the three BED columns
bed_file_data <- long_format_local_controls %>%
  
  # Keep only the unique window strings
  distinct(window) %>%
  
  # Parse the 'window' column into three new columns
  separate(
    col = window, 
    into = c("chr", "start", "end"), 
    sep = "[:-]", 
    convert = TRUE
  )

# 2. Define the full output file path
# (Note: This is commented out because we already provide the GuideScan results that we load in below. 
# However, this is the bed file you would need to generate if you wanted to run GuideScan yourself)
#output_file_path <- "./outputs/TMZ_survival_screen_sgRNA_library/local_control_guide_bedfiles_for_guidescan.bed"

# # 3. Save the data as a proper BED file (tab-separated, no header)
# write.table(
#   bed_file_data,
#   file = output_file_path,
#   sep = "\t",
#   quote = FALSE,
#   row.names = FALSE,
#   col.names = FALSE
# )

##end making bedfile

#load in guidescan dataframe
file_path <- "../data/TMZ_survival_screen/results_guidescan.csv"

guidescan_data <- read_csv(file_path) %>%
  clean_names()

print(colnames(guidescan_data))

unique(str_length(guidescan_data$g_rna_seq))
length(unique(guidescan_data$region_name))
unique(guidescan_data$pam)

head(guidescan_data)


# Perform the left join to add guidescan data
guidescan_local_controls <- left_join(
  long_format_local_controls,
  guidescan_data,
  by = c("window" = "region_name", "sgRNA_per_window" = "g_rna_seq")
)

length(unique(guidescan_local_controls$window)) # want this to stay the same after I filter for specificity > 0.2
# See how many unique guides remain
n_distinct(guidescan_local_controls$sgRNA_per_window)


#filter to keep guides with specificty >=0.2
# Filter the merged dataframe to keep only guides with high specificity
CFD_filtered_guidescan_local_controls <- guidescan_local_controls %>%
  # Keep rows where the specificity score is 0.2 or greater
  filter(specificity >= 0.2)

n_distinct(CFD_filtered_guidescan_local_controls$sgRNA_per_window)
length(unique(CFD_filtered_guidescan_local_controls$window)) 
```


```{r}
#now randomly sample to get 15 local random controls:

# Stage 1: Create a dataframe containing the randomly chosen unique sgRNAs for each ATAC_bin.
sampled_sgrnas <- CFD_filtered_guidescan_local_controls %>%
  # Get the unique sgRNAs for each bin
  distinct(ATAC_bin, sgRNA_per_window) %>%
  # Now, from that unique list, sample 3
  group_by(ATAC_bin) %>%
  slice_sample(n = 3) %>%
  ungroup()

# Stage 2: Use the list from Stage 1 to retrieve the full rows
# from the original dataframe.
final_randomly_chosen_local_controls_CFD_filtered <- CFD_filtered_guidescan_local_controls %>%
  # Keep only rows that match one of our chosen sgRNAs for that bin
  semi_join(sampled_sgrnas, by = c("ATAC_bin", "sgRNA_per_window")) %>%
  
  # If a chosen sgRNA appears in multiple windows, this randomly picks ONE
  group_by(ATAC_bin, sgRNA_per_window) %>%
  slice_sample(n = 1) %>%
  ungroup()

length(unique(final_randomly_chosen_local_controls_CFD_filtered$ATAC_bin))
length(unique(final_randomly_chosen_local_controls_CFD_filtered$window))
length(unique(final_randomly_chosen_local_controls_CFD_filtered$sgRNA_per_window))


```


only save final_randomly_chosen_local_controls_CFD_filtered once because otherwise it is random so it will change each time. And then read back in the one I have statically saved to use (static one should have been made 9/15/25 at 10:30 pm)
```{r}

# 1. Define the full path for your output file
# (Note: You can save the randomly chosen local controls you generated if you want, 
# however, it is random so it will be different than the final ones chosen for our library. 
# Therefore, we load in the final randomly chosen local controls we generated below.)
# output_path <- "./outputs/TMZ_survival_screen_sgRNA_library/final_randomly_chosen_local_controls_CFD_filtered.csv"



# --- STEP 2:  Run this part ONCE to save your random sample --- Comment out the write.csv command after running it ---
# This prevents you from accidentally overwriting your static file in the future.
# write.csv(
#   final_randomly_chosen_local_controls_CFD_filtered,
#   file = output_path,
#   row.names = FALSE
# )


# --- STEP 3: Use this line to read the static file back in for all future analysis ---
input_path <- "../data/TMZ_survival_screen/final_randomly_chosen_local_controls_CFD_filtered.csv"


readcsv_final_randomly_chosen_local_controls_CFD_filtered <- read_csv(
  input_path,
  show_col_types = FALSE
)

# Remove the specific dataframe from your environment so I can't accidentally use it downstream, use readcsv_final_randomly_chosen_local_controls_CFD_filtered instead
rm(final_randomly_chosen_local_controls_CFD_filtered)
```


######### END choosing random local control sgRNAs from regions that were not significant hits




##########################
calculating # of sgRNAs from significant_windows_only

```{r}

all_sgRNA_across_significant_windows <- significant_windows_only %>%
  select(ATAC_bin, window, window_start, window_end, sgRNA_per_window) %>%
  separate_rows(sgRNA_per_window, sep = ",") %>%
  rename(sgRNA.sequence = sgRNA_per_window)

head(all_sgRNA_across_significant_windows)

# Get the final count of unique sgRNAs
num_unique_sgrnas <- length(unique(all_sgRNA_across_significant_windows$sgRNA.sequence))

print(paste("Total number of unique sgRNAs across all significant windows before getting rid of random targeting controls:", num_unique_sgrnas))


```



```{r}
#Get rid of random targeting controls
drop_random_controls <- all_sgRNA_across_significant_windows %>% filter(ATAC_bin != "random.controls")

experimental_sgRNA_pool_642_guides <- drop_random_controls %>% select(ATAC_bin, sgRNA.sequence, window)

#Check that there are 642 unique sgRNAs 
print(paste("Total number of unique sgRNAs across all significant windows:", length(unique(experimental_sgRNA_pool_642_guides$sgRNA.sequence))))

#number of unique windows
print(paste("Total number of unique  significant windows:", length(unique(experimental_sgRNA_pool_642_guides$window))))

#number of unique ATAC bin
print(paste("Total number of unique  significant ATAC_bins:", length(unique(experimental_sgRNA_pool_642_guides$ATAC_bin))))


```
these should be the numbers
[1] "Total number of unique sgRNAs across all significant windows after dropping random targeting controls: 642"
[1] "Total number of unique  significant windows after dropping random targeting controls: 121"
[1] "Total number of unique  significant ATAC_bins after dropping random targeting controls: 36"

```{r}
significant_ATAC_bin_for_bed <- drop_random_controls %>% select(ATAC_bin) %>% distinct() 
dim(significant_ATAC_bin_for_bed)
head(significant_ATAC_bin_for_bed)


#parse ATAC_bin column to get in bed file format
ATAC_parsed <- significant_ATAC_bin_for_bed %>%
  separate(
    col = ATAC_bin, 
    into = c("chr", "ATAC_bin_start", "ATAC_bin_end"), 
    sep = "[:-]", 
    remove = FALSE, 
    convert = TRUE 
  )

head(ATAC_parsed)
dim(ATAC_parsed)




#number of unique ATAC bin
print(paste("Total number of unique significant ATAC_bins in final ATAC_parsed:", length(unique(ATAC_parsed$ATAC_bin))))
print(paste("Total number of unique significant ATAC_bins START in final ATAC_parsed:", length(unique(ATAC_parsed$ATAC_bin_start))))
print(paste("Total number of unique significant ATAC_bins START in final ATAC_parsed:", length(unique(ATAC_parsed$ATAC_bin_end))))
print(paste("unique chromosomes in final ATAC_parsed:", unique(ATAC_parsed$chr)))


```


Read in my original sgRNA library from the big targeted cCRE screen to get control sgRNAs


```{r}
original_largescreen_sgrna_library <- read.csv("../data/common_references/sgRNA_library_FINAL_w_sgOPTI_CROP_adapters.csv")
head(original_largescreen_sgrna_library)


#get subsetted dfs for the types of controls I need

# Subset for hCRISPRa_non_targeting_control
hCRISPRa_non_targeting_control <- original_largescreen_sgrna_library %>%
  filter(df_name == "hCRISPRa_non_targeting_control")

# Subset for hCRISPRi_non_targeting_control
hCRISPRi_non_targeting_control <- original_largescreen_sgrna_library %>%
  filter(df_name == "hCRISPRi_non_targeting_control")

# Subset for AAVS1_negative_control
AAVS1_negative_control <- original_largescreen_sgrna_library %>%
  filter(df_name == "AAVS1_negative_control")

unique(hCRISPRa_non_targeting_control$df_name)
dim(hCRISPRa_non_targeting_control)

unique(hCRISPRi_non_targeting_control$df_name)
dim(hCRISPRi_non_targeting_control)

unique(AAVS1_negative_control$df_name)
dim(AAVS1_negative_control)
```
```{r}
print(paste("number of guides allocated to controls to bring to 1000 sgRNAs:", 1000-length(unique(drop_random_controls$sgRNA.sequence))))
print(paste("number of guides allocated to controls - 56 AAVS1 controls:", 358-nrow(AAVS1_negative_control)))
print(paste("number of guides allocated to NTC controls (358 total - 56 AAVS1 controls - 15 local random controls):", 358-nrow(AAVS1_negative_control)-nrow(readcsv_final_randomly_chosen_local_controls_CFD_filtered)))
print(paste("percent of guide library that is all of negative controls combined:", (358/1000)*100))
```

```{r}
print(paste("number of guides per CRISPRi_NTC and CRISPRa_NTC = 144 &143 respectively:", 287==144+143))

# Take rows 1 through 151 from the hCRISPRa dataframe
hCRISPRa_NTC_subset <- hCRISPRa_non_targeting_control[1:143, ]

# Take rows 1 through 151 from the hCRISPRi dataframe
hCRISPRi_NTC_subset <- hCRISPRi_non_targeting_control[1:144, ]

# Check the number of rows in the new subset
nrow(hCRISPRa_NTC_subset)# Expected output: [1] 143
length(unique(hCRISPRa_NTC_subset$sgRNA.sequence))

nrow(hCRISPRi_NTC_subset)# Expected output: [1] 144
length(unique(hCRISPRi_NTC_subset$sgRNA.sequence))
```



#now combine the negative controls with the 642 experimental sgRNA hit pool:

```{r}
head(experimental_sgRNA_pool_642_guides)
dim(experimental_sgRNA_pool_642_guides)

#get rid of window column so  you can collapse distinct guides together
distinct_experimental_sgRNA_pool_642_guides <- experimental_sgRNA_pool_642_guides %>%
  select(ATAC_bin, sgRNA.sequence) %>%
  distinct()

distinct_experimental_sgRNA_pool_642_guides <- 
  distinct_experimental_sgRNA_pool_642_guides %>%
  mutate(df_name = "hits_for_FACS_followup")

head(distinct_experimental_sgRNA_pool_642_guides)
dim(distinct_experimental_sgRNA_pool_642_guides)

```

```{r}
to_combine_hCRISPRa_non_targeting_control <- hCRISPRa_NTC_subset %>% select(sgRNA.sequence, df_name, Gene.Symbol)
head(to_combine_hCRISPRa_non_targeting_control)
nrow(to_combine_hCRISPRa_non_targeting_control)

to_combine_hCRISPRi_non_targeting_control <- hCRISPRi_NTC_subset %>% select(sgRNA.sequence, df_name, Gene.Symbol)
head(to_combine_hCRISPRi_non_targeting_control)
nrow(to_combine_hCRISPRi_non_targeting_control)



to_combine_AAVS1_negative_control <- AAVS1_negative_control %>% select(sgRNA.sequence, df_name, Gene.Symbol)
head(to_combine_AAVS1_negative_control)
nrow(to_combine_AAVS1_negative_control)


to_combine_randomly_chosen_local_controls_CFD_filtered <-
  readcsv_final_randomly_chosen_local_controls_CFD_filtered %>%
  select(ATAC_bin, sgRNA.sequence = sgRNA_per_window) %>%
  mutate(df_name = "local_negative_controls")



head(to_combine_randomly_chosen_local_controls_CFD_filtered)
nrow(to_combine_randomly_chosen_local_controls_CFD_filtered)
```


#combine the 642 hit sgRNAs and the AAVS1, crispri and crispra ntc dataframes - you should get 1000 unique sgRNA.sequence
```{r}

# Combine the four data frames into one
combined_library_large_pool <- bind_rows(
  distinct_experimental_sgRNA_pool_642_guides,
  to_combine_hCRISPRa_non_targeting_control,
  to_combine_hCRISPRi_non_targeting_control,
  to_combine_AAVS1_negative_control,
  to_combine_randomly_chosen_local_controls_CFD_filtered
)

length(unique(combined_library_large_pool$sgRNA.sequence)) #this should be 1000
```


```{r}
final_sgRNA_lib_w_CROP_adapters <- combined_library_large_pool %>%
  mutate(Left_CROP_sgOPTI_adapter = "atcttgtggaaaggacgaaacaccG") %>%
  mutate(Right_CROP_sgOPTI_adapter = "gtttaagagctatgctggaaacagcatagcaagt") %>%
  mutate(sgRNA_w_CROPsgOPTI_adapters = paste0(Left_CROP_sgOPTI_adapter, sgRNA.sequence, Right_CROP_sgOPTI_adapter)) %>%
  select(
    sgRNA_w_CROPsgOPTI_adapters, 
    sgRNA.sequence, 
    Left_CROP_sgOPTI_adapter, 
    Right_CROP_sgOPTI_adapter, 
    everything()
  )
```



save another version for the supplementary table with cleaned up df_name
```{r}
for_supp_table_final_sgRNA_lib_w_CROP_adapters <- final_sgRNA_lib_w_CROP_adapters %>%
  mutate(df_name = ifelse(df_name == "hits_for_FACS_followup", "hits_for_followup", df_name))

write.csv(
  for_supp_table_final_sgRNA_lib_w_CROP_adapters,
  file = "./outputs/TMZ_survival_screen_sgRNA_library/TMZ_survival_screen_pooled_1000_guides_for_publishing.csv",
  row.names = FALSE
)
```


sanity check that all the guides I just made with adapters do exist in my original 8000+ guide screen pool:
```{r}
# Create a logical vector: TRUE if a match is found in the original library, FALSE if not
has_match <- final_sgRNA_lib_w_CROP_adapters$sgRNA_w_CROPsgOPTI_adapters %in% original_largescreen_sgrna_library$sgRNA_w_CROPsgOPTI_adapters

# Use this logical vector to find the rows in your final library that DO NOT have a match
unmatched_guides <- final_sgRNA_lib_w_CROP_adapters[!has_match, ]

# Check if any unmatched guides were found and print the appropriate message
if (nrow(unmatched_guides) == 0) {
  print("Success! All guides in the final library have a match in the original library.")
} else {
  print("Warning: The following guides were not found in the original library:")
  print(unmatched_guides)
}
```


keep just sgRNA column with adaptersfor genscript order spreadsheet
```{r}
# 1. Create a new data frame with only the column you want to keep
guides_to_order <- final_sgRNA_lib_w_CROP_adapters %>%
  select(sgRNA_w_CROPsgOPTI_adapters)

# 2. Define the full output file path
output_file_path <- "./outputs/TMZ_survival_screen_sgRNA_library/TMZ_survival_screen_pooled_1000_sgRNA_lib_genscript_order.csv"

# 3. Save the new data frame to a CSV file
write.csv(guides_to_order, file = output_file_path, row.names = FALSE, quote = FALSE)
```


```{r}
length(unique(guides_to_order$sgRNA_w_CROPsgOPTI_adapters)) #this should be the 1000 sgRNAs
unique(str_length(guides_to_order$sgRNA_w_CROPsgOPTI_adapters)) #with crop homology all guides should be 79bp long
```













