---
title: "BE4max and KRAB Stacked Jitter Plots Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup and Dependencies

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(plotly)
library(scales)
library(htmlwidgets)
```

## Define Paths

```{r paths}
# Input Files (Using filtered 1.5% removal data)
krab_csv_path <- "./outputs_KRAB_TMZ_survival/bottom_1pt5percent_guides_removed_KRAB_plot_df_stouffer_data_final.csv"
be4max_csv_path <- "./outputs_BE4max_TMZ_survival/bottom_1pt5percent_guides_removed_BE4max_plot_df_stouffer_data_final.csv"

# Output Directory
output_dir <- "./outputs_stacked_jitter_plots"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Filename Suffix
file_suffix <- "_bottom_1.5percent_sgRNAs_removed"
```

## Data Loading and Preparation

```{r load_data}
# 1. Load Data
krab_df <- read.csv(krab_csv_path, stringsAsFactors = FALSE)
be4max_df <- read.csv(be4max_csv_path, stringsAsFactors = FALSE)
```

```{r isolate_STON1_hits}
# Define columns to keep
keep_cols <- c(
    "GuideID", "guide_type", "chr", "genomic_start", "genomic_end",
    "Sequence", "sgRNA", "sgRNA.Strand", "X..edits",
    "Nucleotide.edits", "Amino.acid.edits", "Mutation.category", "is_significant"
)

# Filter KRAB STON1 Hits
KRAB_STON1_stouffers_significant_guides <- krab_df %>%
    filter(guide_type == "chr2:48530040-48530316" & is_significant == TRUE) %>%
    # Convert to 1-based coordinates for UCSC and parse chromosome
    mutate(
        genomic_start = as.numeric(genomic_start) + 1,
        genomic_end = as.numeric(genomic_end) + 1,
        chr = sub(":.*", "", guide_type)
    ) %>%
    dplyr::select(all_of(keep_cols)) %>%
    distinct()

# Filter BE4max STON1 Hits
BE4max_STON1_stouffers_significant_guides <- be4max_df %>%
    filter(guide_type == "chr2:48530040-48530316" & is_significant == TRUE) %>%
    # Convert to 1-based coordinates for UCSC and parse chromosome
    mutate(
        genomic_start = as.numeric(genomic_start) + 1,
        genomic_end = as.numeric(genomic_end) + 1,
        chr = sub(":.*", "", guide_type)
    ) %>%
    dplyr::select(all_of(keep_cols)) %>%
    distinct()

# Optional: Print counts to verify
print(paste("KRAB STON1 Significant Guides (Unique):", nrow(KRAB_STON1_stouffers_significant_guides)))
print(paste("BE4max STON1 Significant Guides (Unique):", nrow(BE4max_STON1_stouffers_significant_guides)))

# Save dataframes as CSV files
csv_out_dir <- "./outputs_stacked_jitter_plots/supplementary_tables_STON1_hit_guides"
if (!dir.exists(csv_out_dir)) dir.create(csv_out_dir, recursive = TRUE)

write.csv(KRAB_STON1_stouffers_significant_guides, file.path(csv_out_dir, "KRAB_STON1_stouffers_significant_guides.csv"), row.names = FALSE)
write.csv(BE4max_STON1_stouffers_significant_guides, file.path(csv_out_dir, "BE4max_STON1_stouffers_significant_guides.csv"), row.names = FALSE)

# ==========================================
# Export as BED Files for UCSC/IGV/GVIZ plotting
# ==========================================

# 1. Define BED output directory
bed_out_dir <- "./outputs_stacked_jitter_plots/final_bed_files_stouffers_significant_sgRNA_hits_for_IGV_type_plots"
if (!dir.exists(bed_out_dir)) dir.create(bed_out_dir, recursive = TRUE)

# 2. Function to save as BED (no header, tab-separated)
# Order: chrom, chromStart, chromEnd, name
save_as_bed <- function(df, filename) {
    bed_data <- df %>%
        dplyr::select(chr, genomic_start, genomic_end, Sequence)

    write.table(bed_data,
        file = file.path(bed_out_dir, filename),
        quote = FALSE,
        sep = "\t",
        row.names = FALSE,
        col.names = FALSE
    )
}

# 3. Save the files
if (nrow(KRAB_STON1_stouffers_significant_guides) > 0) {
    save_as_bed(KRAB_STON1_stouffers_significant_guides, paste0("KRAB_STON1_significant_hits", file_suffix, ".bed"))
}
if (nrow(BE4max_STON1_stouffers_significant_guides) > 0) {
    save_as_bed(BE4max_STON1_stouffers_significant_guides, paste0("BE4max_STON1_significant_hits", file_suffix, ".bed"))
}

print(paste("BED files saved to:", bed_out_dir))
```

```{r thresholds}
# 2. Load Pre-calculated Thresholds (From filtered analysis folders)
krab_thresh_path <- "./outputs_KRAB_TMZ_survival/combined_threshold.csv"
be4max_thresh_path <- "./outputs_BE4max_TMZ_survival/combined_threshold.csv"

krab_threshold <- read.csv(krab_thresh_path)$combined_threshold
be4max_threshold <- read.csv(be4max_thresh_path)$combined_threshold

print(paste("KRAB Threshold (Loaded):", round(krab_threshold, 4)))
print(paste("BE4max Threshold (Loaded):", round(be4max_threshold, 4)))

# 3. Load Manual Coloring from CSV
# We now use pre-assigned colors for global consistency
manual_colors_path <- "../data/ATAC_bin_color_map/manual_colors_ATAC_bins.csv"
manual_colors_df <- read.csv(manual_colors_path, stringsAsFactors = FALSE)

# Filter for rows that have a color assigned and create named vector
# Using guide_type as names and final_colors as values
default_colors <- manual_colors_df$final_colors
names(default_colors) <- manual_colors_df$guide_type

# Override controls to black
static_controls <- c("AAVS1_negative_control", "hCRISPRa_non_targeting_control", "hCRISPRi_non_targeting_control")

# Dynamically find guide_types associated with local controls (by checking GuideID)
# Filter for local controls in both DFs
local_control_rows_krab <- krab_df[grepl("local_negative_controls", krab_df$GuideID), ]
local_control_rows_be4max <- be4max_df[grepl("local_negative_controls", be4max_df$GuideID), ]

# Extract their guide types
local_control_types <- unique(c(local_control_rows_krab$guide_type, local_control_rows_be4max$guide_type))

# Combine all controls
all_control_types <- unique(c(static_controls, local_control_types))

# Assign "black" to all identified control types
for (ctl in all_control_types) {
    if (ctl %in% names(default_colors)) {
        default_colors[ctl] <- "black"
    }
}

# Add "Not Significant" color
stouffer_colors <- default_colors
stouffer_colors["Not Significant"] <- "lightgrey"

# 4. Helper Function to Process Dataframe (Sort + Legend Column + Capping)
process_df <- function(df, dataset_name) {
    # Ensure sorting columns are numeric
    df$genomic_start <- as.numeric(df$genomic_start)
    df$ATAC_bin_start <- as.numeric(df$ATAC_bin_start)

    # Re-apply Significance Column just in case
    if (!"Legend_Label_Stouffer" %in% colnames(df)) {
        df <- df %>%
            mutate(
                is_stouffer_sig = !is.na(q_value) & q_value < 0.01,
                Legend_Label_Stouffer = ifelse(is_stouffer_sig, guide_type, "Not Significant")
            )
    }

    # ---------------------------------------------------------
    # Outlier Capping Logic
    # ---------------------------------------------------------
    # Identify outliers
    outliers <- df %>%
        filter(Scaled_Log2FC < -2.35 | Scaled_Log2FC > 2.35)

    if (nrow(outliers) > 0) {
        print(paste("--- Capped Outliers in", dataset_name, "---"))
        for (i in 1:nrow(outliers)) {
            print(paste(
                "GuideID:", outliers$GuideID[i],
                "| Original Scaled_Log2FC:", round(outliers$Scaled_Log2FC[i], 4)
            ))
        }
    }

    # Apply Capping, Create Plotted_LFC, and Define Shape
    df <- df %>%
        mutate(
            Is_Outlier = Scaled_Log2FC < -2.35 | Scaled_Log2FC > 2.35,
            Plotted_LFC = case_when(
                Scaled_Log2FC < -2.35 ~ -2.33, # Cap bottom
                Scaled_Log2FC > 2.35 ~ 2.33, # Cap top
                TRUE ~ Scaled_Log2FC
            ),
            # Shape Factor: "Normal" vs "Capped"
            Point_Shape = ifelse(Is_Outlier, "Capped", dataset_name)
        )

    # Apply Sorting
    df <- df %>%
        mutate(
            primary_sort = ifelse(guide_type == "AAVS1_negative_control", 0, 1)
        ) %>%
        arrange(
            primary_sort,
            ATAC_bin_start,
            genomic_start,
            GuideID
        ) %>%
        mutate(GuideID = factor(GuideID, levels = unique(GuideID)))

    return(df)
}

krab_plot_df <- process_df(krab_df, "KRAB")
be4max_plot_df <- process_df(be4max_df, "BE4max")

# Combine Dataframes for Faceting
krab_plot_df$Dataset <- "KRAB"
be4max_plot_df$Dataset <- "BE4max"

combined_plot_df <- rbind(krab_plot_df, be4max_plot_df)

# Set Factor Levels to ensure KRAB is on Top
combined_plot_df$Dataset <- factor(combined_plot_df$Dataset, levels = c("KRAB", "BE4max"))

# Define Threshold Dataframe for Faceted lines
threshold_df <- data.frame(
    Dataset = factor(c("KRAB", "BE4max"), levels = c("KRAB", "BE4max")),
    Threshold = c(krab_threshold, be4max_threshold)
)
```

## Generate Combined Faceted Plot

```{r combined_plot, fig.height=14, fig.width=12}
p_combined <- ggplot(combined_plot_df, aes(
    x = GuideID, y = Plotted_LFC, color = Legend_Label_Stouffer, shape = Point_Shape,
    text = paste(
        "Dataset:", Dataset,
        "<br>GuideID:", GuideID,
        "<br>Guide Type:", guide_type,
        "<br>Genomic Start:", genomic_start,
        "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
        "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
        "<br>q-value:", formatC(q_value, format = "e", digits = 2),
        "<br>Capped:", Is_Outlier
    )
)) +
    geom_jitter(width = 0.2, size = 1.5, alpha = 0.7) +

    # Facet Specification: Dataset on rows (Vertical Stack), label on Right side
    facet_grid(Dataset ~ .) +
    scale_color_manual(values = stouffer_colors) +
    # Map shapes: KRAB = 16 (circle), BE4max = 17 (triangle), Capped = 15 (square)
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        # Remove the weird grey box (strip background)
        strip.background = element_blank(),
        # Rotate facet labels 90 degrees (text runs vertical)
        strip.text.y = element_text(size = 12, face = "bold", angle = -90)
    ) +
    labs(
        title = "Stacked Proportion Relative to Average DMSO (Stouffer's FDR < 0.01)",
        x = "GuideID",
        y = "Scaled Log2FC"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +

    # Add Threshold Lines per Facet
    geom_hline(data = threshold_df, aes(yintercept = Threshold), linetype = "dotted", color = "darkgrey") +
    geom_hline(data = threshold_df, aes(yintercept = -Threshold), linetype = "dotted", color = "darkgrey")

plotly_combined <- ggplotly(p_combined, tooltip = "text")

# Save with verbose debugging
print(paste("Checking output directory:", output_dir))
if (!dir.exists(output_dir)) {
    print("Directory does not exist. Creating...")
    dir.create(output_dir, recursive = TRUE)
}

html_filename <- file.path(output_dir, paste0("KRAB_BE4max_Stacked_Faceted_Plot", file_suffix, ".html"))
print(paste("Attempting to save plot to:", html_filename))

tryCatch(
    {
        saveWidget(plotly_combined, html_filename, selfcontained = TRUE)
        if (file.exists(html_filename)) {
            print(paste("SUCCESS: File created at", html_filename))
            print(paste("File size:", file.info(html_filename)$size, "bytes"))
        } else {
            print("WARNING: saveWidget ran without error, but file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving plot:", e$message))
    }
)
```

## Generate Overlaid Jitter Plot

```{r overlaid_plot, fig.height=8, fig.width=12}
# Prepare Data for Overlaid Plot
overlaid_df <- combined_plot_df %>%
    mutate(
        # Shape Logic: Capped (Circle=16) > KRAB (Star=8) > BE4max (Square=15)
        Overlaid_Shape = case_when(
            Is_Outlier ~ "Capped",
            Dataset == "KRAB" ~ "KRAB",
            Dataset == "BE4max" ~ "BE4max"
        ),
        # Size Logic: Significant (Not Sig/Control color check) -> Large (2.5), Else -> Small (1)
        # Using Legend_Label_Stouffer to determine significance (Grey = Not Sig)
        Overlaid_Size = ifelse(Legend_Label_Stouffer == "Not Significant", "Small", "Large"),
        # Manual X Calculation for consistent dodging + ordering
        # Convert factor to integer, then offset KRAB left, BE4max right
        x_manual = as.integer(GuideID) + ifelse(Dataset == "KRAB", -0.25, 0.25)
    ) %>%
    # Sorting: Ensure "Small" (Grey) points are plotted FIRST, "Large" (Sig) LAST (on top)
    arrange(factor(Overlaid_Size, levels = c("Small", "Large")))

# Split into Background and Foreground for Strict Z-ordering
df_bg <- overlaid_df %>% filter(Overlaid_Size == "Small")
df_fg <- overlaid_df %>% filter(Overlaid_Size == "Large")

p_overlaid <- ggplot() +
    # Layer 1: Background (Non-Significant / Small)
    geom_jitter(
        data = df_bg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Legend_Label_Stouffer,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Dataset:", Dataset,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2),
                "<br>Capped:", Is_Outlier
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground (Significant / Large) - Plotted ON TOP
    geom_jitter(
        data = df_fg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Legend_Label_Stouffer,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Dataset:", Dataset,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2),
                "<br>Capped:", Is_Outlier
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +
    scale_color_manual(values = stouffer_colors) +

    # Shape Logic: KRAB=16 (Circle), BE4max=15 (Square), Capped=17 (Triangle)
    # Shape Logic: KRAB=16 (Circle), BE4max=17 (Triangle), Capped=15 (Square)
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +

    # Size Logic: Large=2.5, Small=1.0
    scale_size_manual(values = c("Large" = 2.5, "Small" = 1.0), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    ) +
    labs(
        title = "Overlaid Proportion Relative to Average DMSO (Stouffer's FDR < 0.05)",
        x = "GuideID",
        y = "Scaled Log2FC"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")


# Save Dataframe used for plotting
plot_data_csv <- file.path(output_dir, paste0("Proportion_Rel_Avg_DMSO_Data", file_suffix, ".csv"))
write.csv(combined_plot_df, plot_data_csv, row.names = FALSE)
print(paste("Saved plotting data to:", plot_data_csv))

plotly_overlaid <- ggplotly(p_overlaid, tooltip = "text")

# Save Overlaid Plot
html_filename_overlaid <- file.path(output_dir, paste0("KRAB_BE4max_Overlaid_Stouffers_Plot", file_suffix, ".html"))
print(paste("Attempting to save overlaid plot to:", html_filename_overlaid))

tryCatch(
    {
        saveWidget(plotly_overlaid, html_filename_overlaid, selfcontained = TRUE)
        if (file.exists(html_filename_overlaid)) {
            print(paste("SUCCESS: Overlaid File created at", html_filename_overlaid))
        } else {
            print("WARNING: saveWidget ran without error, but overlaid file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving overlaid plot:", e$message))
    }
)
```

## Generate Simplified Overlaid Plot (Black/Blue)

```{r simple_overlaid_plot, fig.height=8, fig.width=12}
# Prepare Data for Simplified Coloring
# Logic:
# - Not Significant -> "Not Significant" (Light Grey)
# - Significant + KRAB -> "KRAB Significant" (Black)
# - Significant + BE4max -> "BE4max Significant" (Light Blue)
simple_df <- overlaid_df %>%
    mutate(
        Legend_Label_Simple = case_when(
            Legend_Label_Stouffer == "Not Significant" ~ "Not Significant",
            Dataset == "KRAB" ~ "KRAB Significant",
            Dataset == "BE4max" ~ "BE4max Significant"
        )
    )

# Split for Z-ordering (same as before)
# Split for Z-ordering (same as before)
# Split for Z-ordering: Background, Foreground KRAB, Foreground BE4max
df_bg_simple <- simple_df %>% filter(Overlaid_Size == "Small")
df_fg_simple_krab <- simple_df %>% filter(Overlaid_Size == "Large" & Dataset == "KRAB")
df_fg_simple_be4max <- simple_df %>% filter(Overlaid_Size == "Large" & Dataset == "BE4max")

# Define Simple Colors
simple_colors <- c(
    "Not Significant" = "lightgrey",
    "KRAB Significant" = "black",
    "BE4max Significant" = "#1ABA22" # Custom Green 2
)

p_simple <- ggplot() +
    # Layer 1: Background (Non-Significant / Small)
    geom_jitter(
        data = df_bg_simple,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Legend_Label_Simple,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Dataset:", Dataset,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Significance:", Legend_Label_Simple,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2),
                "<br>Capped:", Is_Outlier
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground KRAB (Significant / Large) - Plotted FIRST
    geom_jitter(
        data = df_fg_simple_krab,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Legend_Label_Simple,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Dataset:", Dataset,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Significance:", Legend_Label_Simple,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2),
                "<br>Capped:", Is_Outlier
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 3: Foreground BE4max (Significant / Large) - Plotted ON TOP
    geom_jitter(
        data = df_fg_simple_be4max,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Legend_Label_Simple,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Dataset:", Dataset,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Significance:", Legend_Label_Simple,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>Plotted Log2FC:", round(Plotted_LFC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2),
                "<br>Capped:", Is_Outlier
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +
    scale_color_manual(values = simple_colors) +

    # Shape Logic: KRAB=16 (Circle), BE4max=15 (Square), Capped=17 (Triangle)
    # Shape Logic: KRAB=16 (Circle), BE4max=17 (Triangle), Capped=15 (Square)
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +

    # Size Logic: Large=2.5, Small=1.0
    scale_size_manual(values = c("Large" = 2.5, "Small" = 1.0), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    ) +
    labs(
        title = "Overlaid Proportion (KRAB=Black, BE4max=Blue, FDR < 0.01)",
        x = "GuideID",
        y = "Scaled Log2FC",
        color = "Significance"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")

plotly_simple <- ggplotly(p_simple, tooltip = "text")

# Save Simplified Plot
html_filename_simple <- file.path(output_dir, paste0("KRAB_BE4max_Overlaid_Simple_Colors_Plot", file_suffix, ".html"))
print(paste("Attempting to save simplified plot to:", html_filename_simple))

tryCatch(
    {
        saveWidget(plotly_simple, html_filename_simple, selfcontained = TRUE)
        if (file.exists(html_filename_simple)) {
            print(paste("SUCCESS: Simplified Plot created at", html_filename_simple))
        } else {
            print("WARNING: saveWidget ran without error, but simple file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving simplified plot:", e$message))
    }
)
```

## Generate 4-Panel Vertically Aligned Combined Plot

```{r four_panel_plot, fig.height=24, fig.width=12}
# ==============================================================================
# Goal: 4 Vertically Stacked Plots with PERTLY ALIGNED X-AXIS
# 1. KRAB (Rainbow Color) - but sized/shaped by sig
# 2. KRAB (Sig Color)
# 3. BE4max (Sig Color)
# 4. Overlaid (Sig Color)
# ==============================================================================

# 1. Prepare Base Data Sources
# We can reuse 'overlaid_df' because it already calculates:
# - x_manual (The critical alignment key)
# - Overlaid_Size (Significance sizing)
# - Overlaid_Shape (Dataset/Capped shapes)
# - Legend_Label_Stouffer (Sig vs Grey coloring)

# --- Panel 1: KRAB Rainbow ---
# Filter for KRAB and force Color to be 'guide_type' (Original Rainbow)
# BUT keep Size/Shape/X derived from significance logic
df_p1 <- overlaid_df %>%
    filter(Dataset == "KRAB") %>%
    mutate(
        Panel = "1. KRAB (All Data)",
        # Use guide_type for color (Rainbow), NOT the Stouffer label
        Plot_Color = guide_type
    )

# --- Panel 2: KRAB Significance ---
# Filter for KRAB, use Stouffer coloring
df_p2 <- overlaid_df %>%
    filter(Dataset == "KRAB") %>%
    mutate(
        Panel = "2. KRAB (Significant)",
        Plot_Color = Legend_Label_Stouffer
    )

# --- Panel 3: BE4max Significance ---
# Filter for BE4max, use Stouffer coloring
df_p3 <- overlaid_df %>%
    filter(Dataset == "BE4max") %>%
    mutate(
        Panel = "3. BE4max (Significant)",
        Plot_Color = Legend_Label_Stouffer
    )

# --- Panel 4: Overlaid Significance ---
# Use ALL data, use Stouffer coloring
df_p4 <- overlaid_df %>%
    mutate(
        Panel = "4. Combined Overlaid",
        Plot_Color = Legend_Label_Stouffer
    )

# 2. Combine into One Master DF
four_panel_df <- bind_rows(df_p1, df_p2, df_p3, df_p4)

# 3. Enforce Factor Ordering for Vertical Stack
# Top (1) -> Bottom (4)
four_panel_df$Panel <- factor(four_panel_df$Panel, levels = c(
    "1. KRAB (All Data)",
    "2. KRAB (Significant)",
    "3. BE4max (Significant)",
    "4. Combined Overlaid"
))

# 4. Z-Ordering (Ensure Large/Sig points are ON TOP across all panels)
four_panel_df <- four_panel_df %>%
    arrange(Panel, factor(Overlaid_Size, levels = c("Small", "Large")))

# Split for plotting
df_4p_bg <- four_panel_df %>% filter(Overlaid_Size == "Small")
df_4p_fg <- four_panel_df %>% filter(Overlaid_Size == "Large")

# 5. Define Unified Color Palette
# We need a palette that handles both:
# - The "Stouffer" keys ("Not Significant", + specific significant types...)
# - AND the "Rainbow" keys (All raw guide_types)
# 'default_colors' already has all guide_types.
# 'stouffer_colors' adds "Not Significant".
# So 'stouffer_colors' is safe to use for EVERYTHING (since it's a superset).
# (Panel 1 just won't use the "Not Significant" key).

p_4panel <- ggplot() +
    # Layer 1: Background
    geom_jitter(
        data = df_4p_bg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground (On Top)
    geom_jitter(
        data = df_4p_fg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Facet Specification: 1 Column, Vertical Stack
    facet_grid(Panel ~ .) +
    scale_color_manual(values = stouffer_colors) +
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
    scale_size_manual(values = c("Large" = 2.5, "Small" = 1.0), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text.y = element_text(size = 10, face = "bold", angle = 0, hjust = 0), # Horizontal labels for readability? Or vertical?
        # User requested vertically stacked. Horizontal text on right side often easier to read for long titles.
        # Let's keep rotated -90 or 0 depending on pref. 0 is easier to read.
        legend.position = "right"
    ) +
    # Override strip text to -90 if standard
    theme(strip.text.y = element_text(angle = -90)) +
    labs(
        title = "4-Panel Aligned Proportion (Stouffer's FDR < 0.01)",
        x = "GuideID",
        y = "Scaled Log2FC",
        color = "Guide Type / Sig"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")

plotly_4panel <- ggplotly(p_4panel, tooltip = "text")

# Save 4-Panel Plot
html_filename_4panel <- file.path(output_dir, paste0("KRAB_BE4max_4_Panel_Aligned_Plot", file_suffix, ".html"))
print(paste("Attempting to save 4-panel plot to:", html_filename_4panel))

tryCatch(
    {
        saveWidget(plotly_4panel, html_filename_4panel, selfcontained = TRUE)
        if (file.exists(html_filename_4panel)) {
            print(paste("SUCCESS: 4-Panel File created at", html_filename_4panel))
        } else {
            print("WARNING: saveWidget ran without error, but 4-panel file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving 4-panel plot:", e$message))
    }
)
```

## Generate 4-Panel Vertically Aligned Plot (Smaller Points Variation)

```{r four_panel_plot_small, fig.height=24, fig.width=12}
# ==============================================================================
# Goal: Same as above but with smaller Significant points (1.75 vs 2.5)
# ==============================================================================

# Use same dataframe 'four_panel_df'

p_4panel_small <- ggplot() +
    # Layer 1: Background
    geom_jitter(
        data = df_4p_bg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground (On Top)
    geom_jitter(
        data = df_4p_fg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Facet Specification: 1 Column, Vertical Stack
    facet_grid(Panel ~ .) +
    scale_color_manual(values = stouffer_colors) +
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
    # VARIATION: Smaller Significant Points (1.75 instead of 2.5)
    scale_size_manual(values = c("Large" = 1.75, "Small" = 1.0), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text.y = element_text(size = 10, face = "bold", angle = -90),
        legend.position = "right"
    ) +
    labs(
        title = "4-Panel Aligned Proportion (Stouffer's FDR < 0.01) [Small Points]",
        x = "GuideID",
        y = "Scaled Log2FC",
        color = "Guide Type / Sig"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")

plotly_4panel_small <- ggplotly(p_4panel_small, tooltip = "text")

# Save 4-Panel Plot (Small Variation)
html_filename_4panel_small <- file.path(output_dir, paste0("KRAB_BE4max_4_Panel_Aligned_Plot_SmallPoints", file_suffix, ".html"))
print(paste("Attempting to save small points plot to:", html_filename_4panel_small))

tryCatch(
    {
        saveWidget(plotly_4panel_small, html_filename_4panel_small, selfcontained = TRUE)
        if (file.exists(html_filename_4panel_small)) {
            print(paste("SUCCESS: Small Points File created at", html_filename_4panel_small))
        } else {
            print("WARNING: saveWidget ran without error, but small points file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving small points plot:", e$message))
    }
)
```

## Generate 4-Panel Vertically Aligned Plot (Simplified Colors + Small Points)

```{r four_panel_plot_simple_small, fig.height=24, fig.width=12}
# ==============================================================================
# Goal: 4 Panels, but Panel 4 uses "Simplified Tuning" (Black/Blue)
# AND uses Small Points (1.75)
# ==============================================================================

# 1. Modify Panel 4 Data to use Simplified Coloring Keys
df_p4_simple <- df_p4 %>%
    mutate(
        # Override Plot_Color with Simple Keys
        Plot_Color = case_when(
            Legend_Label_Stouffer == "Not Significant" ~ "Not Significant",
            Dataset == "KRAB" ~ "KRAB Significant",
            Dataset == "BE4max" ~ "BE4max Significant"
        )
    )

# 2. Re-combine Data
# Panels 1-3 remain exactly the same as 'four_panel_df'
# We only swap out the Panel 4 rows
four_panel_df_simple <- bind_rows(df_p1, df_p2, df_p3, df_p4_simple)

# 3. Enforce Factor Ordering
four_panel_df_simple$Panel <- factor(four_panel_df_simple$Panel, levels = c(
    "1. KRAB (All Data)",
    "2. KRAB (Significant)",
    "3. BE4max (Significant)",
    "4. Combined Overlaid"
))

# 4. Z-Order Sort
# Sort by Size (Small -> Large) THEN by Dataset (KRAB -> BE4max)
# This ensures Large BE4max points are the very top layer
four_panel_df_simple <- four_panel_df_simple %>%
    mutate(Dataset = factor(Dataset, levels = c("KRAB", "BE4max"))) %>%
    arrange(Panel, factor(Overlaid_Size, levels = c("Small", "Large")), Dataset)

# Split
df_4p_simple_bg <- four_panel_df_simple %>% filter(Overlaid_Size == "Small")
df_4p_simple_fg_krab <- four_panel_df_simple %>% filter(Overlaid_Size == "Large" & Dataset == "KRAB")
df_4p_simple_fg_be4max <- four_panel_df_simple %>% filter(Overlaid_Size == "Large" & Dataset == "BE4max")

# 5. Define Extended Simple Palette
# We need a palette that covers:
# - Original Stouffer keys (Panel 2, 3)
# - Rainbow keys (Panel 1)
# - NEW Simple keys (Panel 4)

# Start with our robust stouffer_colors (contains Rainbow + Not Sig + Sig Types)
extended_colors <- stouffer_colors

# Add the specific simple keys if they don't map naturally
extended_colors["KRAB Significant"] <- "black"
extended_colors["BE4max Significant"] <- "#1ABA22"

# Note: "Not Significant" is already in stouffer_colors ("lightgrey")

p_4panel_simple <- ggplot() +
    # Layer 1: Background
    geom_jitter(
        data = df_4p_simple_bg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground KRAB (On Top) - Plotted FIRST
    geom_jitter(
        data = df_4p_simple_fg_krab,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 3: Foreground BE4max (On Top of KRAB) - Plotted LAST
    geom_jitter(
        data = df_4p_simple_fg_be4max,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size,
            text = paste(
                "Panel:", Panel,
                "<br>GuideID:", GuideID,
                "<br>Guide Type:", guide_type,
                "<br>Color Group:", Plot_Color,
                "<br>Original Scaled Log2FC:", round(Scaled_Log2FC, 2),
                "<br>q-value:", formatC(q_value, format = "e", digits = 2)
            )
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Facet Stack
    facet_grid(Panel ~ .) +
    scale_color_manual(values = extended_colors) +
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
    # SMALL POINTS VARIATION (1.75)
    scale_size_manual(values = c("Large" = 1.75, "Small" = 1.0), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    theme_classic() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        strip.text.y = element_text(size = 10, face = "bold", angle = -90),
        legend.position = "right"
    ) +
    labs(
        title = "4-Panel Aligned (Simple Colors on Bottom) [Small Points]",
        x = "GuideID",
        y = "Scaled Log2FC",
        color = "Guide Type / Sig"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")

plotly_4panel_simple <- ggplotly(p_4panel_simple, tooltip = "text")

# Save 4-Panel Plot (Simple + Small)
html_filename_simple_small <- file.path(output_dir, paste0("KRAB_BE4max_4_Panel_Aligned_Plot_Simple_SmallPoints", file_suffix, ".html"))
print(paste("Attempting to save simple small points plot to:", html_filename_simple_small))

tryCatch(
    {
        saveWidget(plotly_4panel_simple, html_filename_simple_small, selfcontained = TRUE)
        if (file.exists(html_filename_simple_small)) {
            print(paste("SUCCESS: Simple Small Points File created at", html_filename_simple_small))
        } else {
            print("WARNING: saveWidget ran without error, but file was NOT found.")
        }
    },
    error = function(e) {
        print(paste("ERROR saving simple small points plot:", e$message))
    }
)
```

## Generate Static 4-Panel Plot for Paper (PDF/PNG, No Legend)

```{r static_paper_plot, fig.height=6, fig.width=10}
# ==============================================================================
# Goal: Static Version of "Simple Colors + Small Points" for Paper
# - No Legend
# - No Interaction (text aesthetic removed)
# - Saved as PDF and PNG
# ==============================================================================

print("Generating Static Paper Plot...")

# Reuse the dataframes and colors from the previous chunk:
# - df_4p_simple_bg
# - df_4p_simple_fg_krab
# - df_4p_simple_fg_be4max
# - extended_colors

p_static <- ggplot() +
    # Layer 1: Background
    geom_jitter(
        data = df_4p_simple_bg,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 2: Foreground KRAB (On Top) - Plotted FIRST
    geom_jitter(
        data = df_4p_simple_fg_krab,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Layer 3: Foreground BE4max (On Top of KRAB) - Plotted LAST
    geom_jitter(
        data = df_4p_simple_fg_be4max,
        aes(
            x = x_manual, y = Plotted_LFC,
            color = Plot_Color,
            shape = Overlaid_Shape,
            size = Overlaid_Size
        ),
        width = 0.15, height = 0, alpha = 0.7
    ) +

    # Facet Specification: 1 Column, Vertical Stack
    facet_grid(Panel ~ .) +
    scale_color_manual(values = extended_colors) +
    scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
    # Adjusted Sizes: Large=2.5, Small=1.5
    scale_size_manual(values = c("Large" = 2.5, "Small" = 1.5), guide = "none") +
    scale_y_continuous(limits = c(-2.35, 2.35)) +
    # Reduce white space on sides (Tight X-Axis)
    scale_x_continuous(expand = c(0.005, 0.005)) +
    theme_classic(base_size = 7) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        # Force every element to be exactly 7pt
        plot.title = element_text(size = 7, face = "bold"),
        axis.title = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        strip.text.y = element_text(size = 7, face = "bold", angle = -90),
        # NO LEGEND FOR PAPER
        legend.position = "none"
    ) +
    labs(
        title = "4-Panel Aligned Proportion (Static Paper Version)",
        x = "GuideID",
        y = "Scaled Log2FC"
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey")

# Save as PDF
pdf_filename <- file.path(output_dir, paste0("KRAB_BE4max_4_Panel_Static_Paper", file_suffix, ".pdf"))
print(paste("Saving PDF to:", pdf_filename))
ggsave(pdf_filename, plot = p_static, width = 6.3, height = 4.5, limitsize = FALSE)

# Save as PNG
png_filename <- file.path(output_dir, paste0("KRAB_BE4max_4_Panel_Static_Paper", file_suffix, ".png"))
print(paste("Saving PNG to:", png_filename))
ggsave(png_filename, plot = p_static, width = 6.3, height = 4.5, dpi = 300, limitsize = FALSE)

print("Static plots saved successfully.")
```

## Generate Split Static Plots (AAVS1-only vs Main Figure)

```{r split_static_plots}
# ==============================================================================
# Goal: Split p_static into two separate files with same Y-axis and format
# 1. AAVS1 Only (Supplemental)
# 2. Everything except AAVS1 (Main Fig)
# ==============================================================================

# 1. Define New Output Directories
main_fig_dir <- "./outputs_stacked_jitter_plots/figures_for_paper"
supp_fig_dir <- file.path(main_fig_dir, "Supplemental")

if (!dir.exists(main_fig_dir)) dir.create(main_fig_dir, recursive = TRUE)
if (!dir.exists(supp_fig_dir)) dir.create(supp_fig_dir, recursive = TRUE)

# 2. Universal Plotting Dimensions and Font
static_w <- 6.3
static_h <- 4.5
static_base_size <- 7

# --- HELPER FUNCTION TO GENERATE PLOT ---
# This ensures both plots have identical formatting and Y-axis
generate_static_subset_plot <- function(data_bg, data_fg_krab, data_fg_be4max, title_text) {
    ggplot() +
        geom_jitter(
            data = data_bg, width = 0.15, height = 0, alpha = 0.7,
            aes(x = x_manual, y = Plotted_LFC, color = Plot_Color, shape = Overlaid_Shape, size = Overlaid_Size)
        ) +
        geom_jitter(
            data = data_fg_krab, width = 0.15, height = 0, alpha = 0.7,
            aes(x = x_manual, y = Plotted_LFC, color = Plot_Color, shape = Overlaid_Shape, size = Overlaid_Size)
        ) +
        geom_jitter(
            data = data_fg_be4max, width = 0.15, height = 0, alpha = 0.7,
            aes(x = x_manual, y = Plotted_LFC, color = Plot_Color, shape = Overlaid_Shape, size = Overlaid_Size)
        ) +
        facet_grid(Panel ~ .) +
        scale_color_manual(values = extended_colors) +
        scale_shape_manual(values = c("KRAB" = 16, "BE4max" = 17, "Capped" = 15), guide = "none") +
        scale_size_manual(values = c("Large" = 2.5, "Small" = 1.5), guide = "none") +
        scale_y_continuous(limits = c(-2.35, 2.35)) +
        scale_x_continuous(expand = c(0.005, 0.005)) +
        theme_classic(base_size = static_base_size) +
        theme(
            axis.text.x = element_blank(), axis.ticks.x = element_blank(),
            strip.background = element_blank(),
            plot.title = element_text(size = static_base_size, face = "bold"),
            axis.title = element_text(size = static_base_size),
            axis.text.y = element_text(size = static_base_size),
            strip.text.y = element_text(size = static_base_size, face = "bold", angle = -90),
            legend.position = "none"
        ) +
        labs(title = title_text, x = "GuideID", y = "Scaled Log2FC") +
        geom_hline(yintercept = 0, linetype = "dashed", color = "grey")
}

# 3. GENERATE AND SAVE PLOT 1: AAVS1 ONLY
df_bg_aavs1 <- df_4p_simple_bg %>% filter(guide_type == "AAVS1_negative_control")
df_krab_aavs1 <- df_4p_simple_fg_krab %>% filter(guide_type == "AAVS1_negative_control")
df_be4max_aavs1 <- df_4p_simple_fg_be4max %>% filter(guide_type == "AAVS1_negative_control")

p_aavs1_only <- generate_static_subset_plot(df_bg_aavs1, df_krab_aavs1, df_be4max_aavs1, "AAVS1 Controls Only")

aavs1_pdf <- file.path(supp_fig_dir, "AAVS1_only_stacked_jitter_bottom_1.5_guides_removed.pdf")
aavs1_png <- file.path(supp_fig_dir, "AAVS1_only_stacked_jitter_bottom_1.5_guides_removed.png")

ggsave(aavs1_pdf, plot = p_aavs1_only, width = 1.1, height = static_h)
ggsave(aavs1_png, plot = p_aavs1_only, width = 1.1, height = static_h, dpi = 300)

# 4. GENERATE AND SAVE PLOT 2: ALL EXCEPT AAVS1
df_bg_main <- df_4p_simple_bg %>% filter(guide_type != "AAVS1_negative_control")
df_krab_main <- df_4p_simple_fg_krab %>% filter(guide_type != "AAVS1_negative_control")
df_be4max_main <- df_4p_simple_fg_be4max %>% filter(guide_type != "AAVS1_negative_control")

p_main_fig <- generate_static_subset_plot(df_bg_main, df_krab_main, df_be4max_main, "Main Figure (AAVS1 Removed)")

main_pdf <- file.path(main_fig_dir, "final_main_fig_stacked_jitter_AAVS1_removed_bottom_1.5_guides_removed.pdf")
main_png <- file.path(main_fig_dir, "final_main_fig_stacked_jitter_AAVS1_removed_bottom_1.5_guides_removed.png")

ggsave(main_pdf, plot = p_main_fig, width = static_w, height = static_h)
ggsave(main_png, plot = p_main_fig, width = static_w, height = static_h, dpi = 300)

# 5. SAVE SECOND VERSION OF MAIN FIG (Supplemental, 5-inch width)
supp_main_pdf <- file.path(supp_fig_dir, "AAVS1_removed_stacked_jitter_bottom_1.5_guides_removed.pdf")
supp_main_png <- file.path(supp_fig_dir, "AAVS1_removed_stacked_jitter_bottom_1.5_guides_removed.png")

ggsave(supp_main_pdf, plot = p_main_fig, width = 5.4, height = static_h)
ggsave(supp_main_png, plot = p_main_fig, width = 5.4, height = static_h, dpi = 300)

print("Split static plots generated and saved successfully.")
```
