---
title: "Broad Peaks Final Analysis (Refactored)"
output: html_document
date: "2024-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Setup and Data Import
```{r}
library(dplyr)
library(tidyr)
library(DESeq2)
library(pheatmap)

# --- Define Output Path ---
# Using relative path for GitHub repository
output_dir <- "./outputs/ATAC_analysis"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# --- Define Subfolders for Paper Figures/Tables ---
main_fig_dir <- file.path(output_dir, "final_main_paper_figures")
if (!dir.exists(main_fig_dir)) dir.create(main_fig_dir, recursive = TRUE)

supp_dir <- file.path(output_dir, "final_supplemental_figs_tables_for_paper")
if (!dir.exists(supp_dir)) dir.create(supp_dir, recursive = TRUE)

# --- 1. Load Raw Counts (Skip header line 1) ---
featureCounts_raw <- read.delim(gzfile("../data/A172_GBM_ATAC_seq/GSM9825200_consensus_peaks.mLb.clN.featureCounts.txt.gz"),
  header = TRUE, sep = "\t", skip = 1
)

# --- 2. Filter Bad Samples (Programmatic Removal) ---
# Removing X48hr_50_R3 and X96hr_10_R2 due to QC issues
featureCounts <- featureCounts_raw %>%
  dplyr::select(!c("X48hr_50_R3.mLb.clN.bam", "X96hr_10_R2.mLb.clN.bam"))

# --- 3. Load Annotation File ---
annotate_peaks_raw <- read.delim(gzfile("../data/A172_GBM_ATAC_seq/GSM9825200_consensus_peaks.mLb.clN.annotatePeaks.txt.gz"),
  header = TRUE, sep = "\t"
)

# Rename first column to 'id' to match downstream logic
colnames(annotate_peaks_raw)[1] <- "id"
```


2. Prepare Data for DESeq2
```{r}
# --- 1. Create Count Matrix ---
# Separate metadata (cols 1-6) from counts (cols 7-end)
count_matrix <- as.matrix(featureCounts[, 7:ncol(featureCounts)])
rownames(count_matrix) <- featureCounts$Geneid

# --- 2. Create Metadata (colData) ---
# Extract sample names from matrix columns
sample_names <- colnames(count_matrix)

# Create a dataframe for parsing
AS_cell_metadata_df <- data.frame(sample_id = sample_names)

# Parse string: "X48hr_0_R1.mLb.clN.bam" -> "48hr", "0", "R1"
AS_cell_metadata_df <- AS_cell_metadata_df %>%
  mutate(
    temp_name = gsub("X", "", sample_id), # Remove leading X
    temp_name = gsub(".mLb.clN.bam", "", temp_name) # Remove suffix
  ) %>%
  separate(temp_name, into = c("hours", "TMZ_concentration", "replicate_number"), sep = "_") %>%
  mutate(
    hours_and_TMZ = paste(hours, TMZ_concentration, sep = "_")
  )

# Set Factor Levels (Control Reference is 48hr_0)
AS_cell_metadata_df$TMZ_concentration <- factor(AS_cell_metadata_df$TMZ_concentration, levels = c("0", "10", "50", "100"))
AS_cell_metadata_df$hours <- factor(AS_cell_metadata_df$hours, levels = c("48hr", "96hr"))
AS_cell_metadata_df$replicate_number <- factor(AS_cell_metadata_df$replicate_number, levels = c("R1", "R2", "R3", "R4"))
AS_cell_metadata_df$hours_and_TMZ <- factor(AS_cell_metadata_df$hours_and_TMZ,
  levels = c(
    "48hr_0", "48hr_10", "48hr_50", "48hr_100",
    "96hr_0", "96hr_10", "96hr_50", "96hr_100"
  )
)

# Verify alignment
if (!all(rownames(AS_cell_metadata_df$sample_id) == colnames(count_matrix))) {
  stop("Sample names do not match count matrix columns!")
}
```


3. Run DESeq2
```{r}
# Create DESeq DataSet
AS_dds <- DESeqDataSetFromMatrix(
  countData = count_matrix,
  colData = AS_cell_metadata_df,
  design = ~ hours_and_TMZ + replicate_number
)

# Run Differential Analysis
AS_atacDDS <- DESeq(AS_dds)
```

4. Extract Specific Contrasts
```{r}
# --- 48 Hour Comparisons ---
ds_res_dose_48_10 <- results(AS_atacDDS, c("hours_and_TMZ", "48hr_10", "48hr_0"))
ds_res_dose_48_10$term <- "48_10"
ds_res_dose_48_10$id <- rownames(ds_res_dose_48_10)

ds_res_dose_48_50 <- results(AS_atacDDS, c("hours_and_TMZ", "48hr_50", "48hr_0"))
ds_res_dose_48_50$term <- "48_50"
ds_res_dose_48_50$id <- rownames(ds_res_dose_48_50)

ds_res_dose_48_100 <- results(AS_atacDDS, c("hours_and_TMZ", "48hr_100", "48hr_0"))
ds_res_dose_48_100$term <- "48_100"
ds_res_dose_48_100$id <- rownames(ds_res_dose_48_100)

# --- 96 Hour Comparisons ---
ds_res_dose_96_10 <- results(AS_atacDDS, c("hours_and_TMZ", "96hr_10", "96hr_0"))
ds_res_dose_96_10$term <- "96_10"
ds_res_dose_96_10$id <- rownames(ds_res_dose_96_10)

ds_res_dose_96_50 <- results(AS_atacDDS, c("hours_and_TMZ", "96hr_50", "96hr_0"))
ds_res_dose_96_50$term <- "96_50"
ds_res_dose_96_50$id <- rownames(ds_res_dose_96_50)

ds_res_dose_96_100 <- results(AS_atacDDS, c("hours_and_TMZ", "96hr_100", "96hr_0"))
ds_res_dose_96_100$term <- "96_100"
ds_res_dose_96_100$id <- rownames(ds_res_dose_96_100)

# Combine and Filter for Significance
AS_ds_res_dose <- rbind(
  ds_res_dose_48_10, ds_res_dose_48_50, ds_res_dose_48_100,
  ds_res_dose_96_10, ds_res_dose_96_50, ds_res_dose_96_100
) %>%
  as.data.frame()

AS_ds_res_dose$padj <- p.adjust(AS_ds_res_dose$pvalue, method = "BH")

# Extract Significant IDs (DAS)
AS_das <- AS_ds_res_dose %>%
  filter(padj < 0.05) %>%
  pull(id) %>%
  unique()

print(paste("Number of Significant Peaks:", length(AS_das)))

# Create a supplementary table dataframe keeping all columns without modifying AS_ds_res_dose
supp_table_DAS <- AS_ds_res_dose %>%
  filter(padj < 0.05) %>%
  arrange(id)

# Check that all IDs in supp_table_DAS have a match in AS_das
ids_match_check <- all(supp_table_DAS$id %in% AS_das)
print(paste("Do all IDs in supp_table_DAS have a match in AS_das?", ids_match_check))

# Save the supplementary table to CSV
supp_table_out_dir <- supp_dir
if (!dir.exists(supp_table_out_dir)) dir.create(supp_table_out_dir, recursive = TRUE)
write.csv(supp_table_DAS, file.path(supp_table_out_dir, "DAS_all_sig_peaks.csv"), row.names = FALSE)
```


5. Group-Averaged Heatmap
```{r}
## 5. Heatmap Generation (High-Contrast / Rasterized for Publication)

library(ComplexHeatmap)

# 1. Get VST Data
vst_data <- vst(AS_atacDDS, blind = FALSE)
vst_matrix <- assay(vst_data)

# 2. Subset for Significant Peaks (DAS)
das_vst_matrix <- vst_matrix[rownames(vst_matrix) %in% AS_das, ]

# 3. Calculate Mean VST per Condition
condition_map <- setNames(AS_cell_metadata_df$hours_and_TMZ, AS_cell_metadata_df$sample_id)
unique_conditions <- levels(AS_cell_metadata_df$hours_and_TMZ)

das_vst_mean <- sapply(unique_conditions, function(cond) {
  samples <- names(condition_map)[condition_map == cond]
  if (length(samples) > 1) {
    rowMeans(das_vst_matrix[, samples])
  } else {
    das_vst_matrix[, samples]
  }
})

# 4. Z-Score Scale
z_score_matrix <- t(scale(t(das_vst_mean)))

# 5. Define Column Order Explicitly
col_order <- c(
  "48hr_0", "48hr_10", "48hr_50", "48hr_100",
  "96hr_0", "96hr_10", "96hr_50", "96hr_100"
)
z_score_matrix <- z_score_matrix[, col_order]

# 6. Plot Heatmap for Adobe Illustrator
# CHANGE: Using the pdf() "sandwich" to force the 20-inch height and save properly.
pdf_filename <- file.path(output_dir, "heatmap_final_vst_for_illustrator.pdf")

# Start recording to PDF
pdf(pdf_filename, width = 4, height = 5)

# Draw the heatmap
ComplexHeatmap::pheatmap(z_score_matrix,
  cluster_cols = FALSE,
  show_rownames = FALSE,
  clustering_method = "ward.D2",
  treeheight_row = 15,
  fontsize = 8,
  # NEW: Descriptive label for your legend
  name = "Relative\nAccessibility",
  color = colorRampPalette(c("navy", "white", "red"))(50),
  breaks = seq(-2, 2, length.out = 50),
  cutree_rows = 4,
  gaps_col = 4,
  use_raster = TRUE,
  raster_quality = 10,
  main = "Differential Accessibility\n(Grouped Mean VST)"
)

# Stop recording and save the file
dev.off()
```

6. Verification & Annotation Merge
```{r}
## 6. Verification & Annotation Merge

# Create the summary dataframe you asked to check
# We merge the metadata (Chr, Start, End) with the DAS list

# 1. Prepare IDs (Source: featureCounts)
# CHANGE: We ONLY keep the ID here. We discard featureCounts coordinates
# because they have a 1-bp mismatch with your preferred annotation format.
das_ids_only <- featureCounts %>%
  dplyr::filter(Geneid %in% AS_das) %>%
  dplyr::select(Geneid) %>%
  dplyr::rename(id = Geneid)

# 2. Merge with Full Annotation (Source: annotatePeaks.txt)
# CHANGE: We join the IDs with the FULL annotation file.
# This ensures Chr, Start, End come from the annotation file (Start.y in your screenshot).
merged_DAS_summary_new <- left_join(das_ids_only, annotate_peaks_raw, by = "id")

# Save Summary Table to Figures Folder
summary_file <- file.path(output_dir, "updated_DAS_summary.RDS")
saveRDS(merged_DAS_summary_new, summary_file)

# --- 2b. Export BED files ---
bed_dir <- file.path(output_dir, "FINAL_bed_files_for_genome_browser_3_12_26")
if (!dir.exists(bed_dir)) dir.create(bed_dir, recursive = TRUE)

# Export merged_DAS_summary_new as BED format
bed_das <- merged_DAS_summary_new %>%
  dplyr::select(Chr, Start, End, id) %>%
  dplyr::rename(chr = Chr, start = Start, end = End)

write.table(bed_das,
  file.path(bed_dir, "3_12_26_diff_accessible_peaks_FINAL_for_genomic_browser.bed"),
  sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE
)

# Export annotate_peaks_raw as BED format
bed_raw <- annotate_peaks_raw %>%
  dplyr::select(Chr, Start, End, id) %>%
  dplyr::rename(chr = Chr, start = Start, end = End)

write.table(bed_raw,
  file.path(bed_dir, "3_12_26_all_peaks_FINAL_for_genomic_browser.bed"),
  sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE
)


```


Genomic Annotation Distribution for DAS
```{r}
## 7. Genomic Annotation Distribution (High-Saturation Pie Chart)

library(ggplot2)
library(stringr)

# 1. Simplify Annotations
# CHANGED: Keeps the original 'promoter-TSS' label as requested
merged_DAS_summary_new <- merged_DAS_summary_new %>%
  mutate(Simple_Annotation = str_extract(Annotation, "^[^ ]+"))

# 2. Calculate Percentages for the Table
annotation_stats <- merged_DAS_summary_new %>%
  group_by(Simple_Annotation) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)

# 3. Save the actual table for text referencing
# Filename: DAS_annotation_distribution_table.csv
write.csv(annotation_stats, file.path(supp_dir, "DAS_annotation_distribution_table.csv"), row.names = FALSE)

# 4. Create the Pie Chart
# - CHANGED: Updated title to "Genomic Distribution of Differentially Accessible Sites"
# - CHANGED: Switched to 'Set1' for maximum color saturation.
# - REMOVED: geom_text labels for a cleaner, publication-ready look.
p_pie <- ggplot(annotation_stats, aes(x = "", y = percentage, fill = Simple_Annotation)) +
  geom_bar(stat = "identity", width = 1, color = "white", linewidth = 0.1) +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Differentially\nAccessible Sites",
    fill = "Genomic Feature"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 5),
    legend.title = element_text(face = "bold", size = 5),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3, "mm"),
    legend.spacing.y = unit(1, "mm")
  )

# 5. Save the final PDFs
# Version 1: 4x4 with Legend
ggsave(file.path(main_fig_dir, "DAS_annotation_distribution_pie_chart.pdf"), 
       plot = p_pie, width = 4, height = 4)

# Version 2: 0.75x0.75 NO Legend
ggsave(file.path(main_fig_dir, "DAS_annotation_distribution_pie_chart_no_legend.pdf"), 
       plot = p_pie + theme(legend.position = "none"), width = 0.75, height = 0.75)
```


Genomic Annotation Distribution for all accessible sites
```{r}
## 8. Total Peak Distribution (All nf-core peaks)

# 1. Simplify Annotations for the full raw dataset
# Keeps 'promoter-TSS' as requested
annotate_peaks_raw <- annotate_peaks_raw %>%
  mutate(Simple_Annotation = str_extract(Annotation, "^[^ ]+"))

# 2. Calculate Percentages for the Total Dataset
total_annotation_stats <- annotate_peaks_raw %>%
  group_by(Simple_Annotation) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)

# 3. Save the actual table for the full peakset
write.csv(total_annotation_stats,
  file.path(supp_dir, "All_accessible_Peaks_annotation_distribution_table.csv"),
  row.names = FALSE
)

# 4. Create the Pie Chart
p_total_pie <- ggplot(total_annotation_stats, aes(x = "", y = percentage, fill = Simple_Annotation)) +
  geom_bar(stat = "identity", width = 1, color = "white", linewidth = 0.1) +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "All Accessible\nSites",
    fill = "Genomic Feature"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 5),
    legend.title = element_text(face = "bold", size = 5),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3, "mm"),
    legend.spacing.y = unit(1, "mm")
  )

# 5. Save the final PDFs
# Version 1: 4x4 with Legend
ggsave(file.path(main_fig_dir, "All_accessible_peaks_annotation_distribution_pie_chart.pdf"), 
       plot = p_total_pie, width = 4, height = 4)

# Version 2: 0.75x0.75 NO Legend
ggsave(file.path(main_fig_dir, "All_accessible_peaks_annotation_distribution_pie_chart_no_legend.pdf"), 
       plot = p_total_pie + theme(legend.position = "none"), width = 0.75, height = 0.75)
```

## 9. Experimental: Heatmap with 2Mb Proximity Annotation

```{r}
library(GenomicFeatures)
library(org.Hs.eg.db)
library(ComplexHeatmap)
library(circlize)

# 1. Target genes and MANE loader
target_genes <- c("MSH2", "MSH6", "MLH1", "PMS2")
mane_gff_path <- "../data/MANE/MANE.GRCh38.v1.5.refseq_genomic.gff.gz"

if (!exists("txdb_mane")) {
  txdb_mane <- makeTxDbFromGFF(mane_gff_path, format = "gff")
  seqlevelsStyle(txdb_mane) <- "UCSC"
}

# robust gene lookup helper
get_real_db_key <- function(symbol, txdb) {
  all_keys <- keys(txdb, keytype = "GENEID")
  if (symbol %in% all_keys) {
    return(symbol)
  }
  entrez_id <- suppressMessages(mapIds(org.Hs.eg.db, keys = symbol, column = "ENTREZID", keytype = "SYMBOL", multiVals = "first"))
  if (!is.na(entrez_id)) {
    if (entrez_id %in% all_keys) {
      return(entrez_id)
    }
    if (paste0("GeneID:", entrez_id) %in% all_keys) {
      return(paste0("GeneID:", entrez_id))
    }
  }
  return(NULL)
}

# 2. Get 2Mb windows for targets
target_ranges <- GRanges()
for (g in target_genes) {
  db_key <- get_real_db_key(g, txdb_mane)
  if (!is.null(db_key)) {
    g_range <- genes(txdb_mane)[db_key]
    midpoint <- (start(g_range) + end(g_range)) / 2
    chrom <- as.character(seqnames(g_range))[1]
    w_start <- floor(midpoint - 1000000)
    w_end <- floor(midpoint + 1000000)
    tmp_gr <- GRanges(chrom, IRanges(w_start, w_end))
    tmp_gr$symbol <- g
    target_ranges <- c(target_ranges, tmp_gr)
  }
}

# 3. Calculate total accessible peaks per gene (Regardless of DAS status)
# This block is independent and does not affect the heatmap or DAS calculations.
all_peaks_gr <- GRanges(
  seqnames = annotate_peaks_raw$Chr,
  ranges = IRanges(start = annotate_peaks_raw$Start, end = annotate_peaks_raw$End),
  id = annotate_peaks_raw$id
)

# Create a separate target ranges object just for baseline calculations to not affect downstream code
baseline_target_ranges <- target_ranges
msh2_idx <- which(baseline_target_ranges$symbol == "MSH2")
if (length(msh2_idx) > 0) {
  # Hardcode MSH2 2Mb region for baseline accessibility only
  baseline_target_ranges[msh2_idx] <- GRanges("chr2", IRanges(46533132, 48533132), symbol = "MSH2")
}

overlaps_baseline <- findOverlaps(all_peaks_gr, baseline_target_ranges)

print("--- Baseline Accessibility (All Peaks) Within 2Mb Windows ---")
# Initialize an empty list to store rows for the dataframe
baseline_peaks_list_safe <- list()
# Initialize a vector to track ALL unique peak IDs across all target genes
all_unique_peak_hits_safe <- c()

for (g in target_genes) {
  gene_range <- baseline_target_ranges[baseline_target_ranges$symbol == g]
  if (length(gene_range) > 0) {
    coords_str <- paste0(seqnames(gene_range), ":", start(gene_range), "-", end(gene_range))
    target_idx <- which(baseline_target_ranges$symbol == g)
    gene_hits <- queryHits(overlaps_baseline)[subjectHits(overlaps_baseline) == target_idx]
    unique_peaks <- length(unique(gene_hits))
    
    # Store data
    baseline_peaks_list_safe[[g]] <- data.frame(
        Gene = g,
        Coordinates = coords_str,
        Accessible_Peaks_Within_2Mb = unique_peaks,
        stringsAsFactors = FALSE
    )
    
    # Track all peak hits to find total unique across all genes
    all_unique_peak_hits_safe <- c(all_unique_peak_hits_safe, gene_hits)
    
    print(paste0(g, " [", coords_str, "]: ", unique_peaks, " accessible peaks within 2 Mb"))
  }
}

# Bind list into a single dataframe
baseline_accessible_peaks_df_safe <- do.call(rbind, baseline_peaks_list_safe)
rownames(baseline_accessible_peaks_df_safe) <- NULL

# Print the total unique peaks across all regions (accounting for overlapping 2Mb windows)
total_unique_across_all_safe <- length(unique(all_unique_peak_hits_safe))
print(paste0("Total unique accessible peaks across all combined 2Mb regions: ", total_unique_across_all_safe))

# 4. Convert DAS to GRanges and find overlaps
das_gr <- GRanges(
  seqnames = merged_DAS_summary_new$Chr,
  ranges = IRanges(start = merged_DAS_summary_new$Start, end = merged_DAS_summary_new$End),
  id = merged_DAS_summary_new$id
)

overlaps <- findOverlaps(das_gr, target_ranges)

# 4. Binary Matrix mapping DAS to Genes
das_gene_mat <- matrix(0, nrow = length(das_gr), ncol = length(target_genes))
rownames(das_gene_mat) <- das_gr$id
colnames(das_gene_mat) <- target_genes

for (i in seq_along(overlaps)) {
  das_idx <- queryHits(overlaps)[i]
  gene_name <- target_ranges$symbol[subjectHits(overlaps)[i]]
  col_idx <- which(colnames(das_gene_mat) == gene_name)
  das_gene_mat[das_idx, col_idx] <- 1
}

# align strictly to heatmap
das_gene_mat_aligned <- das_gene_mat[rownames(z_score_matrix), ]

# 5. Extract clusters from z_score_matrix
set.seed(123)
row_dend <- hclust(dist(z_score_matrix), method = "ward.D2")
row_clusters <- cutree(row_dend, k = 4)

# 6. Sum frequencies per cluster
cluster_freq <- matrix(0, nrow = 4, ncol = length(target_genes))
colnames(cluster_freq) <- target_genes
for (k in 1:4) {
  cluster_ids <- names(row_clusters)[row_clusters == k]
  if (length(cluster_ids) == 1) {
    cluster_freq[k, ] <- das_gene_mat_aligned[cluster_ids, ]
  } else {
    cluster_freq[k, ] <- colSums(das_gene_mat_aligned[cluster_ids, ])
  }
}

# 7. Map back to block annotation matrix for the right
right_anno_mat <- matrix(0, nrow = nrow(z_score_matrix), ncol = length(target_genes))
colnames(right_anno_mat) <- target_genes
rownames(right_anno_mat) <- rownames(z_score_matrix)

for (i in 1:nrow(z_score_matrix)) {
  c_id <- row_clusters[i]
  right_anno_mat[i, ] <- cluster_freq[c_id, ]
}

# 8. Plot side-by-side using native ComplexHeatmap
max_val <- max(right_anno_mat)

# Custom palette: exactly 0 is white, >0 transitions into viridis
library(viridis)
my_colors <- c("white", viridis::viridis(49))

color_fun <- colorRamp2(seq(0, max(1, max_val), length = 50), my_colors)

main_col_fun <- colorRamp2(seq(-2, 2, length.out = 50), colorRampPalette(c("navy", "white", "red"))(50))

ht_main <- Heatmap(z_score_matrix,
  name = "Relative\nAccessibility",
  col = main_col_fun,
  cluster_columns = FALSE,
  cluster_rows = row_dend,
  show_row_names = FALSE,
  row_dend_width = unit(10, "mm"),
  row_split = 4,
  column_split = factor(rep(c("48hr", "96hr"), each = 4), levels = c("48hr", "96hr")),
  column_title = "DAS Grouped Mean VST",
  row_title = NULL,
  width = unit(4, "cm"),
  use_raster = TRUE,
  raster_quality = 10,
  heatmap_legend_param = list(title_position = "topleft")
)

ht_anno <- Heatmap(right_anno_mat,
  name = "DAS\nwithin 2Mb",
  col = color_fun,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  column_names_side = "top",
  column_names_rot = 90,
  column_names_gp = gpar(fontsize = 7),
  column_split = factor(target_genes, levels = target_genes),
  column_title = NULL,
  column_gap = unit(1, "mm"),
  width = unit(1.3, "cm"),
  use_raster = TRUE,
  raster_quality = 10,
  heatmap_legend_param = list(direction = "vertical")
)

pdf_filename_anno <- file.path(output_dir, "DAS_heatmap_2Mb_genes_of_interest_no_cluster_label.pdf")
pdf(pdf_filename_anno, width = 7, height = 5)
draw(ht_main + ht_anno, main_heatmap = "Relative\nAccessibility")
dev.off()
```

## 10. Cluster Validation and Export

```{r}
# --- Create Validation Dataframes ---

# 1. Full Peak-Level Binary Matrix with Cluster IDs
# We use the raw statistical cluster assignments directly from cutree
peak_cluster_df <- as.data.frame(das_gene_mat_aligned)
peak_cluster_df$Cluster_ID <- row_clusters
peak_cluster_df$Peak_ID <- rownames(peak_cluster_df)

# Reorder columns to put Peak_ID and Cluster_ID first
peak_cluster_df <- peak_cluster_df[, c("Peak_ID", "Cluster_ID", target_genes)]

# Save to CSV
write.csv(peak_cluster_df,
  file.path(output_dir, "DAS_Peaks_2Mb_Proximity_with_Clusters.csv"),
  row.names = FALSE
)

# Calculate and print the number of unique DAS within 2Mb of specific genes
# Uses the target_genes variable already defined (MSH2, MSH6, MLH1, PMS2)
# DAS that overlap more than one gene of interest are only counted once in the total
unique_mmr_das_count <- sum(rowSums(peak_cluster_df[, target_genes, drop = FALSE]) > 0)
print(paste("Total unique DAS within 2Mb of MMR genes (MSH2, MSH6, MLH1, PMS2):", unique_mmr_das_count))

# 2. Aggregated Summary Dataframe (Genes by Cluster)
cluster_summary_df <- as.data.frame(cluster_freq)
cluster_summary_df$Cluster_ID <- paste0("Cluster_", 1:4)

# Reorder to put Cluster_ID first
cluster_summary_df <- cluster_summary_df[, c("Cluster_ID", target_genes)]

# Add Total row
total_row <- data.frame(Cluster_ID = "Total")
for (g in target_genes) {
  total_row[[g]] <- sum(cluster_summary_df[[g]])
}
cluster_summary_df <- rbind(cluster_summary_df, total_row)

# Save to CSV
write.csv(cluster_summary_df,
  file.path(supp_dir, "Cluster_Summary_DAS_2Mb_Proximity.csv"),
  row.names = FALSE
)


# --- Plot Duplicate Heatmap with Explicit Cluster Labels ---

# Provide a categorical color mapping for the 4 statistical clusters
cluster_colors <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "4" = "#984EA3")
left_anno <- rowAnnotation(
  Cluster = as.character(row_clusters),
  col = list(Cluster = cluster_colors),
  show_annotation_name = FALSE,
  simple_anno_size = unit(2, "mm"),
  annotation_legend_param = list(
    title_gp = gpar(fontsize = 7, fontface = "bold"),
    labels_gp = gpar(fontsize = 7)
  )
)

# We use the exact same matrices and splits, but tell ComplexHeatmap to print the categorical color block on the left
ht_main_labeled <- Heatmap(z_score_matrix,
  name = "Relative\nAccessibility",
  col = main_col_fun,
  cluster_columns = FALSE,
  cluster_rows = row_dend,
  show_row_names = FALSE,
  row_dend_width = unit(7, "mm"),
  row_split = 4,
  left_annotation = left_anno,
  row_title = NULL,
  column_split = factor(rep(c("48hr", "96hr"), each = 4), levels = c("48hr", "96hr")),
  column_title = "DAS Grouped Mean VST",
  column_title_gp = gpar(fontsize = 7, fontface = "bold"),
  column_names_gp = gpar(fontsize = 7),
  width = unit(3, "cm"),
  use_raster = TRUE,
  raster_quality = 10,
  heatmap_legend_param = list(
    title_position = "topleft",
    title_gp = gpar(fontsize = 7, fontface = "bold"),
    labels_gp = gpar(fontsize = 7)
  )
)

ht_anno_labeled <- Heatmap(right_anno_mat,
  name = "DAS\nwithin 2Mb",
  col = color_fun,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  column_names_side = "top",
  column_names_rot = 90,
  column_names_gp = gpar(fontsize = 7),
  column_split = factor(target_genes, levels = target_genes),
  column_title = NULL,
  column_gap = unit(1, "mm"),
  width = unit(1.1, "cm"),
  use_raster = TRUE,
  raster_quality = 10,
  heatmap_legend_param = list(
    direction = "vertical",
    title_gp = gpar(fontsize = 7, fontface = "bold"),
    labels_gp = gpar(fontsize = 7)
  )
)


pdf_filename_val <- file.path(main_fig_dir, "Main_fig_DAS_heatmap_2Mb_genes_of_interest_CLUSTER_LABELS.pdf")
pdf(pdf_filename_val, width = 7, height = 5) # Made slightly wider to fit the text labels
draw(ht_main_labeled + ht_anno_labeled, 
     main_heatmap = "Relative\nAccessibility",
     ht_gap = unit(0.5, "mm"))
dev.off()
print("Validation dataframes saved and labeled heatmap plotted!")
```



## 12. Patient Data Consensus Peaks
```{r}
library(GenomicRanges)

# --- 1. Define Directories & Files ---
patient_dir <- "../data/patient_data/GSE174554"
patient_bed_out_dir <- file.path(output_dir, "Patient_data_consensus_peak_sets")

if (!dir.exists(patient_bed_out_dir)) dir.create(patient_bed_out_dir, recursive = TRUE)

# Find properly formatted patient BED files
patient_bed_files <- list.files(patient_dir, pattern = "_properly_formatted\\.bed$", full.names = TRUE)

# Separate by Primary vs Recurrent
primary_files <- patient_bed_files[grepl("Primary", basename(patient_bed_files))]
recurrent_files <- patient_bed_files[grepl("Recurrent", basename(patient_bed_files))]

# --- 2. Load BED files into GRanges lists ---
load_bed_as_granges <- function(file_path) {
  df <- read.table(file_path, header = FALSE, sep = "\t", stringsAsFactors = FALSE)
  # Standard BED has chr in V1, start in V2, end in V3.
  # BED is 0-indexed, so we add 1 to the start to convert to 1-indexed GRanges
  GRanges(seqnames = df$V1, ranges = IRanges(start = df$V2 + 1, end = df$V3))
}

primary_beds <- lapply(primary_files, load_bed_as_granges)
recurrent_beds <- lapply(recurrent_files, load_bed_as_granges)
all_patient_beds <- c(primary_beds, recurrent_beds)

# --- 3. Consensus Function ---
make_consensus_set <- function(gr_list, min_patients = 1, gap_to_merge = 1) {
  # Flatten all peaks into one massive pool
  all_peaks <- unlist(GRangesList(gr_list))

  # Calculate coverage: creates an Rle object of peak depth per base pair
  cvg <- coverage(all_peaks)

  # Filter for 'consensus': keep segments where coverage >= min_patients
  # Prevent collision with dplyr::slice by explicitly calling IRanges
  consensus_rle <- IRanges::slice(cvg, lower = min_patients)

  # Convert back to GRanges
  consensus_gr <- as(consensus_rle, "GRanges")

  # Always run reduce() to merge directly adjacent fragments.
  # Note: min.gapwidth = 1 is required to merge intervals that are perfectly continuous
  # (e.g., end of interval A is the exact same coordinate as start of interval B)
  consensus_gr <- reduce(consensus_gr, min.gapwidth = gap_to_merge)

  return(consensus_gr)
}

# Create the three reference sets
primary_consensus <- make_consensus_set(primary_beds, min_patients = 1, gap_to_merge = 1)
recurrent_consensus <- make_consensus_set(recurrent_beds, min_patients = 1, gap_to_merge = 1)
global_consensus <- make_consensus_set(all_patient_beds, min_patients = 1, gap_to_merge = 1)

# --- 4. Export to BED ---
export_consensus_bed <- function(gr, filepath) {
  # Convert to standard BED format (0-indexed start)
  df <- data.frame(
    chr = as.character(seqnames(gr)),
    start = start(gr) - 1,
    end = end(gr)
  )
  write.table(df, file = filepath, sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
}

export_consensus_bed(primary_consensus, file.path(patient_bed_out_dir, "Patient_Primary_Consensus_min1.bed"))
export_consensus_bed(recurrent_consensus, file.path(patient_bed_out_dir, "Patient_Recurrent_Consensus_min1.bed"))
export_consensus_bed(global_consensus, file.path(patient_bed_out_dir, "Patient_Global_Consensus_min1.bed"))

print(paste("Generated 3 patient consensus BED files in:", patient_bed_out_dir))
```

## 13. Overlap Analysis against Patient Data
```{r}
# --- 1. Convert to GRanges ---
# Convert the annotate_peaks_raw dataframe to a GRanges object
my_peaks_gr <- GRanges(
  seqnames = annotate_peaks_raw$Chr,
  ranges = IRanges(start = annotate_peaks_raw$Start, end = annotate_peaks_raw$End),
  id = annotate_peaks_raw$id
)

# --- 2. Find Overlaps ---
# The findOverlaps function finds regions that intersect by at least 1 base pair.
# Because coordinates won't match perfectly, this is the standard way to check
# if a peak in your data "hits" a peak in the patient data.
overlaps <- findOverlaps(my_peaks_gr, global_consensus)

# Extract counts
# queryHits tells us which of 'my_peaks_gr' had at least one overlap
my_peaks_overlapping <- length(unique(queryHits(overlaps)))
my_peaks_unique <- length(my_peaks_gr) - my_peaks_overlapping

# subjectHits tells us which of 'global_consensus' had at least one overlap
patient_peaks_overlapping <- length(unique(subjectHits(overlaps)))
patient_peaks_unique <- length(global_consensus) - patient_peaks_overlapping

print(paste("Your total peaks:", length(my_peaks_gr)))
print(paste("Patient total peaks:", length(global_consensus)))
print(paste("My peaks that overlap a patient peak (from my peak perspective, mutltiple patient peaks can fall in one of my peaks):", my_peaks_overlapping))
print(paste("Patient peaks that overlap your peak (from patient peak perspective):", patient_peaks_overlapping))

# Add percentage calculations
print(paste0("Percentage of your peaks that overlap at least one patient peak: ", 
             round((my_peaks_overlapping / length(my_peaks_gr)) * 100, 2), "%"))
print(paste0("Percentage of patient peaks that overlap your peaks: ", 
             round((patient_peaks_overlapping / length(global_consensus)) * 100, 2), "%"))

# Note: Because 1 of your broad peaks might overlap 2 smaller patient peaks,
# the mathematical "intersection" size is slightly nuanced. A standard approach
# for a simple proportional Venn Diagram is to use the count of your overlapping peaks.

# --- 3. Plot Venn Diagram ---
# We will use the 'eulerr' package for accurate proportional Euler/Venn diagrams
if (!requireNamespace("eulerr", quietly = TRUE)) {
  install.packages("eulerr")
}
library(eulerr)

# --- 3. Plot Proportional Venn Diagram ---
# CHANGE: Swapping to patient_peaks_overlapping perspective
my_peaks_unique_perspective <- length(my_peaks_gr) - patient_peaks_overlapping

fit <- euler(c(
  "My Peaks" = my_peaks_unique_perspective,
  "Patient Consensus" = patient_peaks_unique,
  "My Peaks&Patient Consensus" = patient_peaks_overlapping
))

venn_plot <- plot(fit,
  quantities = list(fontsize = 7),
  labels = FALSE,
  fills = list(fill = c("#A5EDFF", "#CCBFFF", "#B2FF8C"), alpha = 0.6),
  main = NULL
)

# Save Proportional Version
pdf_filename_venn <- file.path(main_fig_dir, "My_Peaks_vs_Patient_Consensus_Venn.pdf")
pdf(pdf_filename_venn, width = 3, height = 1.5)
print(venn_plot)
library(grid)
grid.text("Our Peaks", x = 0.35, y = 0.9, gp = gpar(fontsize = 7, fontface = "bold"))
grid.text("Patient Peaks", x = 0.80, y = 0.9, gp = gpar(fontsize = 7, fontface = "bold"))
dev.off()


# --- 4. Plot Non-Proportional Venn Diagram (Equal Sized Circles) ---
fit_same_size <- euler(c("A" = 1, "B" = 1, "A&B" = 0.45))

venn_plot_same_size <- plot(fit_same_size,
  quantities = list(
    labels = c(
      as.character(my_peaks_unique_perspective), 
      as.character(patient_peaks_unique), 
      as.character(patient_peaks_overlapping)
    ),
    fontsize = 7
  ),
  labels = FALSE,
  fills = list(fill = c("#A5EDFF", "#CCBFFF", "#B2FF8C"), alpha = 0.6),
  main = NULL
)

# Save Same-Size Version
pdf_filename_venn_same <- file.path(main_fig_dir, "same_size_circle_venn_My_Peaks_vs_Patient_Consensus.pdf")
pdf(pdf_filename_venn_same, width = 2, height = 1)
print(venn_plot_same_size)
grid.text("Our Peaks", x = 0.30, y = 0.9, gp = gpar(fontsize = 7, fontface = "bold"))
grid.text("Patient Peaks", x = 0.75, y = 0.9, gp = gpar(fontsize = 7, fontface = "bold"))
dev.off()

print("Saved both proportional and same-size Venn diagrams!")
```

## 14. MMR Gene Baseline Accessibility (500 kb Windows)
# This chunk calculates total accessibility in a 500kb window centered on MMR genes,
# matching the exact coordinates used in the 'dense_and_full_gene_tracks_MMR.Rmd' script.

```{r}
# 1. Define target genes and 500kb window logic
target_genes_500k <- c("MSH2", "MSH6", "MLH1", "PMS2")
half_window <- 250000

# 2. Create the 500kb windows
target_ranges_500k <- GRanges()
for (g in target_genes_500k) {
  db_key <- get_real_db_key(g, txdb_mane) # Uses helper from Section 9
  if (!is.null(db_key)) {
    g_range <- genes(txdb_mane)[db_key]
    midpoint <- (start(g_range) + end(g_range)) / 2
    chrom <- as.character(seqnames(g_range))[1]
    
    # Applying the floor(midpoint - 250000) and floor(midpoint + 249999) logic
    w_start <- floor(midpoint - half_window)
    w_end <- floor(midpoint + (half_window - 1))
    
    tmp_gr <- GRanges(chrom, IRanges(w_start, w_end))
    tmp_gr$symbol <- g
    target_ranges_500k <- c(target_ranges_500k, tmp_gr)
  }
}

# 3. Find overlaps with the full peak set
# (all_peaks_gr was already defined in Section 9)
overlaps_500k <- findOverlaps(all_peaks_gr, target_ranges_500k)

print("--- Baseline Accessibility (All Peaks) Within 500 kb Windows ---")
for (g in target_genes_500k) {
  gene_range <- target_ranges_500k[target_ranges_500k$symbol == g]
  if (length(gene_range) > 0) {
    coords_str <- paste0(seqnames(gene_range), ":", start(gene_range), "-", end(gene_range))
    target_idx <- which(target_ranges_500k$symbol == g)
    gene_hits <- queryHits(overlaps_500k)[subjectHits(overlaps_500k) == target_idx]
    unique_peaks <- length(unique(gene_hits))
    print(paste0(g, " [", coords_str, "]: ", unique_peaks, " accessible peaks within 500 kb"))
  }
}
```
